---
title: "Convert OD to titre (SERION assay)"
format: html
---

## Constant

Path to data:

```{r}
path2data <- paste0("C:/Users/ongph/github/data")
```

Files names:

```{r}
inp_sero <- "sero/raw/measles_elisa_Aug2025.xlsx"
```

## Packages

Required packages:

```{r}
# Edit this only
required <- c("tidyverse", "lubridate", "janitor", "openxlsx")
```

Installing those that are not installed yet:

```{r}
to_install <- required[! required %in% installed.packages()[, "Package"]]
if (length(to_install)) install.packages(to_install)
```

Loading the packages for interactive use:

```{r}
invisible(sapply(required, library, character.only = TRUE))
```

## Functions

Convert OD to titre

```{r}
calc_titre <- function(a, b, c, d, od, manustd, ourstd) {
  exp(c - log((d - a)/(od * manustd / ourstd - a) - 1) / b)
}
```

## Data

Read raw data:

```{r}
df <- file.path(path2data, inp_sero) |> 
  read.xlsx(sheet = 2) |> 
  clean_names()
```

## Convert OD to titre

Convert:

```{r}
df <- df |> 
  # Remove empty wells
  filter(!is.na(results)) |> 
  mutate(
    # Subtract blank
    neg = neg - blank,
    std_1 = std_1 - blank,
    std_2 = std_2 - blank,
    results = results - blank,
    # Compute mean of 2 standard samples
    std_mean = rowMeans(pick(starts_with("std")), na.rm = TRUE),
    # Convert to titre
    titre = calc_titre(
      a = a, 
      b = b, 
      c = c, 
      d = d, 
      od = results, 
      manustd = std_ref, 
      ourstd = std_mean
    )
  )
```

Determine cutoff value for positive:

```{r}
cutoff_val <- 200
```

Define sample positive

```{r}
df <- df |> 
  mutate(
    pos = if_else(titre > cutoff_val, 1, 0)
  )
```

## Quality control

The substrate blank must be OD < 0.25

```{r}
df |> 
  filter(blank >= 0.25)
```

The negative control must be negative

```{r}
df |> 
  distinct(neg, .keep_all = TRUE) |> 
  mutate(
    titre = calc_titre(
      a = a, 
      b = b, 
      c = c, 
      d = d, 
      od = neg, # convert the neg OD
      manustd = std_ref, 
      ourstd = std_mean
    ),
    pos = if_else(titre > cutoff_val, 1, 0)
  ) |> 
  # Check if any neg is pos
  filter(pos == 1)
```

The mean OD-value of the standard serum must be within the validity range

```{r}
df |> 
  filter(std_mean < lower | std_mean > upper)
```

## Clean metadata columns

```{r}
df <- df |> 
  mutate(
    across(c(dob, date_collect), dmy)
  )
```

## Save output files

Save file RDS

```{r}
df |> 
  saveRDS(
    file = file.path(
      path2data, 
      inp_sero |> 
        str_replace("sero/raw", "sero/cleaned") |> 
        str_replace("\\.xlsx$", ".rds")
    )
  )
```

Save file xlsx

```{r}
df |> 
  write.xlsx(
    file = file.path(
      path2data, 
      inp_sero |> str_replace("sero/raw", "sero/cleaned")
    ),
    rowNames = FALSE
  )
```


