---
title: "Result 1"
format: html
execute: 
  eval: false
editor_options: 
  chunk_output_type: console
---

## Constant

Path to data:

```{r}
path2data <- paste0("C:/Users/ongph/github/data")
```

Files names:

```{r}
# Case notification
cases_file <- "measles/cleaned/measles_hcmc_18-25.rds"
cases_neg_file <- "measles/cleaned/measles_hcmc_18-25_neg.rds"

# Serology
sero_file <- "sero/cleaned/measles_elisa_hcmc_children.rds"

# Vaccination coverage
## HCMC
vax_hcmc_file <- "vax/vaxreg_hcmc_measles.rds"
## Vietnam
vax_vn_file <- "vax/vaxvn_2023.xlsx"

# Population metadata
map_hcmc_file <- "map/gadm41_hcmc_district.rds"
map_vn_file <- "map/gadm41_vn.rds"
census_file <- "census2019_hcm.rds"
```

## Packages

Required packages:

```{r, message=FALSE}
required <- c("openxlsx", "janitor", "tidyverse", "scales", "mgcv", "stringi", "sf", "binom", "ggsci", "ggspatial", "patchwork", "zoo", "slider", "cowplot", "ggh4x", "gtable", "grid")

# Installing those that are not installed yet:
to_install <- required[! required %in% installed.packages()[, "Package"]]
if (length(to_install)) install.packages(to_install)

# Loading the packages for interactive use:
invisible(sapply(required, library, character.only = TRUE))

source("code/prop_query.R")
```

## Functions

Clean Vietnamese district and province name

```{r}
clean_district <- function(x) {
  x |>
    stri_trans_general("Latin-ASCII") |> 
    str_to_lower() |>                      
    str_squish() |>                        
    str_replace_all("đ", "d") |>
    str_remove("^(quan |huyen |thanh pho |tp )")
}

clean_province <- function(x) {
  x |>
    stri_trans_general("Latin-ASCII") |> 
    str_to_lower() |>
    str_squish() |>
    str_replace_all("đ", "d")
}
```

Correct seroprevalence by sensitivity and specificity

```{r}
prev_correct <- function(est_val, sens, spec) {
  (est_val + spec - 1) / (sens + spec - 1)
}
```

Predict with confidence interval

```{r}
predict_ci <- function(model, data, var, ci = 0.95, length = 1000) {
  p <- (1 - ci) / 2
  link_inv <- model$family$linkinv
  n <- nrow(data) - length(model$coefficients)
  vals <- seq(from = min(data[[var]], na.rm = T), to = max(data[[var]], na.rm = T), length.out = length)
  df <- data.frame(col = vals)
  colnames(df) <- var
  
  pred <- data.frame(predict(model, df, se.fit = TRUE))
  pred[[var]] <- df[[var]]
  pred$lwr <- link_inv(pred$fit + qt(p, n) * pred$se.fit)
  pred$upr <- link_inv(pred$fit + qt(1 - p, n) * pred$se.fit)
  pred$fit <- link_inv(pred$fit)
  pred
}

sero_ci <- function(df, key) {
  df |>
    summarise(n_sero = n(), x_pos = sum(pos), .by = all_of(key)) |>
    mutate(
      sero    = x_pos / n_sero,
      sero_lo = if_else(x_pos == 0, 0, qbeta(0.025, x_pos, n_sero - x_pos + 1)),
      sero_hi = if_else(x_pos == n_sero, 1, qbeta(0.975, x_pos + 1, n_sero - x_pos))
    ) |>
    rename(group = all_of(key))
}

sero_for <- function(lo, hi) {
  d <- sero_src |> filter(age_y >= lo, age_y < hi)
  bind_rows(
    sero_ci(d, "district") |>
      mutate(group = as.character(group), level = "District"),
    sero_ci(mutate(d, region = as.character(region)), "region") |>
      mutate(level = "Region")
  )
}
```

Turn districts to regions:

```{r}
district_region <- tribble(
  ~district,     ~region,
 
  "binh chanh",  "West",
  "binh tan",    "West",
 
  "12",          "North",
  "hoc mon",     "North",
  "cu chi",      "North",
 
  "thu duc",     "East",
  "2",           "East",        # former District 2, now Thu Duc City
  "9",           "East",        # former District 9, now Thu Duc City
 
  "7",           "South",
  "nha be",      "South",
  "can gio",     "South",
 
  "1",           "Central",
  "3",           "Central",
  "4",           "Central",
  "5",           "Central",
  "6",           "Central",
  "8",           "Central",
  "10",          "Central",
  "11",          "Central",
  "binh thanh",  "Central",
  "go vap",      "Central",
  "phu nhuan",   "Central",
  "tan binh",    "Central",
  "tan phu",     "Central"
)

lv_region <- c("West", "North", "East", "South", "Central")
```


## Data

Map:

```{r}
df_map_hcmc <- file.path(path2data, map_hcmc_file) |> 
  readRDS() |> 
  mutate(district_blank = clean_district(district))

df_map_vn <- file.path(path2data, map_vn_file) |> 
  readRDS() |> 
  mutate(province_blank = clean_province(province))
```

Census:

```{r}
df_census <- file.path(path2data, census_file) |> 
  readRDS()

df_census_age <- df_census |> 
  group_by(age) |>
  summarise(n_census = sum(n), .groups = "drop")

df_census_district <- df_census |> 
  group_by(district) |>
  summarise(n_census = sum(n), .groups = "drop") |> 
  mutate(
    district = clean_district(district)
  )
```

Case report data:

```{r}
df_case <- readRDS(file.path(path2data, cases_file))
```

Serology:

```{r}
df_sero <- file.path(path2data, sero_file) |> 
  readRDS() |> 
  mutate(district = clean_district(district))
```

Vaccination coverage

```{r}
df_vax_hcmc <- file.path(path2data, vax_hcmc_file) |> 
  readRDS() |> 
  mutate(district_blank = clean_district(district))

df_vax_vn <- file.path(path2data, vax_vn_file) |> 
  read.xlsx() |> 
  mutate(province_blank = clean_province(province))
```

## Analysis

## Case data

### Proportion by age and cohort

```{r}
col_admission <- "ngay_nv"
col_dob       <- "ngay_sinh"   # set to NA to fall back to year(admission) - age
 
# Week convention. 1 = weeks start Monday (ISO), 7 = Sunday (MMWR)
week_start <- 1
 
# Outbreak period rule
ob_threshold <- 5    # admissions per week
ob_min_run   <- 3    # consecutive weeks at or above the threshold
ob_max_gap   <- 8    # weeks below threshold tolerated within one period
ob_min_cases <- 50   # episodes smaller than this are treated as sporadic
 
# Outbreak response immunisation windows
shade_win <- tibble(
  xmin = dmy(c("01-12-2018", "01-09-2024")),
  xmax = dmy(c("31-05-2019", "30-04-2025"))
)
 
shade_label <- "Outbreak response\nimmunisation"
shade_fill  <- "grey70"
shade_alpha <- 0.35
 
age_cols <- c(
  "<1"    = "#08519C",
  "1-5"   = "#56B4E9",
  "5-10"  = "#FBD87F",
  "10-15" = "#B87400",
  ">15"   = "#009E73"
)
 
cohort_cols <- c(
  "<2009"     = "#999999",
  "2009-2013" = "#CC79A7",
  "2014-2018" = "#F0E442",
  "2019-2023" = "#0072B2",
  "2024"      = "#D55E00",
  "2025"      = "#000000"
)
 
# ---------------------------------------------------------------------------
# 2. Weekly dataset
# ---------------------------------------------------------------------------
 
df_week <- df_case |>
  mutate(
    week = floor_date(.data[[col_admission]], "week", week_start = week_start),
 
    age_group = case_when(
      age <  1              ~ "<1",
      age >= 1  & age <  5  ~ "1-5",
      age >= 5  & age <  10 ~ "5-10",
      age >= 10 & age <= 15 ~ "10-15",
      age >  15             ~ ">15"
    ),
    age_group = factor(age_group, levels = names(age_cols)),
 
    birth_year = if (is.na(col_dob)) {
      year(.data[[col_admission]]) - floor(age)
    } else {
      year(.data[[col_dob]])
    },
 
    cohort = case_when(
      birth_year <  2009              ~ "<2009",
      between(birth_year, 2009, 2013) ~ "2009-2013",
      between(birth_year, 2014, 2018) ~ "2014-2018",
      between(birth_year, 2019, 2023) ~ "2019-2023",
      birth_year == 2024              ~ "2024",
      birth_year == 2025              ~ "2025"
    ),
    cohort = factor(cohort, levels = names(cohort_cols))
  ) |>
  drop_na(week)
 
n_unclassified <- sum(!is.na(df_week$birth_year) & is.na(df_week$cohort))
if (n_unclassified > 0) {
  warning(n_unclassified, " cases have a birth year outside the defined cohorts")
}
 
# ---------------------------------------------------------------------------
# 3. Detect outbreak periods
# ---------------------------------------------------------------------------
 
# Weeks with no admissions must appear as zero, or a stretch of quiet weeks
# reads as continuous transmission
weekly_counts <- function(data, from = NULL, to = NULL, week_col = week) {
  obs <- data |>
    dplyr::count({{ week_col }}, name = "cases") |>
    dplyr::rename(week = {{ week_col }})
 
  from <- if (is.null(from)) min(obs$week) else as.Date(from)
  to   <- if (is.null(to))   max(obs$week) else as.Date(to)
 
  tibble(week = seq(from, to, by = "week")) |>
    left_join(obs, by = "week") |>
    mutate(cases = replace_na(cases, 0))
}
 
# One pass over the whole record. A period starts at the first of `min_run`
# consecutive weeks at or above `threshold`; two such runs separated by fewer
# than `max_gap` weeks below threshold are treated as one period.
find_outbreaks <- function(data, threshold = ob_threshold,
                           min_run = ob_min_run, max_gap = ob_max_gap,
                           week_col = week) {
  wc <- weekly_counts(data, week_col = {{ week_col }})
  r  <- rle(wc$cases >= threshold)
 
  end_i   <- cumsum(r$lengths)
  start_i <- end_i - r$lengths + 1
  seed    <- which(r$values & r$lengths >= min_run)
 
  if (length(seed) == 0) {
    return(tibble(episode = integer(), start = as.Date(character()),
                  end = as.Date(character()), weeks = integer(),
                  cases = integer(), peak_week = as.Date(character()),
                  peak_cases = integer()))
  }
 
  s <- start_i[seed]
  e <- end_i[seed]
 
  merged_s <- s[1]
  merged_e <- e[1]
  out_s <- integer()
  out_e <- integer()
 
  for (i in seq_along(s)[-1]) {
    if (s[i] - merged_e - 1 < max_gap) {
      merged_e <- e[i]
    } else {
      out_s <- c(out_s, merged_s)
      out_e <- c(out_e, merged_e)
      merged_s <- s[i]
      merged_e <- e[i]
    }
  }
  out_s <- c(out_s, merged_s)
  out_e <- c(out_e, merged_e)
 
  tibble(
    episode = seq_along(out_s),
    i_start = out_s,
    i_end   = out_e
  ) |>
    mutate(
      start = wc$week[i_start],
      end   = wc$week[i_end],
      weeks = i_end - i_start + 1L,
      cases = map2_int(i_start, i_end, \(a, b) sum(wc$cases[a:b])),
      peak_week  = map2_vec(i_start, i_end,
                            \(a, b) wc$week[a:b][which.max(wc$cases[a:b])]),
      peak_cases = map2_int(i_start, i_end, \(a, b) max(wc$cases[a:b]))
    ) |>
    select(-i_start, -i_end)
}
 
# Sensitivity of the periods to the threshold. Report this alongside the rule.
outbreak_sensitivity <- function(data, thresholds = c(3, 5, 10, 20, 30)) {
  map_dfr(thresholds, \(x) find_outbreaks(data, threshold = x) |>
            mutate(threshold = x, .before = 1))
}
 
cat("\n--- All episodes at threshold", ob_threshold, "---\n")
episodes_all <- find_outbreaks(df_week)
print(episodes_all)
 
cat("\n--- Sensitivity to the threshold ---\n")
print(outbreak_sensitivity(df_week), n = Inf)
 
episodes <- episodes_all |>
  filter(cases >= ob_min_cases) |>
  arrange(start) |>
  mutate(episode = row_number())
 
cat("\n--- Episodes retained (at least", ob_min_cases, "cases) ---\n")
print(episodes)
 
if (nrow(episodes) != 2) {
  stop("Expected 2 outbreak periods, found ", nrow(episodes),
       ". Adjust ob_threshold, ob_max_gap or ob_min_cases, or extend the ",
       "figure to facet over all episodes.")
}
 
# ---------------------------------------------------------------------------
# 4. Plot window derived from the detected periods
# ---------------------------------------------------------------------------
 
ep1 <- episodes |> slice(1)
ep2 <- episodes |> slice(2)
 
plot_start <- ep1$start
gap_start  <- ep1$end + 7      # week < gap_start keeps ep1's final week
gap_end    <- ep2$start
plot_end   <- ep2$end
 
df_plot <- df_week |>
  filter(
    between(week, plot_start, plot_end),
    week < gap_start | week >= gap_end
  ) |>
  mutate(
    period = factor(if_else(week < gap_start, "p1", "p2"),
                    levels = c("p1", "p2"))
  )
 
# ---------------------------------------------------------------------------
# 5. What the figure leaves out
# ---------------------------------------------------------------------------
 
exclusions <- df_week |>
  summarise(
    total           = n(),
    before_period1  = sum(week < plot_start),
    between_periods = sum(week >= gap_start & week < gap_end),
    after_period2   = sum(week > plot_end),
    plotted         = sum(between(week, plot_start, plot_end) &
                            (week < gap_start | week >= gap_end)),
    missing_age     = sum(is.na(age_group)),
    missing_cohort  = sum(is.na(cohort))
  )
print(exclusions)
 
cat(sprintf(
  paste0("Periods: %s to %s and %s to %s.\n",
         "Plotted %d of %d admissions (%.1f%%). Outside the periods: ",
         "%d before, %d between, %d after.\n"),
  format(plot_start, "%d %b %Y"), format(gap_start - 7, "%d %b %Y"),
  format(gap_end, "%d %b %Y"),    format(plot_end, "%d %b %Y"),
  exclusions$plotted, exclusions$total,
  100 * exclusions$plotted / exclusions$total,
  exclusions$before_period1, exclusions$between_periods,
  exclusions$after_period2
))
 
# ---------------------------------------------------------------------------
# 6. Shared layers, scales and theme
# ---------------------------------------------------------------------------
 
# Axis limits per panel. The 4-day margins contain the half-width of the
# outermost bars, which are 7 days wide and centred on the week start.
period_range <- tibble(
  period = factor(c("p1", "p2"), levels = c("p1", "p2")),
  start  = c(plot_start, gap_end),          # true period boundaries
  end    = c(gap_start - 7, plot_end),
  lo     = c(plot_start - 4, gap_end - 4),  # padded axis limits
  hi     = c(gap_start - 1, plot_end + 4)
)
 
anchors_f <- period_range |>
  pivot_longer(c(lo, hi), values_to = "week") |>
  select(period, week)
 
# Shading clipped to its panel. Unclipped, a rectangle extending past the
# panel limits would widen the scale and break the proportional panel widths.
shade_f <- shade_win |>
  mutate(period = factor(if_else(xmin < gap_start, "p1", "p2"),
                         levels = c("p1", "p2")),
         label = shade_label) |>
  left_join(period_range, by = "period") |>
  mutate(xmin = pmax(xmin, lo), xmax = pmin(xmax, hi)) |>
  filter(xmax > xmin)
 
# Only boundaries that fall inside their panel get a dashed line
vline_f <- shade_f |>
  pivot_longer(c(xmin, xmax), values_to = "x") |>
  filter(between(x, lo, hi))
 
anchor_layer <- geom_blank(data = anchors_f, aes(x = week), inherit.aes = FALSE)
 
draw_key_shade <- function(data, params, size) {
  fill_col <- if (is.null(data$fill)) shade_fill else data$fill
  alp <- if (is.null(data$alpha) || is.na(data$alpha)) 1 else data$alpha
  grid::grobTree(
    grid::rectGrob(gp = grid::gpar(fill = alpha(fill_col, alp), col = NA)),
    grid::segmentsGrob(
      x0 = c(0.05, 0.95), x1 = c(0.05, 0.95),
      y0 = c(0, 0),       y1 = c(1, 1),
      gp = grid::gpar(col = "black", lty = "dashed", lwd = 1)
    )
  )
}
 
# Only panel A maps fill, so only panel A produces a key; panels B and C
# spend their fill scale on the stacked bars.
rect_keyed <- geom_rect(
  data = shade_f,
  aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = label),
  inherit.aes = FALSE, alpha = shade_alpha, key_glyph = draw_key_shade
)
 
rect_plain <- geom_rect(
  data = shade_f,
  aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
  inherit.aes = FALSE, fill = shade_fill, alpha = shade_alpha
)
 
layers_bg_a  <- list(rect_keyed, anchor_layer)
layers_bg_bc <- list(rect_plain, anchor_layer)
 
layers_fg <- list(
  geom_vline(data = vline_f, aes(xintercept = x), inherit.aes = FALSE,
             linetype = "dashed", colour = "black", linewidth = 0.5),
  # space = "free_x" is only available in facet_grid, not facet_wrap
  facet_grid(cols = vars(period), scales = "free_x", space = "free_x")
)
 
date_lab <- function(x) {
  yr  <- format(x, "%Y")
  out <- if_else(!duplicated(yr) & !is.na(yr),
                 format(x, "%b\n%Y"), format(x, "%b"))
  replace_na(out, "")
}
 
axis_required <- sort(unique(c(
  period_range$start, period_range$end,
  shade_win$xmin, shade_win$xmax
)))

x_scale <- scale_x_date(
  breaks = \(limits) {
    b <- axis_required[axis_required >= as.Date(limits[1]) &
                         axis_required <= as.Date(limits[2])]
    sort(unique(b))
  },
  minor_breaks = NULL,
  labels = \(x) format(round_date(x, "month"), "%b\n%Y"),
  expand = expansion(mult = c(0.01, 0.01))
)
 
base_theme <- theme_minimal(base_size = 12) +
  theme(
    panel.spacing.x = unit(1.5, "lines"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.y = element_line(colour = "grey90", linewidth = 0.4),
    axis.line.x  = element_line(colour = "black", linewidth = 0.6),
    axis.ticks.x = element_line(colour = "black"),
    axis.text    = element_text(colour = "black"),
    axis.title   = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    strip.background = element_blank(),
    strip.text       = element_blank(),
    plot.tag = element_text(face = "bold", size = 13),
    plot.tag.position = "topleft",
    plot.margin = margin(t = 5, r = 15, b = 5, l = 10)
  )
 
drop_x <- theme(
  axis.title.x = element_blank(),
  axis.text.x  = element_blank(),
  axis.ticks.x = element_blank()
)
 
# ---------------------------------------------------------------------------
# 7. Panel data and panel builders
# ---------------------------------------------------------------------------
 
prep_stack <- function(d, var) {
  d |>
    drop_na({{ var }}) |>
    count(period, week, grp = {{ var }}) |>
    mutate(prop = n / sum(n), .by = week)
}
 
d_count  <- df_plot |> count(period, week)
d_age    <- prep_stack(df_plot, age_group)
d_cohort <- prep_stack(df_plot, cohort)
 
stacked_panel <- function(d, cols, legend_title, reverse_stack, tag) {
  ggplot(d, aes(x = week, y = prop)) +
    layers_bg_bc +
    geom_col(aes(fill = grp), width = 7, colour = "white", linewidth = 0.2,
             position = position_stack(reverse = reverse_stack)) +
    layers_fg +
    scale_fill_manual(values = cols) +
    scale_y_continuous(labels = label_percent(), limits = c(0, 1),
                       breaks = seq(0, 1, by = 0.25), expand = c(0, 0)) +
    x_scale +
    labs(y = "Proportion", fill = legend_title, tag = tag) +
    base_theme
}
 
pA <- ggplot(d_count, aes(x = week, y = n)) +
  layers_bg_a +
  geom_col(width = 7, fill = "grey35") +
  layers_fg +
  scale_fill_manual(
    values = setNames(shade_fill, shade_label),
    name = NULL,
    guide = guide_legend(keywidth = unit(20, "pt"), keyheight = unit(24, "pt"))
  ) +
  scale_y_continuous(labels = label_comma(),
                     expand = expansion(mult = c(0, 0.05))) +
  x_scale +
  labs(y = "Case counts", tag = "A") +
  base_theme + drop_x
 
pB <- stacked_panel(d_age, age_cols, "Age group",
                    reverse_stack = TRUE, tag = "B") +
  guides(fill = guide_legend(reverse = TRUE)) +
  drop_x
 
pC <- stacked_panel(d_cohort, cohort_cols, "Birth cohort",
                    reverse_stack = FALSE, tag = "C") +
  labs(x = "Admission week")
 
# ---------------------------------------------------------------------------
# 8. Assemble and save
# ---------------------------------------------------------------------------
 
fig <- pA / pB / pC +
  plot_layout(heights = c(0.7, 1, 1), guides = "collect") &
  theme(
    legend.position      = "right",
    legend.justification = "top",
    legend.box.just      = "left",
    legend.spacing.y     = unit(12, "pt")
  )
 
# print(fig)
#  
# ggsave("figs/chap1/case_age_cohort.png", fig,
#        width = 8, height = 5, dpi = 300, bg = "white")
```

### 2024-2025

```{r}
df_ep2 <- df_week |>
  filter(between(week, ep2$start, ep2$end))
```

```{r}
# ---------------------------------------------------------------------------
# 1. Parameters
# ---------------------------------------------------------------------------

col_date     <- "ngay_nv"   # df_ep2 carries admission date, not onset
col_district <- "district"        # rename if the district column differs
date_axis_label <- "Date of admission"   # must match col_date

# Vaccination campaign start dates (date_vax_cp and date_vax_cp2 previously)
campaign_dates <- ymd(c("2024-09-01", "2024-10-01"))

ma_days <- 7               # moving average window, applied to both C and B

# Districts merged to match df_map_hcmc, which already has Thu Duc City
thuduc_parts <- c("2", "9", "thu duc")

age_levels <- c("<1", "1-5", "5-10", "10-15", ">15")

# ---------------------------------------------------------------------------
# 2. District name normalisation
# ---------------------------------------------------------------------------

merge_thuduc <- function(x) if_else(x %in% thuduc_parts, "thu duc", x)

census_dist <- df_census_district |>
  mutate(district = merge_thuduc(district)) |>
  summarise(n_census = sum(n_census), .by = district)

map_dist <- df_map_hcmc |>
  mutate(district = merge_thuduc(district_blank)) |>
  select(district)

cat("In census but not map:", setdiff(census_dist$district, map_dist$district), "\n")
cat("In map but not census:", setdiff(map_dist$district, census_dist$district), "\n")

# ---------------------------------------------------------------------------
# 3. Daily district counts
# ---------------------------------------------------------------------------

df_dist <- df_ep2 |>
  mutate(
    district = merge_thuduc(.data[[col_district]]),
    date = .data[[col_date]]
  ) |>
  drop_na(district, date)

cat("Cases with no district match:",
    sum(!df_dist$district %in% census_dist$district), "\n")

date_grid <- seq(min(df_dist$date), max(df_dist$date), by = "day")

# The ORI window falling inside this episode, clipped to the data range
ori_win <- shade_win |>
  filter(xmax >= min(date_grid), xmin <= max(date_grid)) |>
  mutate(xmin = pmax(xmin, min(date_grid)),
         xmax = pmin(xmax, max(date_grid)))

shade_label <- "Outbreak response immunisation"

draw_key_shade <- function(data, params, size) {
  fill_col <- if (is.null(data$fill)) shade_fill else data$fill
  alp <- if (is.null(data$alpha) || is.na(data$alpha)) 1 else data$alpha
  grid::grobTree(
    grid::rectGrob(gp = grid::gpar(fill = scales::alpha(fill_col, alp),
                                   col = NA)),
    grid::segmentsGrob(
      x0 = c(0.05, 0.95), x1 = c(0.05, 0.95),
      y0 = c(0, 0),       y1 = c(1, 1),
      gp = grid::gpar(col = "black", lty = "dashed", lwd = 1)
    )
  )
}

# Same shading as shade_layer, but with fill mapped so it produces a key
shade_layer_keyed <- geom_rect(
  data = mutate(ori_win, label = shade_label),
  aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = label),
  inherit.aes = FALSE, alpha = shade_alpha, key_glyph = draw_key_shade
)

# Drawn before the lines so the shading sits behind them
shade_layer <- geom_rect(
  data = ori_win,
  aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
  inherit.aes = FALSE, fill = shade_fill, alpha = shade_alpha
)

vline_layer <- geom_vline(
  xintercept = c(ori_win$xmin, ori_win$xmax),
  linetype = "dashed", colour = "black", linewidth = 0.5
)

# Same four ticks as the age and cohort figure: episode bounds and ORI bounds
axis_required_ep2 <- sort(unique(c(range(date_grid), ori_win$xmin, ori_win$xmax)))

x_scale_ep2 <- scale_x_date(
  breaks = \(limits) {
    b <- axis_required_ep2[axis_required_ep2 >= as.Date(limits[1]) &
                             axis_required_ep2 <= as.Date(limits[2])]
    sort(unique(b))
  },
  minor_breaks = NULL,
  labels = \(x) format(round_date(x, "month"), "%b\n%Y"),
  expand = expansion(mult = c(0.02, 0.05))
)

daily_dist <- df_dist |>
  count(district, date) |>
  # Days with no cases anywhere are absent from count(); without completing
  # them the cumulative sums and moving averages skip those days entirely
  complete(district = census_dist$district, date = date_grid,
           fill = list(n = 0))
```

```{r}
# ---------------------------------------------------------------------------
# 4. Panel A: cumulative incidence by district, quarterly snapshots
# ---------------------------------------------------------------------------

snapshot_unit <- "quarter"

snapshot_dist <- daily_dist |>
  mutate(bin = floor_date(date, snapshot_unit)) |>
  summarise(n = sum(n), .by = c(district, bin)) |>
  arrange(district, bin) |>
  mutate(cum_n = cumsum(n), .by = district) |>
  left_join(census_dist, by = "district") |>
  mutate(
    cum_inc = 1e5 * cum_n / n_census,
    cum_inc = if_else(cum_inc == 0, NA, cum_inc),
    bin_end = pmin(ceiling_date(bin, snapshot_unit) - 1, max(date_grid)),
    yr = factor(year(bin_end)),
    # Levels from month.abb, not from order of appearance: "Jun" occurs in
    # both years, so reordering by first occurrence would put Mar after Dec
    # inside 2025
    mo = factor(format(bin_end, "%b"), levels = month.abb)
  )

map_snapshot <- map_dist |>
  left_join(snapshot_dist, by = "district")

pA_maps <- ggplot(map_snapshot) +
  geom_sf(aes(fill = cum_inc),
          colour = "grey40",
          linewidth = 0.15) +
  scale_fill_gsea(
    rescaler = function(x, to = c(0, 1), from = NULL)
      scales::rescale(x^0.5),
    na.value = "white",
    guide = guide_colourbar(
      title.position = "top",
      title.hjust = 0.5,
      barwidth  = unit(55, "mm"),
      barheight = unit(3, "mm"),
      frame.colour = "black",
      frame.linewidth = 0.3,
      ticks.colour = "black"
    )
  ) +
  # Outer strip spans the panels sharing a year, inner strip gives the month
  facet_nested_wrap(
    vars(yr, mo),
    nrow = 1,
    nest_line = element_line(colour = "grey30", linewidth = 0.4),
    strip = strip_nested(
      background_x = list(
        element_blank(),
        element_rect(
          fill = "grey93",
          colour = "grey40",
          linewidth = 0.3
        )
      ),
      text_x = list(
        element_text(
          hjust = 0,
          face = "bold",
          colour = "black",
          margin = margin(t = 2, b = 3)
        ),
        element_text(
          hjust = 0.5,
          colour = "black",
          margin = margin(t = 2, b = 3)
        )
      ), 
      by_layer_x = TRUE
    )
  ) +
  labs(fill = "Cumulative incidence rate (per 100,000 population)", tag = "A") +
  theme_bw(base_size = 11) +
  theme(
    legend.position = "bottom",
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.ticks.length = unit(0, "pt"),
    panel.grid = element_blank(),
    panel.border = element_rect(
      colour = "grey40",
      linewidth = 0.3,
      fill = NA
    ),
    panel.spacing.x = unit(3, "pt"),
    ggh4x.facet.nestline = element_line(colour = "grey25", linewidth = 0.4),
    plot.tag = element_text(face = "bold", size = 14)
  )

# Same coordinate limits as the map facets, geometry drawn invisibly, so the
# scale bar is computed against a matching coordinate range
p_arrow <- ggplot() +
  annotation_north_arrow(
    location = "tr", which_north = "grid",
    height = unit(7, "mm"), width = unit(6, "mm"),
    style = north_arrow_fancy_orienteering(text_size = 6)
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
  theme_void() +
  theme(plot.margin = margin(4, 4, 4, 4))

pA <- pA_maps +
  theme(
    legend.position = "bottom",
    legend.justification = "left",
    legend.margin = margin(t = 4, b = 0),
    legend.title = element_text(hjust = 0),
    plot.margin = margin(2, 2, 2, 2)
  ) + patchwork::inset_element(
  p_arrow,
  left = 0.88, bottom = 0.06, right = 0.99, top = 0.30,
  align_to = "full", clip = FALSE
)

# pA
```

```{r}
# ---------------------------------------------------------------------------
# 5. Attack rate helper
#
# rate = cases on day t / (population - cumulative cases up to day t-1).
# This is a daily hazard among those not yet recorded as cases, not a
# cumulative attack rate; name it accordingly in the caption.
# ---------------------------------------------------------------------------

add_attack_rate <- function(d, group, n_days = ma_days) {
  d |>
    arrange({{ group }}, date) |>
    mutate(
      cum_cases = cumsum(n),
      lag_cum   = lag(cum_cases, default = 0),
      at_risk   = n_census - lag_cum,
      rate      = n / at_risk,
      rate_ma   = slide_dbl(rate, mean, .before = n_days - 1, .complete = TRUE),
      .by = {{ group }}
    )
}

# ---------------------------------------------------------------------------
# 6. Panel B: district attack rate
# ---------------------------------------------------------------------------

region_lookup <- tribble(
  ~district,     ~region,
  "binh chanh",  "West",
  "binh tan",    "West",
  "hoc mon",     "North",
  "12",          "North",
  "cu chi",      "North",
  "thu duc",     "East",
  "nha be",      "South",
  "7",           "South",
  "can gio",     "South"
)

rate_dist <- daily_dist |>
  left_join(census_dist, by = "district") |>
  add_attack_rate(district) |>
  left_join(region_lookup, by = "district") |>
  mutate(
    region = replace_na(region, "Central"),
    region = factor(region, levels = c("West", "North", "East", "South", "Central")),
    highlight = if_else(region == "West", "West", "Others")
  )

# log() of a zero moving average returns -Inf, which ggplot drops with a
# warning. Count them so the gaps in the lines are expected rather than a
# surprise.
cat("District days with a zero moving average:",
    sum(rate_dist$rate_ma == 0, na.rm = TRUE), "\n")

district_pals <- c("South" = "#332288", "North" = "#88CCEE", "Central" = "#117733", "East" = "#DDCC77", "West" = "#CC6677")

pB <- rate_dist |>
  ggplot(aes(x = date, y = log(rate_ma * 1e5),
             group = district, colour = region, alpha = highlight)) +
  shade_layer +
  geom_line() +
  vline_layer +
  scale_colour_manual(values = district_pals) +
  scale_alpha_discrete(limits = c("West", "Others"), range = c(1, 0.3),
                       guide = "none") +
  x_scale_ep2 +
  labs(x = NULL, y = "Log attack rate", colour = "Region", tag = "B") +
  theme_classic() +
  theme(legend.position = "bottom",
        plot.tag = element_text(face = "bold", size = 14))

pB
```

```{r}
# ---------------------------------------------------------------------------
# 7. Panel C: age group attack rate
# ---------------------------------------------------------------------------

# Built with the same case_when as df_week, so the denominator bands match
# the numerator bands exactly. A parallel cut() with right = TRUE would shift
# every boundary by one year.
census_age <- df_census |>
  mutate(
    # census codes age 1 as under 1, so census age n is completed age n - 1
    age_years = age - 1,
    age_group = case_when(
      age_years <  1                     ~ "<1",
      age_years >= 1  & age_years <  5   ~ "1-5",
      age_years >= 5  & age_years <  10  ~ "5-10",
      age_years >= 10 & age_years <= 15  ~ "10-15",
      age_years >  15                    ~ ">15"
    ),
    age_group = factor(age_group, levels = age_levels)
  ) |>
  summarise(n_census = sum(n), .by = age_group)

print(census_age)

daily_age <- df_ep2 |>
  mutate(date = .data[[col_date]]) |>
  drop_na(age_group, date) |>
  count(age_group, date) |>
  complete(age_group, date = date_grid, fill = list(n = 0))

rate_age <- daily_age |>
  left_join(census_age, by = "age_group") |>
  add_attack_rate(age_group)

pC <- rate_age |>
  mutate(age_group = fct_rev(age_group)) |>
  ggplot(aes(x = date, y = log(rate_ma * 1e5), colour = age_group)) +
  shade_layer_keyed +
  geom_line() +
  vline_layer +
  scale_color_observable() +
  scale_fill_manual(
    values = setNames(shade_fill, shade_label),
    name = NULL,
    guide = guide_legend(order = 2, keywidth = unit(16, "pt"),
                         keyheight = unit(16, "pt"))
  ) +
  x_scale_ep2 +
  guides(colour = guide_legend(reverse = TRUE, order = 1)) +
  labs(x = date_axis_label, y = "Log attack rate", colour = "Age (years)",
       tag = "C") +
  theme_classic() +
  theme(legend.position = "bottom",
        legend.box = "vertical",
        plot.tag = element_text(face = "bold", size = 14))

# pC
```

```{r}
# ---------------------------------------------------------------------------
# 8. Assemble
# ---------------------------------------------------------------------------

fig_ep2 <- plot_grid(pA, pB, pC, ncol = 1, rel_heights = c(1, 1, 1.3))

fig_ep2

# ggsave("figs/chap1/ep2_spatial_attack.png", fig_ep2,
#        width = 5, height = 7.5, dpi = 500, bg = "white")
```

### Values cases

```{r}
# ===========================================================================
# SUMMARY VALUES FOR THE EPIDEMIC CURVE FIGURE
# ===========================================================================
 
pc <- \(x) sprintf("%.1f", 100 * x)
 
# ---- 0. Duplicates --------------------------------------------------------
# Case totals are quoted in the text, so settle this first. Keep dedupe FALSE
# if the figure was built from the un-deduplicated frame, so text and figure
# agree; if TRUE, the figure must be rebuilt too.
 
df_week |> summarise(rows = n(), unique_ids = n_distinct(id))
df_week |> count(id) |> filter(n > 1) |> nrow()
 
dedupe <- FALSE
cases  <- if (dedupe) distinct(df_week, id, .keep_all = TRUE) else df_week
 
# ---- 1. Outbreak periods, totals, and peaks -------------------------------
 
ep_win <- tibble(
  episode = c("2018-2020", "2024-2025"),
  start   = c(ep1$start, ep2$start),
  end     = c(ep1$end,   ep2$end)
)
 
ep_win |>
  mutate(weeks  = as.numeric(end - start) / 7,
         months = as.numeric(end - start) / 30.44) |>
  print()
 
ep_totals <- ep_win |>
  reframe(cases |> filter(between(week, start, end)) |> summarise(n = n()),
          .by = episode)
 
ep_totals |> print()
 
# weekly peak: date and size
cases |>
  inner_join(ep_win, by = join_by(between(week, start, end))) |>
  count(episode, week) |>
  slice_max(n, n = 1, by = episode) |>
  print()
 
# ---- 2. Cases between the two outbreaks -----------------------------------
 
gap_start <- ep1$end
gap_end   <- ep2$start
 
cases |>
  filter(week > gap_start, week < gap_end) |>
  summarise(n = n(), first = min(week), last = max(week),
            n_weeks_with_cases = n_distinct(week))
 
# ---- 3. Age composition, monthly, with denominators -----------------------
# <5 = "<1" + "1-5"; <10 adds "5-10".
 
age_mo <- cases |>
  inner_join(ep_win, by = join_by(between(week, start, end))) |>
  drop_na(age_group) |>
  mutate(month = floor_date(week, "month")) |>
  summarise(n_tot   = n(),
            n_u5    = sum(age_group %in% c("<1", "1-5")),
            n_u10   = sum(age_group %in% c("<1", "1-5", "5-10")),
            n_u1    = sum(age_group == "<1"),
            .by = c(episode, month)) |>
  mutate(pct_u1 = n_u1 / n_tot, pct_u5 = n_u5 / n_tot, pct_u10 = n_u10 / n_tot) |>
  arrange(episode, month) |> 
  as_tibble()
 
age_mo |> print(n = Inf)
 
# the specific window in your draft: May to July 2024
age_mo |>
  filter(month >= ymd("2024-05-01"), month <= ymd("2024-07-01")) |>
  summarise(n_tot = sum(n_tot),
            pct_u5  = sum(n_u5) / sum(n_tot),
            pct_u10 = sum(n_u10) / sum(n_tot))
 
# equivalent window at the start of the first outbreak, for the parallel claim
age_mo |>
  filter(episode == "2018-2020") |>
  slice_head(n = 3) |>
  summarise(n_tot = sum(n_tot),
            pct_u5  = sum(n_u5) / sum(n_tot),
            pct_u10 = sum(n_u10) / sum(n_tot))
 
# full age breakdown for the first three months of each outbreak
cases |>
  inner_join(ep_win, by = join_by(between(week, start, end))) |>
  mutate(month = floor_date(week, "month")) |>
  filter(month <= min(month) %m+% months(2), .by = episode) |>
  count(episode, age_group) |>
  mutate(pct = n / sum(n), .by = episode) |>
  as_tibble() |> 
  print(n = Inf)
 
# ---- 4. Cohort composition, monthly, with denominators -------------------
 
coh_mo <- cases |>
  inner_join(ep_win, by = join_by(between(week, start, end))) |>
  drop_na(cohort) |>
  mutate(month = floor_date(week, "month")) |>
  count(episode, month, cohort) |>
  mutate(n_tot = sum(n), pct = n / sum(n), .by = c(episode, month)) |>
  arrange(episode, month, desc(pct))
 
coh_mo |>
  as_tibble() |> 
  mutate(lab = sprintf("%s%% (%d/%d)", pc(pct), n, n_tot)) |>
  select(episode, month, cohort, lab) |>
  print(n = Inf)
 
# leading cohort each month, which is the trajectory your text describes
coh_mo |>
  slice_max(pct, n = 1, by = c(episode, month)) |>
  mutate(lab = sprintf("%s: %s%% (%d/%d)", cohort, pc(pct), n, n_tot)) |>
  select(episode, month, n_tot, lab) |>
  print(n = Inf)
 
# ---- 5. ORI windows and the post-ORI cohort rise -------------------------
 
shade_win |>
  mutate(days = as.numeric(xmax - xmin),
         months = days / 30.44) |>
  print()
 
# monthly share of one named cohort, with its denominator
coh_track <- function(coh, from, to) {
  coh_mo |>
    filter(cohort == coh, month >= ymd(from), month <= ymd(to)) |>
    mutate(lab = sprintf("%s: %s%% (%d/%d)",
                         format(month, "%b %Y"), pc(pct), n, n_tot)) |>
    select(month, n, n_tot, pct, lab)
}
 
coh_track("2019-2023", "2019-06-01", "2020-03-31") |> print(n = Inf)
coh_track("2019-2023", "2024-05-01", "2024-08-31") |> print(n = Inf)
coh_track("2024",      "2024-09-01", "2025-12-31") |> print(n = Inf)
coh_track("2025",      "2025-01-01", "2025-12-31") |> print(n = Inf)
 
# CAUTION: a rising share can occur while counts fall. Check both.
coh_mo |>
  filter(cohort == "2019-2023", month >= ymd("2019-06-01")) |>
  select(month, n, n_tot, pct) |>
  print(n = Inf)
 
# ---- 6. Sentence-ready strings -------------------------------------------
 
ep_totals |>
  left_join(ep_win, by = "episode") |>
  mutate(txt = sprintf("%s outbreak: %s to %s, %d cases",
                       episode, format(start, "%B %Y"), format(end, "%B %Y"), n)) |>
  pull(txt) |>
  writeLines()
 
age_mo |>
  filter(month %in% ymd(c("2018-09-01", "2024-05-01", "2024-06-01", "2024-07-01"))) |>
  mutate(txt = sprintf("%s: under 5 %s%%, under 10 %s%% (n = %d)",
                       format(month, "%b %Y"), pc(pct_u5), pc(pct_u10), n_tot)) |>
  pull(txt) |>
  writeLines()
```

### Values cases 2

```{r}
# ===========================================================================
# SUMMARY VALUES FOR THE 2024-2025 OUTBREAK FIGURE
# ===========================================================================
# Panel B and C plot log(rate_ma * 1e5), a daily hazard among those not yet
# recorded as cases. Everything below back-transforms to a rate per 100,000
# per day, which is the quotable form.
 
r100k <- \(x) x * 1e5
fm    <- \(x) sprintf("%.2f", x)
 
ori_start <- ori_win$xmin[1]
ori_end   <- ori_win$xmax[1]
 
c(as.character(ori_start), as.character(ori_end),
  as.character(as.numeric(ori_end - ori_start)))
 
# ---- 1. Panel A: cumulative incidence by district and snapshot -----------
 
snap_tab <- snapshot_dist |>
  as_tibble() |>
  select(district, bin, bin_end, cum_n, n_census, cum_inc)
 
# city-wide cumulative incidence at each snapshot
snap_tab |>
  summarise(cum_n = sum(cum_n), n_census = sum(n_census), .by = bin_end) |>
  mutate(cum_inc = 1e5 * cum_n / n_census) |>
  arrange(bin_end) |>
  print(n = Inf)
 
# highest and lowest district at each snapshot, and the ratio between them
snap_tab |>
  filter(!is.na(cum_inc)) |>
  summarise(
    hi_d = district[which.max(cum_inc)], hi = max(cum_inc),
    lo_d = district[which.min(cum_inc)], lo = min(cum_inc),
    ratio = max(cum_inc) / min(cum_inc),
    n_dist = n(),
    .by = bin_end
  ) |>
  arrange(bin_end) |>
  print(n = Inf)
 
# final snapshot, every district ranked
snap_tab |>
  filter(bin_end == max(bin_end)) |>
  arrange(desc(cum_inc)) |>
  print(n = Inf)
 
# which districts had recorded cases by the first snapshot
snap_tab |>
  filter(bin_end == min(bin_end), cum_n > 0) |>
  arrange(desc(cum_inc)) |>
  print(n = Inf)
 
# ---- 2. Panel B: district attack rate ------------------------------------
 
rate_dist_t <- rate_dist |> as_tibble()
 
# peak daily hazard per district, and when it occurred
rate_dist_t |>
  filter(!is.na(rate_ma), rate_ma > 0) |>
  slice_max(rate_ma, n = 1, by = district) |>
  transmute(region, district, date, peak_per100k = r100k(rate_ma)) |>
  arrange(desc(peak_per100k)) |>
  print(n = Inf)
 
# by region: when regions reached their peak, to check for a spatial wave
rate_dist_t |>
  filter(!is.na(rate_ma), rate_ma > 0) |>
  slice_max(rate_ma, n = 1, by = district) |>
  summarise(n_dist = n(),
            first_peak = min(date), last_peak = max(date),
            med_peak = median(date),
            .by = region) |>
  arrange(med_peak) |>
  print(n = Inf)
 
# were western districts earlier? date each district first exceeded a
# reference hazard, here the 1 per 100,000 per day level
thr <- 1 / 1e5
 
rate_dist_t |>
  filter(!is.na(rate_ma), rate_ma >= thr) |>
  slice_min(date, n = 1, by = district) |>
  transmute(region, district, first_crossing = date) |>
  arrange(first_crossing) |>
  print(n = Inf)
 
# rate before, during, and after the ORI window, pooled by region
rate_dist_t |>
  mutate(phase = case_when(date <  ori_start ~ "Before ORI",
                           date <= ori_end   ~ "During ORI",
                           .default          = "After ORI")) |>
  summarise(cases = sum(n), n_census = first(n_census),
            .by = c(region, district, phase)) |>
  summarise(cases = sum(cases), n_census = sum(n_census),
            .by = c(region, phase)) |>
  mutate(inc_per100k = 1e5 * cases / n_census) |>
  pivot_wider(names_from = phase, values_from = c(cases, inc_per100k)) |>
  print(n = Inf)
 
# ---- 3. Panel C: age group attack rate -----------------------------------
 
rate_age_t <- rate_age |> as_tibble()
 
# peak daily hazard per age group, and when
rate_age_t |>
  filter(!is.na(rate_ma), rate_ma > 0) |>
  slice_max(rate_ma, n = 1, by = age_group) |>
  transmute(age_group, date, peak_per100k = r100k(rate_ma), n_census) |>
  arrange(desc(peak_per100k)) |>
  print(n = Inf)
 
# ratio of the <1 hazard to every other group, at each group's own peak
rate_age_t |>
  filter(!is.na(rate_ma), rate_ma > 0) |>
  slice_max(rate_ma, n = 1, by = age_group) |>
  mutate(ratio_to_u1 = max(rate_ma) / rate_ma) |>
  transmute(age_group, peak_per100k = r100k(rate_ma), ratio_to_u1) |>
  arrange(desc(peak_per100k)) |>
  print(n = Inf)
 
# same-day comparison at the city-wide peak, which is the cleaner statement
peak_day <- rate_age_t |>
  summarise(tot = sum(n), .by = date) |>
  slice_max(tot, n = 1) |>
  pull(date)
 
peak_day
 
rate_age_t |>
  filter(date == peak_day) |>
  transmute(age_group, rate_per100k = r100k(rate_ma),
            ratio_to_u1 = max(rate_ma, na.rm = TRUE) / rate_ma) |>
  arrange(desc(rate_per100k)) |>
  print(n = Inf)
 
# cumulative incidence by age over the whole episode, and by ORI phase
rate_age_t |>
  mutate(phase = case_when(date <  ori_start ~ "Before ORI",
                           date <= ori_end   ~ "During ORI",
                           .default          = "After ORI")) |>
  summarise(cases = sum(n), n_census = first(n_census),
            .by = c(age_group, phase)) |>
  mutate(inc_per100k = 1e5 * cases / n_census) |>
  pivot_wider(names_from = phase, values_from = c(cases, inc_per100k)) |>
  print(n = Inf)
 
rate_age_t |>
  summarise(cases = sum(n), n_census = first(n_census), .by = age_group) |>
  mutate(inc_per100k = 1e5 * cases / n_census,
         pct_of_cases = 100 * cases / sum(cases)) |>
  arrange(desc(inc_per100k)) |>
  print(n = Inf)
 
# rate decline after ORI ended: hazard at ORI end vs three months later
rate_age_t |>
  filter(date %in% c(ori_end, ori_end + 90)) |>
  mutate(when = if_else(date == ori_end, "at_end", "plus_90d")) |>
  select(age_group, when, rate_ma) |>
  pivot_wider(names_from = when, values_from = rate_ma) |>
  mutate(at_end = r100k(at_end), plus_90d = r100k(plus_90d),
         fold_drop = at_end / plus_90d) |>
  print(n = Inf)
 
# ---- 4. Sentence-ready strings -------------------------------------------
 
snap_tab |>
  filter(bin_end == max(bin_end)) |>
  slice_max(cum_inc, n = 3) |>
  mutate(txt = sprintf("%s: %.0f per 100,000 (%d cases)",
                       district, cum_inc, cum_n)) |>
  pull(txt) |>
  writeLines()
 
rate_age_t |>
  summarise(cases = sum(n), n_census = first(n_census), .by = age_group) |>
  mutate(txt = sprintf("%s: %.0f per 100,000 over the episode (%d cases)",
                       age_group, 1e5 * cases / n_census, cases)) |>
  pull(txt) |>
  writeLines()

# ===========================================================================
# IS THE WESTERN ATTACK RATE THE HIGHEST, AND OVER WHAT RANGE?
# ===========================================================================
 
r100k <- \(x) x * 1e5
 
rate_dist_t <- rate_dist |> as_tibble()
 
# ---- 1. Cumulative incidence over the whole episode, by region -----------
# Recomputed from summed cases and population, not averaged district rates.
 
rate_dist_t |>
  summarise(cases = sum(n), n_census = first(n_census),
            .by = c(region, district)) |>
  summarise(n_dist = n(), cases = sum(cases), n_census = sum(n_census),
            .by = region) |>
  mutate(inc_per100k = 1e5 * cases / n_census) |>
  arrange(desc(inc_per100k)) |>
  print(n = Inf)
 
# and by district, so you can see whether West leads as a block or through
# one or two districts
rate_dist_t |>
  summarise(cases = sum(n), n_census = first(n_census),
            .by = c(region, district)) |>
  mutate(inc_per100k = 1e5 * cases / n_census) |>
  arrange(desc(inc_per100k)) |>
  print(n = Inf)
 
# ---- 2. Range of the daily hazard within each region ---------------------
# This is the "ranging from ... to ..." you want to quote.
 
rate_dist_t |>
  filter(!is.na(rate_ma), rate_ma > 0) |>
  summarise(min_per100k = r100k(min(rate_ma)),
            med_per100k = r100k(median(rate_ma)),
            max_per100k = r100k(max(rate_ma)),
            .by = region) |>
  arrange(desc(max_per100k)) |>
  print(n = Inf)
 
# peak hazard per district, which is the sharper version of the same claim
rate_dist_t |>
  filter(!is.na(rate_ma), rate_ma > 0) |>
  slice_max(rate_ma, n = 1, by = district) |>
  transmute(region, district, date, peak_per100k = r100k(rate_ma)) |>
  arrange(desc(peak_per100k)) |>
  print(n = Inf)
 
# ---- 3. Was West highest consistently, or only sometimes? ---------------
# Regional daily hazard, pooled within region, then ranked each day.
 
reg_daily <- rate_dist_t |>
  summarise(n = sum(n), n_census = sum(n_census), .by = c(region, date)) |>
  arrange(region, date) |>
  mutate(cum_cases = cumsum(n),
         at_risk   = n_census - lag(cum_cases, default = 0),
         rate      = n / at_risk,
         rate_ma   = slide_dbl(rate, mean, .before = ma_days - 1, .complete = TRUE),
         .by = region)
 
# share of days on which each region held the highest hazard
reg_daily |>
  filter(!is.na(rate_ma), rate_ma > 0) |>
  mutate(rk = rank(-rate_ma), .by = date) |>
  summarise(days_top = sum(rk == 1), days = n(), .by = region) |>
  mutate(pct_top = 100 * days_top / days) |>
  arrange(desc(pct_top)) |>
  print(n = Inf)
 
# monthly regional hazard, so you can see whether the lead holds over time
reg_daily |>
  filter(!is.na(rate_ma)) |>
  summarise(med_per100k = r100k(median(rate_ma)), .by = c(region, mo = floor_date(date, "month"))) |>
  pivot_wider(names_from = region, values_from = med_per100k) |>
  arrange(mo) |>
  print(n = Inf)
 
# ---- 4. Sentence-ready string -------------------------------------------
 
rate_dist_t |>
  filter(date < ori_start, !is.na(rate_ma), rate_ma > 0) |>
  summarise(min_per100k = r100k(min(rate_ma)),
            max_per100k = r100k(max(rate_ma)),
            n_days      = n_distinct(date),
            .by = region) |>
  mutate(txt = sprintf("%s: %.2f to %.2f per 100,000 per day (%d days)",
                       region, min_per100k, max_per100k, n_days)) |>
  arrange(desc(max_per100k)) |>
  pull(txt) |>
  writeLines()
```


## Serology

Get children in vaccination age in 3 hospitals, sampling between 01-08-2022 to 30-04-2024

```{r}
df_sero_3ch <- df_sero |> 
  filter(
    hospital %in% c("Children's hospital 1", "Children's hospital 2", "City Children's Hospital"),
    date_collection <= dmy("30-04-2024"),
    date_collection >= dmy("01-08-2022")
  )

df_sero_3ch_mod <- df_sero_3ch |> 
  filter(
    exact_age >= 9/12
  ) |> 
  mutate(
    age_5y = if_else(exact_age < 5, "<5", ">=5"),
    age_5y = fct_relevel(age_5y, "<5", ">=5"),
    age_ceiling = ceiling(exact_age)
  ) |> 
  left_join(df_census_age, by = join_by(age_ceiling == age)) |> 
  mutate(
    weight = n_census/mean(n_census)
  )
```

Check age by period

```{r}
brks     <- c(0, 9/12, 1, 2, 3, 4, 5, 10, 15)
age_labs <- c("<9m", "9-11m", "1y", "2y", "3y", "4y", "5-9y", "10-15y")

df_plot <- df_sero_3ch |>
  mutate(
    age_band   = factor(findInterval(exact_age, brks, rightmost.closed = TRUE),
                        levels = seq_along(age_labs), labels = age_labs),
    period_lab = factor(format(period_collect, "%b %Y"),
                        levels = format(sort(unique(period_collect)), "%b %Y")),
    hospital   = factor(case_when(str_detect(hospital, "1")    ~ "CH1",
                                  str_detect(hospital, "2")    ~ "CH2",
                                  str_detect(hospital, "City") ~ "CCH"),
                        levels = c("CH1", "CH2", "CCH"))
  ) |>
  count(hospital, period_lab, age_band) |>
  complete(nesting(hospital, period_lab), age_band, fill = list(n = 0))

xmax <- max(df_plot$n)

plot_hosp <- function(d, x_axis = FALSE, y_lab = NULL) {
  ggplot(d, aes(x = n, y = age_band, fill = age_band == "<9m")) +
    geom_col() +
    facet_grid(hospital ~ period_lab, switch = "y") +
    scale_fill_manual(values = c(`FALSE` = "grey40", `TRUE` = "#D55E00"), guide = "none") +
    scale_x_continuous(limits = c(0, xmax), breaks = scales::breaks_pretty(3),
                       expand = expansion(mult = c(0, 0.08))) +
    labs(x = if (x_axis) "Number of samples" else NULL, y = y_lab) +
    theme_bw() +
    theme(panel.grid.minor = element_blank(),
          strip.placement  = "outside",
          axis.text.x      = if (x_axis) element_text() else element_blank(),
          axis.ticks.x     = if (x_axis) element_line() else element_blank())
}

p_list <- df_plot |> group_split(hospital) |> map(droplevels)
n_h    <- length(p_list)

design <- p_list |>
  map_int(\(d) nlevels(d$period_lab)) |>
  imap(\(k, i) area(i, 1, i, k)) |>
  reduce(c)

p_age <- p_list |>
  imap(\(d, i) plot_hosp(d,
                         x_axis = i == n_h,
                         y_lab  = if (i == ceiling(n_h / 2)) "Age band" else NULL)) |>
  wrap_plots(design = design)
```

## Hospital coverage

```{r}
recode_hosp <- function(x) {
  factor(case_when(str_detect(x, "1")    ~ "CH1",
                   str_detect(x, "2")    ~ "CH2",
                   str_detect(x, "City") ~ "CCH"),
         levels = c("CH1", "CH2", "CCH"))
}

## ---------------- checks ----------------

stopifnot(!anyNA(recode_hosp(df_sero_3ch$hospital)))
setdiff(unique(df_sero_3ch$district), df_map_hcmc$district_blank)   # want character(0)
stopifnot(!any(duplicated(df_map_hcmc$district_blank)))

## ---------------- A: age panels ----------------

brks_age <- c(0, 9/12, 1, 2, 3, 4, 5, 10, 15)
age_labs <- c("<9m", "9-11m", "1y", "2y", "3y", "4y", "5-9y", "10-15y")

df_age <- df_sero_3ch |>
  mutate(
    age_band   = factor(findInterval(exact_age, brks_age, rightmost.closed = TRUE),
                        levels = seq_along(age_labs), labels = age_labs),
    period_lab = factor(format(period_collect, "%b %Y"),
                        levels = format(sort(unique(period_collect)), "%b %Y")),
    hospital   = recode_hosp(hospital)
  ) |>
  count(hospital, period_lab, age_band) |>
  complete(nesting(hospital, period_lab), age_band, fill = list(n = 0))

xmax <- max(df_age$n)

plot_hosp <- function(d, x_axis = FALSE, y_lab = FALSE) {
  ggplot(d, aes(x = n, y = age_band, fill = age_band == "<9m")) +
    geom_col() +
    facet_grid(hospital ~ period_lab, switch = "y") +
    scale_fill_manual(values = c(`FALSE` = "grey40", `TRUE` = "#D55E00"), guide = "none") +
    scale_x_continuous(
      limits = c(0, xmax),
      breaks = \(l) { b <- scales::breaks_pretty(3)(l); b[b <= l[2] * 0.85] },
      expand = expansion(mult = c(0, 0.08))
    ) +
    labs(x = if (x_axis) "Number of samples" else NULL, y = "Age band") +
    theme_bw() +
    theme(panel.grid.minor = element_blank(),
          strip.placement  = "outside",
          axis.title.y = element_text(colour = if (y_lab) "black" else NA),
          axis.text.x  = if (x_axis) element_text() else element_blank(),
          axis.ticks.x = if (x_axis) element_line() else element_blank())
}

p_list <- df_age |> group_split(hospital) |> map(droplevels)
n_h    <- length(p_list)

grobs <- p_list |>
  imap(\(d, i) plot_hosp(d, x_axis = i == n_h, y_lab = i == ceiling(n_h / 2))) |>
  map(ggplotGrob)

pan_l <- \(g) g$layout$l[startsWith(g$layout$name, "panel")]
g_ref <- grobs[[which.max(map_int(grobs, \(g) length(pan_l(g))))]]
gap   <- g_ref$widths[max(pan_l(g_ref)) - 1]

pad_to <- function(g, target) {
  need <- target - ncol(g)
  if (need <= 0) return(g)
  pos <- max(pan_l(g))
  w   <- rep(list(gap, unit(1, "null")), length.out = need)
  if (need %% 2 == 1) w[[need]] <- unit(0, "cm")
  for (i in seq_len(need)) {
    g   <- gtable_add_cols(g, w[[i]], pos)
    pos <- pos + 1
  }
  g
}

grobs       <- map(grobs, pad_to, target = max(map_int(grobs, ncol)))
grobs[-n_h] <- map(grobs[-n_h], \(g) gtable_add_rows(g, unit(4, "mm")))

p_age <- reduce(grobs, \(a, b) rbind(a, b, size = "max"))

cm_h  <- \(u) convertHeight(u, "cm", valueOnly = TRUE)
pan_t <- \(g) unique(g$layout$t[startsWith(g$layout$name, "panel")])
top_h <- \(g) cm_h(sum(g$heights[seq_len(pan_t(g) - 1)]))
bot_h <- \(g) cm_h(sum(g$heights[-seq_len(pan_t(g))]))

m_top <- top_h(grobs[[1]])
m_bot <- bot_h(grobs[[n_h]])
m_sp  <- bot_h(grobs[[1]]) + top_h(grobs[[2]])

## ---------------- B: maps, stacked, no legend inside ----------------

pct_min <- 1                                             # below this % draws grey
mid     <- 100 / n_distinct(df_map_hcmc$district_blank)  # equal-share reference
lab_low <- paste0("<", pct_min, "%")

df_map_pct <- df_sero_3ch |>
  mutate(hospital = recode_hosp(hospital)) |>
  count(hospital, district) |>
  complete(hospital, district = df_map_hcmc$district_blank, fill = list(n = 0)) |>
  mutate(pct = 100 * n / sum(n), .by = hospital) |>
  mutate(lfc = if_else(pct < pct_min, NA_real_, log2(pct / mid))) |>
  left_join(select(df_map_hcmc, district_blank), by = c("district" = "district_blank")) |>
  st_as_sf()

lo <- log2(pct_min / mid)                          # bar starts exactly at pct_min
hi <- max(df_map_pct$lfc, na.rm = TRUE)

brks_pct <- c(1, 2, 5, 10, 20, 40)
brks_pct <- brks_pct[brks_pct >= pct_min & log2(brks_pct / mid) <= hi]

bar_h <- 2.8
bar_w <- 0.35

cbar <- if (packageVersion("ggplot2") >= "3.5.0") {
  guide_colourbar(order = 1,
                  theme = theme(legend.key.height = unit(bar_h, "cm"),
                                legend.key.width  = unit(bar_w, "cm")))
} else {
  guide_colourbar(order = 1, barheight = unit(bar_h, "cm"), barwidth = unit(bar_w, "cm"))
}

lyr_low <- if (anyNA(df_map_pct$lfc)) {
  geom_sf(data = filter(df_map_pct, is.na(lfc)), aes(colour = lab_low),
          fill = "grey88", linewidth = 0.15, show.legend = "polygon")
}
 
p_map_col <- df_map_pct |>
  ggplot() +
  geom_sf(aes(fill = lfc), colour = "grey40", linewidth = 0.15) +
  facet_wrap(~ hospital, ncol = 1) +
  coord_sf(expand = FALSE) +
  scale_fill_gradientn(
    colours  = pal_gsea()(12),
    limits   = c(lo, hi),
    na.value = "grey88",
    rescaler = \(x, to = c(0, 1), from = range(x, na.rm = TRUE))
                 scales::rescale_mid(x, to, from, mid = 0),
    breaks   = log2(brks_pct / mid), labels = paste0(brks_pct, "%"),
    name     = "% of samples",
    guide    = guide_colourbar(barheight = unit(2.6, "cm"),
                               barwidth  = unit(0.35, "cm"))
  ) +
  theme_void() +
  theme(strip.text        = element_blank(),
        legend.position   = "right",
        legend.margin     = margin(l = 0.25, unit = "cm"),
        legend.title      = element_text(size = 10),
        legend.text       = element_text(size = 9),
        panel.spacing.y   = unit(0, "cm"),
        plot.margin       = margin(t = m_top, b = m_bot, l = 4, r = 0, unit = "cm"))
 
## ---------------- combine ----------------
# The age figure is a raw gtable; wrap it. The maps stay a normal ggplot so
# patchwork can size them. widths of 1.5:1 gives the map column enough room
# that coord_sf is height-limited rather than width-limited.
 
p_all <- wrap_elements(full = p_age) + p_map_col +
  plot_layout(widths = c(2, 1)) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag          = element_text(face = "bold", size = 14),
        plot.tag.position = c(0, 1),
        plot.margin       = margin(t = 10, r = 4, b = 4, l = 4))
 
p_all
# 
# ggsave("figs/chap1/age_and_map_combined.png", p_all, width = 9, height = 5, dpi = 300, bg = "white")
```

### GAM model

```{r}
sens   <- 0.99
spec   <- 0.95
critp  <- 0.95
n_grid <- 200                                  # points per fitted curve
w_date <- 1.4                                  # width of the time panel
pad_v  <- 1                                    # pt of padding above/below each row
pad_s  <- 2                                    # pt of padding inside facet strips

pals     <- hue_pal()(3)
col_age  <- c("<5" = pals[2], ">=5" = pals[3])
lab_all  <- "All periods"
lab_time <- "Over time"
lab_data <- "Data mean (95% CI)"
lab_gam  <- "GAM fit"

# ---- Helpers --------------------------------------------------------
prev_correct <- function(est_val, sens, spec) {
  (est_val + spec - 1) / (sens + spec - 1)
}

clamp01 <- function(x) pmin(pmax(x, 0), 1)

# Vectorised prop.test CI (Wilson score with continuity correction)
prop_ci <- function(x, n, conf = 0.95) {
  z     <- qnorm((1 + conf) / 2)
  est   <- x / n
  yates <- pmin(0.5, abs(x - n / 2))
  z22n  <- z^2 / (2 * n)
  pc_u  <- est + yates / n
  pc_l  <- est - yates / n
  su    <- pmin(pc_u, 1)
  sl    <- pmax(pc_l, 0)
  tibble(
    ci_lwr = if_else(pc_l <= 0, 0,
                     clamp01((sl + z22n - z * sqrt(sl * (1 - sl) / n + z22n / (2 * n))) / (1 + 2 * z22n))),
    ci_upr = if_else(pc_u >= 1, 1,
                     clamp01((su + z22n + z * sqrt(su * (1 - su) / n + z22n / (2 * n))) / (1 + 2 * z22n)))
  )
}

predict_ci <- function(model, data, var, ci = 0.95, n_grid = 200) {
  link_inv <- model$family$linkinv
  df_res   <- nrow(data) - length(model$coefficients)
  tq       <- qt(1 - (1 - ci) / 2, df_res)
  rng      <- range(data[[var]], na.rm = TRUE)
  newd     <- tibble(!!var := seq(rng[1], rng[2], length.out = n_grid))
  pr       <- predict(model, newdata = newd, se.fit = TRUE, type = "link")
  newd |>
    mutate(
      fit = link_inv(pr$fit),
      lwr = link_inv(pr$fit - tq * pr$se.fit),
      upr = link_inv(pr$fit + tq * pr$se.fit)
    )
}

adjust_prev <- function(d) {
  d |>
    mutate(across(c(fit, lwr, upr),
                  \(x) clamp01(prev_correct(x, sens, spec)),
                  .names = "{.col}_true"))
}

# ---- Data -----------------------------------------------------------
per_lvl <- df_sero_3ch_mod |>
  distinct(period_collect) |>
  arrange(period_collect) |>
  mutate(txt = format(period_collect, "%b %Y")) |>
  pull(txt) |>
  unique()

# Single place where period labels become a factor: keeps facets in date order
fct_period <- function(x) factor(as.character(x), levels = c(lab_all, per_lvl))

df_mod <- df_sero_3ch_mod |>
  mutate(
    pos        = as.integer(titer > 200),
    hospital   = factor(hospital),
    age_5y     = factor(age_5y, levels = c("<5", ">=5")),
    period_lab = fct_period(format(period_collect, "%b %Y"))
  ) |>
  filter(!is.na(pos), !is.na(exact_age), !is.na(date_collection))

hosp  <- levels(df_mod$hospital)
mid_h <- hosp[ceiling(length(hosp) / 2)]
end_h <- hosp[length(hosp)]
x_age <- c(0, ceiling(max(df_mod$exact_age, na.rm = TRUE)))

# Strip labels for the rows. Swap in the short version if the names crowd.
hosp_lab <- setNames(hosp, hosp)
# hosp_lab <- setNames(c("CH1", "CH2", "CCH"), hosp)

# ---- 1. Seroprevalence by collection date, per hospital x age group --
pred_date <- df_mod |>
  nest(.by = c(hospital, age_5y)) |>
  mutate(pred = map(data, \(d) {
    mod <- gam(pos ~ s(as.numeric(date_collection)),
               family = binomial, weights = weight,
               data = d, method = "REML")
    predict_ci(mod, d, "date_collection", n_grid = n_grid) |> adjust_prev()
  })) |>
  select(hospital, age_5y, pred) |>
  unnest(pred)

# Observed prevalence per period, for the point ranges
agg_sero <- df_mod |>
  summarise(npos = sum(pos), ntot = n(),
            .by = c(hospital, age_5y, period_collect)) |>
  mutate(prop_ci(npos, ntot)) |>
  mutate(
    prev      = npos / ntot,
    prev_true = clamp01(prev_correct(prev, sens, spec)),
    lwr_true  = clamp01(prev_correct(ci_lwr, sens, spec)),
    upr_true  = clamp01(prev_correct(ci_upr, sens, spec))
  )

# ---- 2. Seroprevalence by exact age: pooled + per period -------------
fit_age <- function(d) {
  nu <- n_distinct(d$exact_age)
  if (nrow(d) < 10 || nu < 4) return(NULL)          # too thin for a smooth
  mod <- gam(pos ~ s(exact_age, k = max(3, min(10, nu - 1))),
             family = binomial, data = d, method = "REML")
  predict_ci(mod, d, "exact_age", n_grid = n_grid) |> adjust_prev()
}

pred_age_per <- df_mod |>
  nest(.by = c(hospital, period_lab)) |>
  mutate(pred = map(data, possibly(fit_age, otherwise = NULL))) |>
  select(hospital, period_lab, pred) |>
  unnest(pred)

pred_age_all <- df_mod |>
  nest(.by = hospital) |>
  mutate(pred = map(data, possibly(fit_age, otherwise = NULL))) |>
  select(hospital, pred) |>
  unnest(pred) |>
  mutate(period_lab = fct_period(lab_all))

pred_age <- bind_rows(pred_age_all, pred_age_per) |>
  mutate(
    period_lab = fct_period(period_lab),
    agegr      = factor(if_else(exact_age < 5, "<5", ">=5"),
                        levels = c("<5", ">=5"))
  )

# Rug data: each observation once per period panel, once in the pooled panel
rug_age <- df_mod |> select(hospital, exact_age, pos, period_lab)
rug_age <- bind_rows(rug_age, mutate(rug_age, period_lab = fct_period(lab_all))) |>
  mutate(period_lab = fct_period(period_lab))

# ---- Shared theme ---------------------------------------------------
theme_sero <- theme_light() +
  theme(
    strip.placement  = "outside",
    strip.clip       = "off",
    strip.text       = element_text(color = "grey30",
                                    margin = margin(pad_s, pad_s, pad_s, pad_s)),
    strip.background = element_rect(fill = NA, color = NA),
    panel.spacing    = unit(0, "cm"),
    panel.border     = element_rect(color = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line        = element_line(color = "black")
  )

guides_sero <- guides(
  colour   = guide_legend(order = 1, override.aes = list(linetype = 1)),
  shape    = guide_legend(order = 2, override.aes = list(colour = "grey30", alpha = 1)),
  linetype = guide_legend(order = 3, override.aes = list(colour = "grey30"))
)

# ---- 3a. Left panel of each row: prevalence over time ---------------
# facet_grid gives a left strip for the hospital (inside the y axis title)
# and a top strip, which keeps this panel the same height as the age panels.
p_date <- map(hosp, \(h) {
  d_pt <- filter(df_mod,    hospital == h) |> mutate(facet = lab_time)
  d_pr <- filter(pred_date, hospital == h) |> mutate(facet = lab_time)
  d_ag <- filter(agg_sero,  hospital == h) |> mutate(facet = lab_time)
  
  ym <- tibble(d = sort(unique(d_pt$period_collect))) |>
    mutate(txt = if_else(!duplicated(year(d)),
                         format(d, "%b\n%Y"), format(d, "%b")))
  
  # Breaks sit at the start of each period, which can fall before the first
  # sample date, so span the limits over both or the label gets dropped
  x_rng <- range(c(ym$d, d_pt$date_collection, d_ag$period_collect + days(15)))
  
  ggplot() +
    geom_hline(yintercept = critp, colour = "darkred", linetype = "dashed") +
    geom_point(data = d_pt, aes(date_collection, pos), shape = 124, alpha = 0.3) +
    geom_ribbon(data = d_pr,
                aes(date_collection, ymin = lwr_true, ymax = upr_true, fill = age_5y),
                alpha = 0.15, show.legend = FALSE) +
    geom_line(data = d_pr,
              aes(date_collection, fit_true, colour = age_5y, linetype = lab_gam)) +
    geom_pointrange(data = d_ag,
                    aes(period_collect + days(15), prev_true,
                        ymin = lwr_true, ymax = upr_true,
                        colour = age_5y, shape = lab_data),
                    alpha = 0.6,
                    show.legend = c(colour = FALSE, fill = FALSE, shape = TRUE)) +
    facet_grid(hospital ~ facet, switch = "y",
               labeller = labeller(hospital = hosp_lab)) +
    scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(-0.1, 1.1),
                       labels = label_percent(suffix = "")) +
    scale_x_date(breaks = ym$d, labels = ym$txt, limits = x_rng,
                 expand = expansion(mult = 0.04)) +
    scale_colour_manual(values = col_age, name = "Age group") +
    scale_fill_manual(values = col_age, guide = "none") +
    scale_shape_manual(values = setNames(16, lab_data), name = NULL) +
    scale_linetype_manual(values = setNames(1, lab_gam), name = NULL) +
    labs(x = NULL, y = "Seroprevalence (%)") +
    guides_sero +
    theme_sero +
    # NA keeps the horizontal space, so panels start at the same x in every row
    theme(axis.title.y = element_text(colour = if (h == mid_h) "black" else NA))
})
names(p_date) <- hosp

# ---- 3b. Right panels of each row: pooled + one panel per period ----
p_age <- map(hosp, \(h) {
  d_pt <- filter(rug_age,  hospital == h)
  d_pr <- filter(pred_age, hospital == h)
  
  ggplot() +
    geom_hline(yintercept = critp, colour = "darkred", linetype = "dashed") +
    geom_point(data = d_pt, aes(exact_age, pos), shape = 124, alpha = 0.3) +
    geom_ribbon(data = d_pr,
                aes(exact_age, ymin = lwr_true, ymax = upr_true, fill = agegr),
                alpha = 0.15, show.legend = FALSE) +
    geom_line(data = d_pr,
              aes(exact_age, fit_true, colour = agegr,
                  linetype = lab_gam, group = 1)) +
    facet_wrap(~period_lab, nrow = 1, drop = TRUE) +
    scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(-0.1, 1.1),
                       labels = label_percent(suffix = "")) +
    scale_x_continuous(breaks = seq(x_age[1], x_age[2], 5), limits = x_age,
                       expand = expansion(mult = 0.02)) +
    scale_colour_manual(values = col_age, name = "Age group") +
    scale_fill_manual(values = col_age, guide = "none") +
    scale_linetype_manual(values = setNames(1, lab_gam), name = NULL) +
    labs(x = "Age (years)", y = NULL) +
    guides_sero +
    theme_sero +
    # x title dropped entirely on upper rows: it only costs height, not alignment
    theme(axis.text.y     = element_blank(),
          axis.ticks.y    = element_blank(),
          panel.spacing.x = unit(0.4, "lines"),
          axis.title.x    = if (h == end_h) element_text() else element_blank())
})
names(p_age) <- hosp

# ---- 4. Assemble: one row per hospital ------------------------------
n_per <- df_mod |> distinct(hospital, period_lab) |> count(hospital) |> deframe()
n_per <- n_per + 1                             # the pooled panel
n_max <- max(n_per)

fig <- map(hosp, \(h) {
  k  <- n_per[[h]]
  ps <- list(p_date[[h]], p_age[[h]])
  ws <- c(w_date, k)
  if (k < n_max) {                             # pad so panels align across rows
    ps <- c(ps, list(plot_spacer()))
    ws <- c(ws, n_max - k)
  }
  wrap_plots(ps, nrow = 1, widths = ws)
}) |>
  wrap_plots(ncol = 1, guides = "collect") &
  theme(legend.position = "bottom",
        plot.margin     = margin(t = pad_v, r = 2, b = pad_v, l = 2))

# fig
# 
# ggsave("figs/chap1/fig_seroprev_by_hospital_period.png", fig, width = 9, height = 4.5, dpi = 300)
```

### Sero values

```{r}
# ===========================================================================
# SUMMARY VALUES FOR THE SEROPREVALENCE-BY-AGE FIGURE
# ===========================================================================
# All *_true columns are corrected for assay sensitivity and specificity;
# quote those, not the raw values.
 
pc <- \(x) sprintf("%.1f", 100 * x)
 
# ---- 1. Sample sizes behind each panel ------------------------------------
 
df_mod |>
  summarise(n = n(), npos = sum(pos), .by = c(hospital, age_5y)) |>
  pivot_wider(names_from = age_5y, values_from = c(n, npos)) |>
  print(n = Inf)
 
df_mod |> summarise(n = n(), min_age = min(exact_age), max_age = max(exact_age))
 
# ---- 2. Observed seroprevalence per period --------------------------------
 
agg_sero |>
  mutate(lab = sprintf("%s%% (%s to %s), %d/%d",
                       pc(prev_true), pc(lwr_true), pc(upr_true), npos, ntot)) |>
  select(hospital, age_5y, period_collect, ntot, prev_true, lab) |>
  arrange(hospital, age_5y, period_collect) |>
  print(n = Inf)
 
# pooled across periods, per hospital and age group
df_mod |>
  summarise(npos = sum(pos), ntot = n(), .by = c(hospital, age_5y)) |>
  mutate(prop_ci(npos, ntot)) |>
  mutate(prev_true = clamp01(prev_correct(npos / ntot, sens, spec)),
         lwr_true  = clamp01(prev_correct(ci_lwr, sens, spec)),
         upr_true  = clamp01(prev_correct(ci_upr, sens, spec)),
         lab = sprintf("%s%% (%s to %s), %d/%d",
                       pc(prev_true), pc(lwr_true), pc(upr_true), npos, ntot)) |>
  select(hospital, age_5y, lab) |>
  print(n = Inf)
 
# range across periods: how much the point estimates moved
agg_sero |>
  summarise(n_periods = n(),
            min = min(prev_true), max = max(prev_true),
            range_pp = 100 * (max(prev_true) - min(prev_true)),
            .by = c(hospital, age_5y)) |>
  print(n = Inf)
 
# how many period estimates sit below the critical threshold
agg_sero |>
  summarise(n = n(), n_below = sum(prev_true < critp),
            n_ci_below = sum(upr_true < critp),      # CI entirely below
            .by = c(hospital, age_5y)) |>
  print(n = Inf)
 
# ---- 3. Fitted age curve: values at reference ages ------------------------
# Nearest grid point to each reference age, from the pooled ("All periods")
# fit per hospital.
 
ref_ages <- c(1, 2, 3, 5, 10, 15)
 
pred_age |>
  filter(period_lab == lab_all) |>
  cross_join(tibble(ref = ref_ages)) |>
  slice_min(abs(exact_age - ref), n = 1, by = c(hospital, ref)) |>
  mutate(lab = sprintf("%s%% (%s to %s)",
                       pc(fit_true), pc(lwr_true), pc(upr_true))) |>
  select(hospital, ref, age = exact_age, fit_true, lab) |>
  arrange(hospital, ref) |>
  print(n = Inf)
 
# minimum of the pooled curve and where it occurs: the susceptibility trough
pred_age |>
  filter(period_lab == lab_all) |>
  slice_min(fit_true, n = 1, by = hospital) |>
  mutate(lab = sprintf("%s%% (%s to %s) at age %.1f y",
                       pc(fit_true), pc(lwr_true), pc(upr_true), exact_age)) |>
  select(hospital, exact_age, fit_true, lab) |>
  print(n = Inf)
 
# age range over which the fitted curve sits below the critical threshold
pred_age |>
  filter(period_lab == lab_all, fit_true < critp) |>
  summarise(age_from = min(exact_age), age_to = max(exact_age),
            n_grid = n(), .by = hospital) |>
  print(n = Inf)
 
# and where the upper CI is entirely below it, which is the stronger claim
pred_age |>
  filter(period_lab == lab_all, upr_true < critp) |>
  summarise(age_from = min(exact_age), age_to = max(exact_age), .by = hospital) |>
  print(n = Inf)
 
# ---- 4. Fitted time curve: start, end, and change -------------------------
 
pred_date |>
  slice(c(which.min(date_collection), which.max(date_collection)),
        .by = c(hospital, age_5y)) |>
  mutate(point = if_else(date_collection == min(date_collection), "start", "end"),
         .by = c(hospital, age_5y)) |>
  select(hospital, age_5y, point, fit_true, lwr_true, upr_true) |>
  pivot_wider(names_from = point, values_from = c(fit_true, lwr_true, upr_true)) |>
  mutate(change_pp = 100 * (fit_true_end - fit_true_start)) |>
  print(n = Inf)
 
# ---- 5. Under-5 versus 5-and-over, the headline contrast ------------------
 
df_mod |>
  summarise(npos = sum(pos), ntot = n(), .by = age_5y) |>
  mutate(prop_ci(npos, ntot)) |>
  mutate(prev_true = clamp01(prev_correct(npos / ntot, sens, spec)),
         lwr_true  = clamp01(prev_correct(ci_lwr, sens, spec)),
         upr_true  = clamp01(prev_correct(ci_upr, sens, spec)),
         lab = sprintf("%s%% (%s to %s), %d/%d",
                       pc(prev_true), pc(lwr_true), pc(upr_true), npos, ntot)) |>
  select(age_5y, lab) |>
  print()
 
# ---- 6. Sentence-ready strings --------------------------------------------
 
pred_age |>
  filter(period_lab == lab_all) |>
  slice_min(fit_true, n = 1, by = hospital) |>
  mutate(txt = sprintf("%s: fitted seroprevalence reaches a minimum of %s%% (%s to %s) at age %.1f years",
                       hospital, pc(fit_true), pc(lwr_true), pc(upr_true), exact_age)) |>
  pull(txt) |>
  writeLines()
```


## Vaccination coverage

Count children born since 1/1/2017

```{r}
df_vax_hcmc |> 
  filter(dob >= dmy("01-01-2017")) |> 
  nrow()
```

## Sero vs coverage

```{r}
set.seed(2026)
 
study_start <- as.Date("2022-09-05")
study_end   <- as.Date("2024-04-30")
 
# IMPORTANT: set these to the sero collection window, not the whole study.
# MCV2 coverage rises ~23 points over 21 months, so a full-period min-max
# bar spans most of the axis and the x position stops meaning anything.
#   win_start <- min(as.Date(df_mod$date_collect), na.rm = TRUE)
#   win_end   <- max(as.Date(df_mod$date_collect), na.rm = TRUE)
win_start <- study_start
win_end   <- study_end
 
m_draws     <- 1000L   # Monte Carlo draws for VE parameters
n_min       <- 5       # minimum sera per region to plot
diag_expand <- 0.35    # fraction of the x range the 1:1 line should cross
 
wk1    <- floor_date(study_start, "week", week_start = 1)
weeks  <- seq(wk1, study_end, by = "7 days")
n_week <- length(weeks)
 
# index of the first week whose Monday falls on or after date d
wk_ge <- function(d) (as.integer(unclass(d) - unclass(wk1)) + 6L) %/% 7L + 1L
 
# calendar-month arithmetic on distinct values only, then matched back
add_months_fast <- function(d, n) {
  u <- unique(d)
  (u %m+% months(n))[match(d, u)]
}
 
# ===========================================================================
# 1. CHILD DATA
# ===========================================================================
 
assign_region <- function(x) {
  r <- case_when(
    is.na(x) | x == ""                   ~ NA_character_,
    str_detect(x, "binh chanh|binh tan") ~ "west",
    str_detect(x, "hoc mon|cu chi|12")   ~ "north",
    str_detect(x, "thu duc")             ~ "east",
    str_detect(x, "nha be|can gio|7")    ~ "south",
    .default = "central"
  )
  factor(str_to_title(r), levels = c("West", "North", "East", "South", "Central"))
}
 
child <- df_vax_hcmc |>
  as_tibble() |>
  transmute(
    district = district_blank,
    region   = assign_region(district_blank),
    dob = as.Date(dob),
    d1  = as.Date(date_m1),
    d2  = as.Date(date_m2),
    t9  = add_months_fast(dob,  9L),     # eligible for dose 1
    t18 = add_months_fast(dob, 18L),     # eligible for dose 2
    t60 = add_months_fast(dob, 60L)      # turns 5, leaves the denominator
  ) |>
  filter(dob <= study_end, t60 > wk1, !is.na(district), district != "")
 
# CHECK: "central" is the catch-all, so any typo lands there. Read this.
child |> count(region, district, sort = TRUE) |> print(n = Inf)
 
# ---- dose-level quantities ------------------------------------------------
# EVC denominator = the dose 1 population (9-59 mo). Waning is timed from the
# dose date; a child with only a dose 2 record is treated as single-dose.
 
a_den      <- wk_ge(child$t9)
b_exit     <- wk_ge(child$t60) - 1L
t12        <- add_months_fast(child$dob, 12L)
first_date <- if_else(is.na(child$d1), child$d2, child$d1)
s_first    <- wk_ge(first_date)
s_second   <- if_else(is.na(child$d1), NA_integer_, wk_ge(child$d2))
band_first <- if_else(first_date < t12, 1L, 2L)   # 1 = 9-11 mo, 2 = >= 12 mo
b_one      <- if_else(is.na(s_second), b_exit, pmin(b_exit, s_second - 1L))
 
sum(is.na(child$d1) & !is.na(child$d2))   # orphan dose 2 records
 
# ===========================================================================
# 2. VE PARAMETERS AND MONTE CARLO DRAWS
# ===========================================================================
# Normal on the logit scale for the VE proportions (keeps draws inside (0,1),
# which matters at 99.6%), log scale for the reversion rate (keeps it
# positive). VE depends on dose count, not a product across doses, because
# the MCV2 estimate is a two-dose effectiveness.
# NOTE the 9-11 mo CI is asymmetric on the logit scale, so centring on the
# best estimate gives a sampled interval of about (66%, 94%) against the
# reported (72%, 95%): wider and lower, i.e. conservative.
 
ci_sd <- function(lo, hi, trans) (trans(hi) - trans(lo)) / 3.92
 
draws <- tibble(
  ve1_L = plogis(rnorm(m_draws, qlogis(0.840), ci_sd(0.720, 0.950, qlogis))),
  ve1_H = plogis(rnorm(m_draws, qlogis(0.925), ci_sd(0.848, 0.970, qlogis))),
  ve2   = plogis(rnorm(m_draws, qlogis(0.996), ci_sd(0.993, 0.998, qlogis))),
  p_rev = exp(rnorm(m_draws, log(0.0022), ci_sd(0.0017, 0.0028, log)))
)
 
q_wk <- -log1p(-draws$p_rev) * 7 / 365.25   # weekly decay from annual reversion
 
draws |>                                    # sampled vs reported intervals
  summarise(across(everything(),
                   list(med = median,
                        lo  = \(x) quantile(x, 0.025),
                        hi  = \(x) quantile(x, 0.975))))
 
# ===========================================================================
# 3. COUNTING ENGINE
# ===========================================================================
# Each child contributes intervals to numerator and denominator, so counting
# is a +1/-1 event table plus a cumsum: O(n_children), not O(n x weeks).
# For EVC the table gains a dose-week dimension, because
#   sum_i VE_i(w) = mult * exp(-q w) * sum_s N[w, s] exp(q s),
# so the registry is scanned once and the M draws cost two small matrix
# products per term rather than M passes over the children.
 
cover_counts_g <- function(a, b, g, ng) {
  a  <- pmax(a, 1L)
  b  <- pmin(b, n_week)
  ok <- !is.na(a) & !is.na(b) & !is.na(g) & b >= a
  a <- a[ok]; b <- b[ok]; g <- g[ok]
 
  nb  <- ng * (n_week + 1L)
  off <- (g - 1L) * (n_week + 1L)
  d   <- tabulate(off + a, nb) - tabulate(off + b + 1L, nb)
 
  m <- matrix(d, nrow = n_week + 1L, ncol = ng) |> apply(2, cumsum)
  m[seq_len(n_week), , drop = FALSE]
}
 
key_cover_counts_g <- function(a, b, g, key, ng) {
  a  <- pmax(a, 1L)
  b  <- pmin(b, n_week)
  ok <- !is.na(a) & !is.na(b) & !is.na(g) & !is.na(key) & b >= a
  a <- a[ok]; b <- b[ok]; g <- g[ok]; key <- key[ok]
 
  z   <- min(key):max(key)
  nk  <- length(z)
  k   <- key - z[1L] + 1L
  nb  <- ng * nk * (n_week + 1L)
  off <- ((g - 1L) * nk + (k - 1L)) * (n_week + 1L)
  d   <- tabulate(off + a, nb) - tabulate(off + b + 1L, nb)
 
  m <- matrix(d, nrow = n_week + 1L) |> apply(2, cumsum)
  list(n = m[seq_len(n_week), , drop = FALSE], z = z, nk = nk)
}
 
term_matrices_g <- function(a, b, g, key, mult, ng) {
  kc <- key_cover_counts_g(a, b, g, key, ng)
  Ez <- exp(outer(kc$z, q_wk))                 # n_key  x M
  Ew <- exp(outer(seq_len(n_week), -q_wk))     # n_week x M
  lapply(seq_len(ng), \(gg) {
    cols <- ((gg - 1L) * kc$nk + 1L):(gg * kc$nk)
    (kc$n[, cols, drop = FALSE] %*% Ez) * Ew * rep(mult, each = n_week)
  })
}
 
# ===========================================================================
# 4. COVERAGE AND EVC FOR ANY GROUPING
# ===========================================================================
 
build_series <- function(grp) {
  lev <- levels(droplevels(grp))
  ng  <- length(lev)
  gi  <- as.integer(factor(as.character(grp), levels = lev))
 
  mk <- function(entry, exit, vax, dose_lab) {
    b   <- wk_ge(exit) - 1L
    a_d <- wk_ge(entry)
    a_n <- pmax(a_d, wk_ge(vax))      # doses before entry count from entry
    tibble(
      week    = rep(weeks, times = ng),
      group   = rep(lev, each = n_week),
      dose    = dose_lab,
      n_denom = as.vector(cover_counts_g(a_d, b, gi, ng)),
      n_vax   = as.vector(cover_counts_g(a_n, b, gi, ng))
    )
  }
 
  cov_df <- bind_rows(
    mk(child$t9,  child$t60, child$d1, "Dose 1"),
    mk(child$t18, child$t60, child$d2, "Dose 2")
  ) |>
    mutate(group    = factor(group, levels = lev),
           coverage = if_else(n_denom > 0, n_vax / n_denom, NA_real_))
 
  numM <- replicate(ng, matrix(0, n_week, m_draws), simplify = FALSE)
  for (bb in 1:2) {                                   # single-dose periods
    sel <- which(!is.na(s_first) & band_first == bb)
    if (length(sel)) numM <- Map(`+`, numM, term_matrices_g(
      pmax(a_den, s_first)[sel], b_one[sel], gi[sel], s_first[sel],
      if (bb == 1L) draws$ve1_L else draws$ve1_H, ng))
  }
  sel <- which(!is.na(s_second))                      # two-dose periods
  if (length(sel)) numM <- Map(`+`, numM, term_matrices_g(
    pmax(a_den, s_second)[sel], b_exit[sel], gi[sel], s_second[sel],
    draws$ve2, ng))
 
  denE <- cover_counts_g(a_den, b_exit, gi, ng)
 
  evc_mat <- map(seq_len(ng), \(gg) {
    dv <- denE[, gg]; dv[dv == 0] <- NA_real_
    numM[[gg]] / dv
  })
 
  evc_df <- map2(evc_mat, lev, \(E, gg) {
    qs <- apply(E, 1, quantile, probs = c(0.5, 0.025, 0.975), na.rm = TRUE)
    tibble(week = weeks, group = gg,
           evc = qs[1, ], evc_lo = qs[2, ], evc_hi = qs[3, ])
  }) |>
    bind_rows() |>
    mutate(group = factor(group, levels = lev))
 
  # ---- period summaries over the comparison window ----
  idx <- which(weeks >= win_start & weeks <= win_end)
 
  evc_win <- map2(seq_len(ng), lev, \(gg, nm) {
    dsum <- sum(denE[idx, gg])
    pa   <- if (dsum > 0) colSums(numM[[gg]][idx, , drop = FALSE]) / dsum
            else rep(NA_real_, m_draws)
    wk <- evc_df |> filter(group == nm, week >= win_start, week <= win_end)
    tibble(group = nm,
           x     = median(pa),
           xu_lo = quantile(pa, 0.025, na.rm = TRUE),
           xu_hi = quantile(pa, 0.975, na.rm = TRUE),
           x_lo  = min(wk$evc, na.rm = TRUE),
           x_hi  = max(wk$evc, na.rm = TRUE))
  }) |>
    bind_rows()
 
  cov_win <- cov_df |>
    filter(week >= win_start, week <= win_end, n_denom > 0) |>
    summarise(x = sum(n_vax) / sum(n_denom),
              x_lo = min(coverage), x_hi = max(coverage),
              .by = c(group, dose)) |>
    mutate(xu_lo = NA_real_, xu_hi = NA_real_)
 
  list(cov = cov_df, evc = evc_df, cov_win = cov_win, evc_win = evc_win)
}
 
# ===========================================================================
# 5. RUN, RANK, PALETTE
# ===========================================================================
 
ser_region <- build_series(child$region)
 
# Dose 1 is near-flat across regions, so dose 2 separates them. Pooled over
# the period, so the order does not hinge on one week.
rank_order <- ser_region$cov |>
  filter(dose == "Dose 2") |>
  summarise(m = sum(n_vax) / sum(n_denom), .by = group) |>
  arrange(desc(m)) |>
  pull(group) |>
  as.character()
 
rank_order
 
pal_region <- setNames(
  c("#332288", "#88CCEE", "#117733", "#DDCC77", "#CC6677"),
  rank_order
)
 
# ---- checks ---------------------------------------------------------------
 
ser_region$cov |> filter(coverage > 1) |> nrow()                  # should be 0
ser_region$evc |> summarise(lo = min(evc_lo), hi = max(evc_hi))   # in [0, 1]
 
# ===========================================================================
# 6. SEROPREVALENCE, AGE-MATCHED TO EACH DENOMINATOR
# ===========================================================================
# MCV1 and EVC use a 9-59 month denominator; MCV2 uses 18-59 months. Sera are
# restricted to the matching window for each panel so x and y describe the
# same population.
 
sero_src <- df_mod |>
  as_tibble() |>
  mutate(age_y  = exact_age,          # age in years, decimal
         region = assign_region(district)) |>
  filter(!is.na(pos), !is.na(age_y), !is.na(district), district != "")
 
sero_src |>                                      # sanity-check the age column
  summarise(n = n(), min = min(age_y), med = median(age_y), max = max(age_y))
 
# CHECK: unmatched district strings drop silently
setdiff(unique(sero_src$district), unique(child$district))
 
sero_ci <- function(df, key) {
  df |>
    summarise(n_sero = n(), x_pos = sum(pos), .by = all_of(key)) |>
    mutate(
      sero    = x_pos / n_sero,
      sero_lo = if_else(x_pos == 0, 0, qbeta(0.025, x_pos, n_sero - x_pos + 1)),
      sero_hi = if_else(x_pos == n_sero, 1, qbeta(0.975, x_pos + 1, n_sero - x_pos))
    ) |>
    rename(group = all_of(key))
}
 
sero_for <- function(lo, hi) {
  sero_src |>
    filter(age_y >= lo, age_y < hi) |>
    mutate(region = as.character(region)) |>
    sero_ci("region")
}
 
x_labels <- c("Dose 1" = "MCV1 coverage",
              "Dose 2" = "MCV2 coverage",
              "EVC"    = "Estimated immunity")
x_short  <- c("MCV1 coverage"      = "MCV1",
              "MCV2 coverage"      = "MCV2",
              "Estimated immunity" = "EVC")
 
age_win <- tibble(
  metric = unname(x_labels),
  age_lo = c(9 / 12, 18 / 12, 9 / 12),   # years, inclusive
  age_hi = c(5, 5, 5)                    # years, exclusive
)
 
sero_all <- age_win |>
  mutate(res = map2(age_lo, age_hi, sero_for)) |>
  unnest(res) |>
  select(metric, group, n_sero, x_pos, sero, sero_lo, sero_hi)
 
# how many sera survive each age window, and how many regions clear n_min
sero_all |>
  summarise(n_sera = sum(n_sero), n_groups = sum(n_sero >= n_min), .by = metric)
 
# ===========================================================================
# 7. SCATTER DATA
# ===========================================================================
 
scat <- bind_rows(
  ser_region$cov_win |>
    mutate(metric = unname(x_labels[as.character(dose)])) |> select(-dose),
  ser_region$evc_win |> mutate(metric = x_labels[["EVC"]])
) |>
  mutate(group = as.character(group)) |>
  inner_join(sero_all, by = c("metric", "group")) |>
  filter(n_sero >= n_min) |>
  mutate(metric = factor(metric, levels = unname(x_labels)),
         region = factor(group, levels = rank_order))
 
# the gap that dominates panel B: seroprevalence minus x
scat |>
  summarise(med_gap = median(sero - x), lo = min(sero - x), hi = max(sero - x),
            .by = metric)
 
# ===========================================================================
# 8. THEME
# ===========================================================================
 
theme_epi <- function() {
  theme_bw() +
    theme(
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background  = element_rect(fill = "white", colour = NA),
      panel.border     = element_rect(fill = NA, colour = "black"),
      panel.grid       = element_blank(),
 
      axis.line   = element_blank(),
      axis.ticks  = element_line(colour = "black"),
      axis.text   = element_text(colour = "black"),
      axis.text.x = element_text(vjust = 1),
      axis.title  = element_text(face = "bold", colour = "black"),
 
      strip.background = element_blank(),
      strip.text       = element_text(colour = "black"),
 
      legend.position   = "bottom",
      legend.title      = element_text(face = "bold", colour = "black"),
      legend.text       = element_text(colour = "black"),
      legend.key        = element_blank(),
      legend.background = element_blank()
    )
}
 
make_x_breaks <- function(rng, by = "3 months", min_gap_days = 70) {
  inner <- seq(floor_date(rng[1], "month"), rng[2], by = by)
  keep  <- as.numeric(inner - rng[1]) >= min_gap_days &
           as.numeric(rng[2] - inner)  >= min_gap_days
  sort(unique(c(rng[1], inner[keep], rng[2])))
}
 
x_breaks <- make_x_breaks(range(weeks))
 
# month on line 1, year on line 2 for the first break of each year
label_month_year <- function(x) {
  m <- format(x, "%b"); y <- format(x, "%Y")
  out <- if_else(!duplicated(y), paste0(m, "\n", y), paste0(m, "\n "))
  out[is.na(x)] <- NA_character_
  out
}
 
# ===========================================================================
# 9. PANEL A: WEEKLY SERIES BY REGION, WITH EVC UNCERTAINTY RIBBON
# ===========================================================================
 
lines_region <- bind_rows(
  ser_region$cov |> transmute(week, group = as.character(group),
                              series = unname(x_labels[as.character(dose)]),
                              value = coverage),
  ser_region$evc |> transmute(week, group = as.character(group),
                              series = x_labels[["EVC"]], value = evc)
) |>
  mutate(series = factor(series, levels = unname(x_labels)),
         region = factor(group, levels = rank_order))
 
rib_region <- ser_region$evc |>
  transmute(week, group = as.character(group), evc_lo, evc_hi) |>
  mutate(region = factor(group, levels = rank_order))
 
p_cov_evc <- ggplot(lines_region,
                    aes(week, value, colour = region, linetype = series,
                        group = interaction(group, series))) +
  geom_ribbon(data = rib_region, inherit.aes = FALSE,
              aes(week, ymin = evc_lo, ymax = evc_hi,
                  fill = region, group = group),
              alpha = 0.15, colour = NA, show.legend = FALSE) +
  geom_line() +
  scale_colour_manual(values = pal_region) +
  scale_fill_manual(values = pal_region) +
  scale_linetype_manual(values = setNames(c("solid", "dashed", "dotted"),
                                          unname(x_labels)),
                        labels = x_short) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_x_date(breaks = x_breaks, labels = label_month_year,
               expand = expansion(mult = 0.03)) +
  guides(colour   = guide_legend(order = 1,
                                 override.aes = list(linetype = "solid")),
         linetype = guide_legend(order = 2,
                                 override.aes = list(colour = "grey30"))) +
  labs(x = NULL, y = "Coverage", colour = "Region", linetype = "Series") +
  theme_epi()
 
# ===========================================================================
# 10. PANEL B: SEROPREVALENCE AGAINST COVERAGE AND IMMUNITY
# ===========================================================================
# Horizontal bar = range over the window, NOT a confidence interval.
# Faded bar on the EVC panel = 95% VE-parameter interval on the period mean.
# Dotted diagonal = 1:1.
# The 1:1 line only enters a panel where the x and y ranges overlap, which
# they do not here, so the y limit is extended just enough for the line to
# cross the top-left corner.
 
diag_lim <- scat |>
  summarise(x1 = min(c(x_lo, xu_lo), na.rm = TRUE),
            x2 = max(c(x_hi, xu_hi), na.rm = TRUE),
            y1 = min(sero_lo, na.rm = TRUE),
            y2 = max(sero_hi, na.rm = TRUE),
            .by = metric) |>
  mutate(span = x2 - x1,
         y2n  = pmin(1, pmax(y2, x1 + diag_expand * span)),
         y1n  = pmax(0, pmin(y1, x2 - diag_expand * span))) |>
  pivot_longer(c(y1n, y2n), values_to = "sero") |>
  transmute(metric, x = x1, sero)
 
p_sero <- ggplot(scat, aes(x, sero, colour = region)) +
  {if (diag_expand > 0) geom_blank(data = diag_lim, inherit.aes = FALSE,
                                   aes(x = x, y = sero))} +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dotted", colour = "grey50") +
  geom_segment(aes(x = xu_lo, xend = xu_hi, y = sero, yend = sero),
               alpha = 0.35, na.rm = TRUE) +
  geom_segment(aes(x = x_lo, xend = x_hi, y = sero, yend = sero)) +
  geom_segment(aes(x = x, xend = x, y = sero_lo, yend = sero_hi)) +
  facet_wrap(~ metric, nrow = 1, scales = "free") +
  scale_colour_manual(values = pal_region) +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  guides(colour = "none") +          # Region legend collected from A
  labs(x = NULL, y = "Seroprevalence") +
  theme_epi()
 
# ===========================================================================
# 11. PANEL B: MCV COVERAGE ACROSS SOUTHERN VIETNAM
# ===========================================================================
 
province_south_raw <- c("Bình Dương", "Bình Phước", "Tây Ninh",
                        "Bà Rịa Vũng Tàu", "Đồng Nai", "Long An", "Tiền Giang",
                        "Bến Tre", "Vĩnh Long", "Trà Vinh", "Đồng Tháp",
                        "Hậu Giang", "An Giang", "Kiên Giang", "Bạc Liêu",
                        "Sóc Trăng", "Cà Mau", "Hồ Chí Minh", "Cần Thơ")
 
# same cleaning as applied to the dataframe
province_south_blank <- province_south_raw |>
  stri_trans_general("Latin-ASCII") |>
  str_to_lower() |>
  str_squish() |>
  str_replace_all("đ", "d")
 
map_south <- df_map_vn |>
  left_join(df_vax_vn, by = "province_blank") |>
  filter(province_blank %in% province_south_blank)
 
# CHECK: provinces failing the name match drop silently and draw grey
setdiff(province_south_blank, map_south$province_blank)
map_south |> st_drop_geometry() |> filter(is.na(measles_pct)) |> pull(province_blank)
 
hcmc_poly <- map_south |> filter(province_blank == "ho chi minh")
 
p_map_south <- ggplot() +
  geom_sf(data = map_south, aes(fill = measles_pct),
          colour = "white", linewidth = 0.2) +
  geom_sf(data = hcmc_poly, fill = NA, colour = "white", linewidth = 1.5) +
  geom_sf(data = hcmc_poly, fill = NA, colour = "grey10", linewidth = 0.6) +
  scale_fill_gsea(reverse = TRUE,
                  labels = scales::percent_format(scale = 1, accuracy = 1),
                  na.value = "grey90",
                  breaks = scales::pretty_breaks(4),
                  guide = guide_colourbar(barwidth  = unit(0.3, "cm"),
                                          barheight = unit(2.5, "cm"),
                                          ticks.colour = "white",
                                          frame.colour = NA)) +
  labs(fill = NULL) +
  theme_void(base_size = 11) +
  theme(legend.position    = "bottom",
        legend.direction   = "horizontal",
        legend.text        = element_text(colour = "black", size = 8),
        legend.box.spacing = unit(2, "pt"),
        plot.margin        = margin(2, 2, 2, 2))
 
# ===========================================================================
# 12. COMBINED FIGURE
# ===========================================================================
# No guide collection anywhere. The region and series legend is extracted
# from panel A as its own object and placed in an explicit grid cell, so it
# behaves like a plot and cannot be repositioned by patchwork.
 
# --- legend, extracted from A -----------------------------------------------
 
p_a <- p_cov_evc + theme(legend.position = "none")
 
p_b <- p_map_south +
  guides(fill = guide_colourbar(barwidth  = unit(3.5, "cm"),
                                barheight = unit(0.3, "cm"),
                                ticks.colour = "white",
                                frame.colour = NA)) +
  theme(legend.position      = "top",
        legend.direction     = "horizontal",
        legend.justification = "center",
        legend.box.spacing   = unit(3, "pt"),
        legend.margin        = margin(0, 0, 0, 0),
        legend.text          = element_text(size = 8, colour = "black"),
        plot.margin          = margin(2, 2, 2, 2))
 
p_c <- p_sero + theme(legend.position = "none")
 
# --- legend column, anchored to the top of its cell ------------------------

p_a_leg <- p_cov_evc +
  guides(colour   = guide_legend(order = 1, ncol = 1,
                                 override.aes = list(linetype = "solid")),
         linetype = guide_legend(order = 2, ncol = 1,
                                 override.aes = list(colour = "grey30"))) +
  theme(legend.position  = "right",
        legend.direction = "vertical",
        legend.box       = "vertical",
        legend.spacing.y = unit(4, "pt"),
        legend.margin    = margin(0, 0, 0, 0),
        legend.title     = element_text(size = 10, face = "bold"),
        legend.text      = element_text(size = 9),
        legend.key.width = unit(1.2, "lines"))

leg_grob <- cowplot::get_legend(p_a_leg)
p_leg <- wrap_elements(full = grid::grobTree(
  leg_grob, vp = grid::viewport(y = 1, just = "top", height = grid::unit(1, "npc"))
)) + theme(plot.margin = margin(0, 0, 0, 20))
 
# --- grid: A and B on top, C below, legend in its own right column ---------
# B is wrapped so patchwork does not align A's panel top with B's panel top,
# which is what left the gap above A when B's colourbar sat on top.
 
design <- "
AABL
CCCL
"
 
p_all <- p_a + wrap_elements(full = p_b) + p_c + p_leg +
  plot_layout(design = design, widths = c(1, 1, 1.05, 0.32),
              heights = c(1, 1)) +
  plot_annotation(tag_levels = list(c("A", "B", "C", ""))) &
  theme(plot.tag = element_text(face = "bold"))
 
p_all
 
# ggsave("figs/chap1/fig_cov_evc_sero.png", p_all, width = 8, height = 5, dpi = 300)
```

## Values

```{r}
# ===========================================================================
# SUMMARY VALUES FOR THE PANEL A DESCRIPTION
# ===========================================================================
 
pp  <- \(x) sprintf("%.1f", 100 * x)      # proportion as a percentage
ppd <- \(x) sprintf("%+.1f", 100 * x)     # change in percentage points
 
wk_first <- min(lines_region$week)
wk_last  <- max(lines_region$week)
 
c(as.character(wk_first), as.character(wk_last))   # the period being described
 
# ---- 1. Start, end, and change for every region and series ----------------
 
ends <- lines_region |>
  filter(week %in% c(wk_first, wk_last)) |>
  mutate(point = if_else(week == wk_first, "start", "end")) |>
  select(series, region, point, value) |>
  pivot_wider(names_from = point, values_from = value) |>
  mutate(change_pp = 100 * (end - start)) |>
  arrange(series, desc(end))
 
ends |> print(n = Inf)
 
# ---- 2. Leading and trailing region, and the spread between them ----------
 
extremes <- lines_region |>
  filter(week %in% c(wk_first, wk_last)) |>
  mutate(point = factor(if_else(week == wk_first, "Start", "End"),
                        levels = c("Start", "End"))) |>
  summarise(
    highest   = as.character(region[which.max(value)]),
    high_val  = max(value),
    lowest    = as.character(region[which.min(value)]),
    low_val   = min(value),
    spread_pp = 100 * (max(value) - min(value)),
    .by = c(series, point)
  ) |>
  arrange(series, point)
 
extremes |> print(n = Inf)
 
# did the ranking of regions change between start and end?
lines_region |>
  filter(week %in% c(wk_first, wk_last)) |>
  mutate(point = if_else(week == wk_first, "start", "end"),
         rk    = rank(-value), .by = c(series, week)) |>
  select(series, region, point, rk) |>
  pivot_wider(names_from = point, values_from = rk) |>
  filter(start != end) |>
  print(n = Inf)
 
# ---- 3. EVC with its uncertainty interval --------------------------------
 
evc_ci <- rib_region |>
  filter(week %in% c(wk_first, wk_last)) |>
  left_join(lines_region |>
              filter(series == x_labels[["EVC"]]) |>
              select(week, group, evc = value),
            by = c("week", "group")) |>
  mutate(point    = factor(if_else(week == wk_first, "Start", "End"),
                           levels = c("Start", "End")),
         width_pp = 100 * (evc_hi - evc_lo),
         lab      = sprintf("%s%% (%s to %s%%)", pp(evc), pp(evc_lo), pp(evc_hi))) |>
  select(region, point, evc, evc_lo, evc_hi, width_pp, lab) |>
  arrange(point, desc(evc))
 
evc_ci |> print(n = Inf)
 
# band width across the whole period: all regions share the same draws, so
# this is VE-parameter uncertainty, not between-region precision
rib_region |>
  summarise(width_min = 100 * min(evc_hi - evc_lo),
            width_med = 100 * median(evc_hi - evc_lo),
            width_max = 100 * max(evc_hi - evc_lo))
 
# ---- 4. Gap between MCV1 coverage and estimated immunity -----------------
# The vertical distance between the two lines is primary vaccine failure
# plus waning, expressed in percentage points.
 
gap <- lines_region |>
  filter(series %in% unname(x_labels[c("Dose 1", "EVC")])) |>
  select(week, region, series, value) |>
  pivot_wider(names_from = series, values_from = value) |>
  rename(mcv1 = !!x_labels[["Dose 1"]], evc = !!x_labels[["EVC"]]) |>
  mutate(gap_pp = 100 * (mcv1 - evc))
 
gap |>
  summarise(gap_start = gap_pp[which.min(week)],
            gap_end   = gap_pp[which.max(week)],
            gap_min   = min(gap_pp),
            gap_max   = max(gap_pp),
            .by = region) |>
  print(n = Inf)
 
# ---- 5. Largest week-on-week step, e.g. the early-2024 jump --------------
 
lines_region |>
  arrange(region, series, week) |>
  mutate(step_pp = 100 * (value - lag(value)), .by = c(region, series)) |>
  slice_max(step_pp, n = 1, by = c(region, series)) |>
  select(series, region, week, step_pp) |>
  arrange(series, desc(step_pp)) |>
  print(n = Inf)
 
# ---- 6. Sentence-ready strings -------------------------------------------
 
ends |>
  mutate(txt = sprintf("%s, %s: %s%% to %s%% (%s pp)",
                       series, region, pp(start), pp(end), ppd(end - start))) |>
  pull(txt) |>
  writeLines()
 
evc_ci |>
  filter(point == "End") |>
  mutate(txt = sprintf("%s: EVC %s", region, lab)) |>
  pull(txt) |>
  writeLines()
```

## Values sero vs cov

```{r}
# ===========================================================================
# SUMMARY VALUES FOR THE PANEL C DESCRIPTION
# ===========================================================================
 
pp <- \(x) sprintf("%.1f", 100 * x)
 
# ---- 1. Seroprevalence by region, with sample sizes -----------------------
# Same sera for MCV1 and EVC (9-59 mo); MCV2 uses 18-59 mo, so its
# seroprevalence and n differ.
 
scat |>
  select(metric, region, x_pos, n_sero, sero, sero_lo, sero_hi) |>
  mutate(lab = sprintf("%s%% (%s to %s%%), %d/%d",
                       pp(sero), pp(sero_lo), pp(sero_hi), x_pos, n_sero)) |>
  arrange(metric, desc(sero)) |>
  print(n = Inf)
 
# range across regions, per age window
scat |>
  summarise(n_regions = n(),
            n_sera    = sum(n_sero),
            sero_min  = min(sero), sero_max = max(sero),
            reg_min   = as.character(region[which.min(sero)]),
            reg_max   = as.character(region[which.max(sero)]),
            .by = metric) |>
  print(n = Inf)
 
# widest and narrowest confidence intervals, which drive how much the
# scatter can be read at all
scat |>
  mutate(ci_pp = 100 * (sero_hi - sero_lo)) |>
  slice_max(ci_pp, n = 1, by = metric) |>
  select(metric, region, n_sero, sero, ci_pp) |>
  print(n = Inf)
 
# ---- 2. The x metrics: period value and temporal range -------------------
 
scat |>
  mutate(lab = sprintf("%s%% (range %s to %s%%)", pp(x), pp(x_lo), pp(x_hi))) |>
  select(metric, region, x, x_lo, x_hi, lab) |>
  arrange(metric, desc(x)) |>
  print(n = Inf)
 
scat |>
  summarise(x_min = min(x), x_max = max(x),
            spread_pp = 100 * (max(x) - min(x)),
            reg_min = as.character(region[which.min(x)]),
            reg_max = as.character(region[which.max(x)]),
            .by = metric) |>
  print(n = Inf)
 
# EVC parameter interval, which only the EVC panel carries
scat |>
  filter(!is.na(xu_lo)) |>
  mutate(u_pp = 100 * (xu_hi - xu_lo)) |>
  select(metric, region, x, xu_lo, xu_hi, u_pp) |>
  print(n = Inf)
 
# ---- 3. The gap: how far every point sits below the 1:1 line -------------
# This is the dominant feature of panel C and the number to lead with.
 
scat |>
  mutate(gap_pp = 100 * (x - sero)) |>
  summarise(gap_med = median(gap_pp),
            gap_min = min(gap_pp), gap_max = max(gap_pp),
            n_above = sum(gap_pp < 0),          # regions above the 1:1 line
            .by = metric) |>
  print(n = Inf)
 
# per region, for the metric-specific sentences
scat |>
  mutate(gap_pp = 100 * (x - sero)) |>
  select(metric, region, x, sero, gap_pp) |>
  arrange(metric, desc(gap_pp)) |>
  print(n = Inf)
 
# does any region's seroprevalence CI reach its x value?
scat |>
  filter(sero_hi >= x) |>
  select(metric, region, x, sero, sero_hi) |>
  print(n = Inf)
 
# ---- 4. Do the two axes rank the regions the same way? -------------------
# Ranking agreement is the descriptive version of the correlation, and is
# more honest than a p-value at n = 5.
 
scat |>
  mutate(rk_x = rank(-x), rk_y = rank(-sero), .by = metric) |>
  select(metric, region, rk_x, rk_y) |>
  arrange(metric, rk_x) |>
  print(n = Inf)
 
scat |>
  summarise(rho = cor(x, sero, method = "spearman"),
            n_concordant = sum(rank(-x) == rank(-sero)),
            .by = metric) |>
  print(n = Inf)
 
# ---- 5. Sentence-ready strings -------------------------------------------
 
scat |>
  filter(metric == x_labels[["EVC"]]) |>
  mutate(txt = sprintf("%s: EVC %s%%, seroprevalence %s%% (%s to %s%%), gap %.1f pp",
                       region, pp(x), pp(sero), pp(sero_lo), pp(sero_hi),
                       100 * (x - sero))) |>
  arrange(desc(x)) |>
  pull(txt) |>
  writeLines()
```


## Sero vs cum inc

```{r}
# ===========================================================================
# 12. REGION-LEVEL CUMULATIVE INCIDENCE, BY BIN
# ===========================================================================
# Region incidence is recomputed from summed cases and population, not
# averaged across district rates.

sero_age_lo <- 9 / 12      # years; matches the EVC denominator
sero_age_hi <- 5
 
bin_keep <- as.Date(c("2024-07-01", "2024-10-01", "2025-01-01",
                      "2025-04-01", "2025-07-01", "2025-10-01"))
 
inc_src <- map_snapshot |>
  as_tibble() |>
  mutate(bin = as.Date(bin), region = assign_region(district)) |>
  filter(bin %in% bin_keep)
 
setdiff(unique(inc_src$district), levels(child$district))
 
inc_all <- inc_src |>
  summarise(cum_n = sum(cum_n), n_census = sum(n_census), .by = c(region, bin)) |>
  mutate(
    group  = as.character(region),
    inc    = cum_n / n_census * 1e5,
    inc_lo = qchisq(0.025, 2 * cum_n) / 2 / n_census * 1e5,
    inc_hi = qchisq(0.975, 2 * (cum_n + 1)) / 2 / n_census * 1e5
  )
 
inc_all |> select(group, bin, cum_n, inc) |> arrange(bin, group) |> print(n = Inf)
 
# ===========================================================================
# 13. IMMUNITY MEASURES VS CUMULATIVE INCIDENCE, REGION LEVEL
# ===========================================================================
# Rows are the pre-outbreak immunity measures, columns the outbreak bins, so
# each row shows how one measure's association with incidence develops as the
# outbreak accumulates. Every panel has five points.
 
snap_week   <- max(weeks)     # last study week; coverage and EVC snapshot
n_min_round <- 5              # minimum sera per region within the last round
common_only <- TRUE
 
hosp_col <- "hospital"        # <- set to your hospital column name
 
# ---- last collection period, defined within each hospital -----------------
 
sero_src |>                                   # verify the ordering first
  count(.data[[hosp_col]], period_collect) |>
  print(n = Inf)
 
sero_last <- sero_src |>
  filter(period_collect == max(period_collect), .by = all_of(hosp_col))
 
sero_last |>                                  # which period each hospital gave
  summarise(period = first(period_collect), n = n(), .by = all_of(hosp_col))
 
sero_last |>
  filter(age_y >= sero_age_lo, age_y < sero_age_hi) |>
  count(region) |>
  print(n = Inf)
 
sero_snap <- sero_last |>
  filter(age_y >= sero_age_lo, age_y < sero_age_hi) |>
  mutate(region = as.character(region)) |>
  sero_ci("region") |>
  filter(n_sero >= n_min_round)
 
# ---- coverage and EVC at the snapshot week --------------------------------
 
metric_lev <- c(unname(x_labels), "Seroprevalence")
 
x_all <- bind_rows(
  ser_region$cov |>
    filter(week == snap_week) |>
    transmute(group  = as.character(group),
              metric = unname(x_labels[as.character(dose)]),
              x = coverage, x_lo = NA_real_, x_hi = NA_real_),
  ser_region$evc |>
    filter(week == snap_week) |>
    transmute(group  = as.character(group),
              metric = x_labels[["EVC"]],
              x = evc, x_lo = evc_lo, x_hi = evc_hi),
  sero_snap |>
    transmute(group = as.character(group), metric = "Seroprevalence",
              x = sero, x_lo = sero_lo, x_hi = sero_hi)
)
 
if (common_only) {
  keep <- x_all |>
    summarise(k = n_distinct(metric), .by = group) |>
    filter(k == length(metric_lev))
  x_all <- x_all |> semi_join(keep, by = "group")
}
 
bin_lab <- setNames(format(bin_keep, "%b %Y"), as.character(bin_keep))
 
dat_inc <- x_all |>
  inner_join(inc_all |> select(group, bin, region, inc, inc_lo, inc_hi),
             by = "group") |>
  mutate(metric = factor(metric, levels = metric_lev),
         region = factor(as.character(region), levels = rank_order),
         bin    = factor(bin_lab[as.character(bin)], levels = unname(bin_lab)))
 
dat_inc |> summarise(n = n(), .by = c(metric, bin)) |> print(n = Inf)
 
# ---- Spearman rho ---------------------------------------------------------
# n = 5 in every panel, and the six bins are nested cumulative counts, so
# these are not six independent tests. Treat the row as a trajectory.
 
cor_inc_all <- dat_inc |>
  summarise(
    n   = sum(is.finite(x) & is.finite(inc)),
    rho = if (n[1] >= 3) cor(x, inc, method = "spearman",
                             use = "complete.obs") else NA_real_,
    p   = if (n[1] >= 3) suppressWarnings(
            cor.test(x, inc, method = "spearman", exact = FALSE)$p.value
          ) else NA_real_,
    .by = c(metric, bin)
  ) |>
  mutate(lab = if_else(is.na(rho), "rho = NA",
                       sprintf("rho = %.2f\np = %.3f", rho, p)))
 
cor_inc_all |> print(n = Inf)
 
# ---- figure ---------------------------------------------------------------
# x varies by row and y by column, which is the opposite of what facet_grid
# can free, so ggh4x is used when available to free every panel.
# Horizontal bars: 95% VE-parameter interval (EVC) or exact 95% CI
# (seroprevalence). Coverage is a registry census and carries no bar.
# Vertical bars: exact 95% Poisson CI for the incidence rate.

theil_sen <- function(x, y) {
  ok <- is.finite(x) & is.finite(y)
  x <- x[ok]; y <- y[ok]
  if (length(x) < 3) return(NULL)
  cb   <- combn(length(x), 2)
  dx   <- x[cb[2, ]] - x[cb[1, ]]
  dy   <- y[cb[2, ]] - y[cb[1, ]]
  keep <- dx != 0
  if (!any(keep)) return(NULL)
  b <- median(dy[keep] / dx[keep])
  tibble(a = median(y - b * x), b = b,
         x1 = min(x), x2 = max(x))
}
 
fit_ts <- dat_inc |>
  reframe(theil_sen(inc, x), .by = c(metric, bin)) |>
  transmute(metric, bin, inc = x1, inc_end = x2,
            x = a + b * x1, x_end = a + b * x2)
 
free_all <- requireNamespace("ggh4x", quietly = TRUE)
 
p_inc_all <- ggplot(dat_inc, aes(inc, x, colour = region)) +
  geom_segment(aes(x = inc_lo, xend = inc_hi, y = x, yend = x)) +
  geom_segment(aes(x = inc, xend = inc, y = x_lo, yend = x_hi), na.rm = TRUE) +
  geom_point(size = 1.2) +
  geom_segment(data = fit_ts, inherit.aes = FALSE,
               aes(x = inc, xend = inc_end, y = x, yend = x_end),
               linetype = "dashed", colour = "grey40", linewidth = 0.4) +
  geom_text(data = cor_inc_all, aes(x = -Inf, y = Inf, label = lab),
            inherit.aes = FALSE, colour = "black", hjust = -0.08, vjust = 1.2,
            lineheight = 0.9) +
  {if (free_all)
    ggh4x::facet_grid2(metric ~ bin, scales = "free", independent = "all", , switch = "y")
   else
    facet_grid(metric ~ bin, scales = "free", , switch = "y")} +
  scale_colour_manual(values = pal_region) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  labs(x = "Cumulative incidence per 100,000", y = NULL, colour = "Region") +
  theme_epi() +
  theme(strip.placement = "outside")
 
# p_inc_all
#  
# ggsave("figs/chap1/fig_immunity_vs_incidence.png", p_inc_all,
#        width = 9.5, height = 6.5, dpi = 300)
```


## Under5: Sero vs cum

```{r}
# ===========================================================================
# 14. UNDER-5 CUMULATIVE INCIDENCE BY REGION
# ===========================================================================
# The all-age version compares child immunity against an outbreak measured in
# the whole population. Restricting both numerator and denominator to young
# children makes the incidence measure match what the immunity measures
# describe.
 
case_age_lo <- 0     # years, inclusive
case_age_hi <- 5     # years, exclusive
 
# Setting case_age_lo <- 1 excludes infants, who are not yet MCV1-eligible and
# so sit outside the coverage and EVC denominators (9-59 months). That is the
# cleaner age match; 0-5 keeps the group epidemiologically complete.
 
dedupe_cases <- TRUE
 
# ---- duplicate records ----------------------------------------------------
 
df_ep2 |> summarise(rows = n(), unique_ids = n_distinct(id))
df_ep2 |> count(id) |> filter(n > 1) |> nrow()
 
# ---- census denominator ---------------------------------------------------
# Census age band k covers [k-1, k), so under 5 is bands 1:5, not 1:4.
 
census_u5 <- df_census |>
  filter(age >= case_age_lo + 1, age <= case_age_hi) |>
  summarise(n = sum(n), .by = district) |>
  mutate(district = merge_thuduc(clean_district(district)),
         region   = assign_region(district)) |>
  summarise(n_census = sum(n), .by = region) |>
  drop_na(region)
 
census_u5
 
# sanity: under-5s should be a few percent of the all-age total
tibble(u5 = sum(census_u5$n_census), all_age = sum(df_census$n)) |>
  mutate(pct = 100 * u5 / all_age)
 
# ---- case numerator -------------------------------------------------------
 
cases_u5 <- df_ep2 |>
  (\(d) if (dedupe_cases) distinct(d, id, .keep_all = TRUE) else d)() |>
  filter(age >= case_age_lo, age < case_age_hi) |>
  mutate(district = merge_thuduc(.data[[col_district]]),
         region   = assign_region(district),
         date     = .data[[col_date]]) |>
  drop_na(region, date)
 
cases_u5 |> count(region)
cat("under-5 cases:", nrow(cases_u5), "of", nrow(df_ep2), "rows\n")
 
# ---- quarterly cumulative incidence ---------------------------------------
# Bins built from the full date range so the under-5 and all-age versions
# share a bin grid; regions with no cases in a quarter still get a row.
 
date_rng <- range(df_ep2[[col_date]], na.rm = TRUE)
bin_grid <- seq(floor_date(date_rng[1], "quarter"),
                floor_date(date_rng[2], "quarter"), by = "quarter")
 
inc_u5 <- cases_u5 |>
  mutate(bin = floor_date(date, "quarter")) |>
  count(region, bin) |>
  complete(region = census_u5$region, bin = bin_grid, fill = list(n = 0)) |>
  arrange(region, bin) |>
  mutate(cum_n = cumsum(n), .by = region) |>
  left_join(census_u5, by = "region") |>
  mutate(
    group  = as.character(region),
    inc    = cum_n / n_census * 1e5,
    inc_lo = qchisq(0.025, 2 * cum_n) / 2 / n_census * 1e5,
    inc_hi = qchisq(0.975, 2 * (cum_n + 1)) / 2 / n_census * 1e5
  )
 
inc_u5 |> select(group, bin, cum_n, n_census, inc) |> arrange(bin) |> print(n = Inf)
 
# ===========================================================================
# 15. IMMUNITY MEASURES VS UNDER-5 INCIDENCE
# ===========================================================================
# Reuses x_all from section 13 (coverage and EVC at snap_week, seroprevalence
# from each hospital's last collection period).
 
bin_keep <- as.Date(c("2024-07-01", "2024-10-01", "2025-01-01",
                      "2025-04-01", "2025-07-01", "2025-10-01"))
bin_lab  <- setNames(format(bin_keep, "%b %Y"), as.character(bin_keep))
 
dat_u5 <- x_all |>
  inner_join(inc_u5 |> filter(bin %in% bin_keep) |>
               select(group, bin, region, inc, inc_lo, inc_hi),
             by = "group") |>
  mutate(metric = factor(metric, levels = metric_lev),
         region = factor(as.character(region), levels = rank_order),
         bin    = factor(bin_lab[as.character(bin)], levels = unname(bin_lab)))
 
dat_u5 |> summarise(n = n(), .by = c(metric, bin)) |> print(n = Inf)
 
# ---- Spearman rho ---------------------------------------------------------
# Five regions per panel, and the six bins are nested cumulative counts, so
# read each row as a trajectory rather than as six independent tests.
 
cor_u5 <- dat_u5 |>
  summarise(
    n   = sum(is.finite(x) & is.finite(inc)),
    rho = if (n[1] >= 3) cor(x, inc, method = "spearman",
                             use = "complete.obs") else NA_real_,
    p   = if (n[1] >= 3) suppressWarnings(
            cor.test(x, inc, method = "spearman", exact = FALSE)$p.value
          ) else NA_real_,
    .by = c(metric, bin)
  ) |>
  mutate(lab = if_else(is.na(rho), "rho = NA",
                       sprintf("rho = %.2f\np = %.3f", rho, p)))
 
cor_u5 |> print(n = Inf)
 
# side by side with the all-age version, if section 13 was run
if (exists("cor_inc_all")) {
  cor_inc_all |>
    select(metric, bin, rho_all = rho) |>
    left_join(cor_u5 |> select(metric, bin, rho_u5 = rho), by = c("metric", "bin")) |>
    print(n = Inf)
}
 
# ---- Theil-Sen line, consistent with a rank correlation -------------------
 
theil_sen <- function(x, y) {
  ok <- is.finite(x) & is.finite(y)
  x <- x[ok]; y <- y[ok]
  if (length(x) < 3) return(NULL)
  cb   <- combn(length(x), 2)
  dx   <- x[cb[2, ]] - x[cb[1, ]]
  dy   <- y[cb[2, ]] - y[cb[1, ]]
  keep <- dx != 0
  if (!any(keep)) return(NULL)
  b <- median(dy[keep] / dx[keep])
  tibble(a = median(y - b * x), b = b, x1 = min(x), x2 = max(x))
}
 
fit_u5 <- dat_u5 |>
  reframe(theil_sen(inc, x), .by = c(metric, bin)) |>
  transmute(metric, bin, inc = x1, inc_end = x2,
            x = a + b * x1, x_end = a + b * x2)
 
# ---- figure ---------------------------------------------------------------
# Horizontal bars: exact 95% Poisson CI for the incidence rate.
# Vertical bars: 95% VE-parameter interval (EVC) or exact 95% CI
# (seroprevalence). Coverage is a registry census and carries no bar.
# Dashed line: Theil-Sen fit, an aid to reading direction, not a model.
 
p_u5 <- ggplot(dat_u5, aes(inc, x, colour = region)) +
  geom_segment(aes(x = inc_lo, xend = inc_hi, y = x, yend = x)) +
  geom_segment(aes(x = inc, xend = inc, y = x_lo, yend = x_hi), na.rm = TRUE) +
  geom_segment(data = fit_u5, inherit.aes = FALSE,
               aes(x = inc, xend = inc_end, y = x, yend = x_end),
               linetype = "dashed", colour = "grey40", linewidth = 0.4) +
  geom_point(size = 1.2) +
  geom_text(data = cor_u5, aes(x = -Inf, y = Inf, label = lab),
            inherit.aes = FALSE, colour = "black", hjust = -0.08, vjust = 1.2,
            lineheight = 0.9) +
  facet_grid(metric ~ bin, scales = "free", switch = "y") +
  scale_colour_manual(values = pal_region) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  labs(x = "Cumulative incidence per 100,000 children under 5",
       y = NULL, colour = "Region") +
  theme_epi() +
  theme(strip.placement = "inside")
 
p_u5
 
ggsave("figs/chap1/fig_immunity_vs_incidence_u5.png", p_u5,
       width = 9.5, height = 6.5, dpi = 300)
```



## Case negative coverage

```{r}
df_case_neg <- readRDS(file.path(path2data, cases_neg_file))

df_case_neg <- df_case_neg |> 
  filter(ngay_nv >= ep2$start, ngay_nv <= ep2$end)

df_summary <- df_case_neg |>
  mutate(
    age_group = cut(
      age,
      breaks = c(-Inf, 9 / 12, 5, 15, Inf),
      labels = c("<9m", "9m-5", "5-15", "15+")
    ),
    vax = factor(tiem_chung,
                 levels = c("Không", "Có", "Không rõ"),
                 labels = c("No", "Yes", "Unknown"))
  ) |>
  count(age_group, vax) |>
  pivot_wider(names_from = vax, values_from = n, values_fill = 0) |>
  mutate(
    n_total = No + Yes + Unknown,
    # complete case: unknown excluded from numerator and denominator
    pct_cc  = 100 * Yes / (Yes + No),
    # lower bound: unknown counted as unvaccinated
    pct_lo  = 100 * Yes / n_total,
    # upper bound: unknown counted as vaccinated
    pct_hi  = 100 * (Yes + Unknown) / n_total
  ) |>
  transmute(
    age_group,
    No      = sprintf("%d", No),
    Yes     = sprintf("%d", Yes),
    Unknown = sprintf("%d", Unknown),
    n_total,
    `Complete case`  = sprintf("%.1f%%", pct_cc),
    `Unknown as no`  = sprintf("%.1f%%", pct_lo),
    `Unknown as yes` = sprintf("%.1f%%", pct_hi)
  )
 
df_summary

# write.csv(df_summary, "figs/chap1/vax_cases_2425.csv", quote = F, row.names = F)
```

## Catchment vs cases address

```{r}
# ===========================================================================
# 16. SEROSURVEY CATCHMENT VS HOSPITAL ADMISSION CATCHMENT
# ===========================================================================
# Three maps per hospital: the district composition of collected serum
# samples, and the district composition of measles admissions in each
# outbreak period, plus a direct scatter of the two shares.
 
col_date <- "ngay_nv"     # admission date column in df_case
 
# ---- 16a. Hospital name normalisation -------------------------------------
# "Nhi đồng 1/2" anchored at end of string so "Nhi Đồng Đồng Nai" and
# "Nhi đồng Tp.Cần Thơ" cannot match. Diacritic case varies, so ignore_case.
 
hosp_map <- df_case |>
  as_tibble() |>
  mutate(
    hosp = case_when(
      str_detect(ten_bv, regex("nhi\\s*đồng\\s*1\\s*$", ignore_case = TRUE)) ~ "CH1",
      str_detect(ten_bv, regex("nhi\\s*đồng\\s*2\\s*$", ignore_case = TRUE)) ~ "CH2",
      str_detect(ten_bv, regex("nhi\\s*đồng\\s*thành\\s*phố\\s*$",
                               ignore_case = TRUE))                          ~ "CCH",
      .default = NA_character_
    )
  )
 
# CHECK: which names mapped where, and the totals
hosp_map |> count(hosp, ten_bv) |> filter(!is.na(hosp)) |> print(n = Inf)
hosp_map |> count(hosp)      # expect CH1 4422, CH2 3415, CCH 1185
 
df_case_3h <- hosp_map |> filter(!is.na(hosp))
 
# ---- 16b. Percentages from all three sources ------------------------------
 
ep_win <- tibble(
  panel = c("2018-2020", "2024-2025"),
  start = c(ep1$start, ep2$start),
  end   = c(ep1$end,   ep2$end)
)
 
ep_win      # confirm the labels match the actual episode date ranges
 
adm_pct <- ep_win |>
  reframe(
    df_case_3h |>
      filter(between(.data[[col_date]], start, end)) |>
      drop_na(district) |>
      count(hosp, district),
    .by = panel
  ) |>
  mutate(pct = 100 * n / sum(n), .by = c(panel, hosp)) |>
  transmute(panel, hospital = hosp, district, pct)
 
adm_pct |> summarise(n_dist = n(), total = sum(pct), .by = c(panel, hospital))
 
# CHECK: district strings must line up across sources
setdiff(unique(adm_pct$district), unique(df_map_pct$district))
setdiff(unique(df_map_pct$district), unique(adm_pct$district))
setdiff(unique(adm_pct$district), unique(df_map_hcmc$district_blank))
 
panel_lev <- c("Serum samples", "2018-2020", "2024-2025")
 
pct_all <- bind_rows(
  df_map_pct |> as_tibble() |>
    transmute(panel = "Serum samples", hospital = as.character(hospital),
              district, pct),
  adm_pct
) |>
  mutate(panel    = factor(panel, levels = panel_lev),
         hospital = factor(hospital, levels = levels(df_map_pct$hospital)))
 
# ---- 16c. Panel A: maps ---------------------------------------------------
# log2 of each district's share against the equal-share reference
# (100 / n districts), so white marks proportional representation, red above,
# blue below. Districts below pct_min draw grey. One shared scale.
 
pct_min <- 1                                             # below this % draws grey
mid     <- 100 / n_distinct(df_map_hcmc$district_blank)  # equal-share reference
 
map_pct <- pct_all |>
  complete(nesting(panel, hospital), district = df_map_hcmc$district_blank,
           fill = list(pct = 0)) |>
  mutate(lfc = if_else(pct < pct_min, NA_real_, log2(pct / mid))) |>
  left_join(select(df_map_hcmc, district_blank),
            by = c("district" = "district_blank")) |>
  st_as_sf()
 
lo <- log2(pct_min / mid)                 # bar starts exactly at pct_min
hi <- max(map_pct$lfc, na.rm = TRUE)
 
# four breaks only: six labels collide on a bar this width
brks_pct <- c(1, 5, 20, 40)
brks_pct <- brks_pct[log2(brks_pct / mid) >= lo & log2(brks_pct / mid) <= hi]
 
p_catch <- map_pct |>
  ggplot() +
  geom_sf(aes(fill = lfc), colour = "grey60", linewidth = 0.25) +
  facet_grid(hospital ~ panel, switch = "y") +
  coord_sf(expand = FALSE) +
  scale_fill_gradientn(
    colours  = pal_gsea()(12),
    limits   = c(lo, hi),
    na.value = "grey92",
    rescaler = \(x, to = c(0, 1), from = range(x, na.rm = TRUE))
                 scales::rescale_mid(x, to, from, mid = 0),
    breaks   = log2(brks_pct / mid), labels = paste0(brks_pct, "%"),
    name     = "Share of samples or admissions",
    guide    = guide_colourbar(title.position = "top",
                               barwidth  = unit(5, "cm"),
                               barheight = unit(0.3, "cm"),
                               ticks.colour = "white",
                               frame.colour = NA)
  ) +
  theme_void(base_size = 11) +
  theme(
    strip.text.x       = element_text(colour = "grey15", size = 11,
                                      margin = margin(b = 5)),
    strip.text.y.left  = element_text(colour = "grey15", size = 11, angle = 0,
                                      margin = margin(r = 6)),
    strip.placement    = "outside",
    strip.clip         = "off",
    legend.position    = "bottom",
    legend.title       = element_text(colour = "grey15", hjust = 0.5, size = 10),
    legend.text        = element_text(colour = "grey30", size = 9),
    legend.box.spacing = unit(2, "pt"),
panel.spacing.x    = unit(0.3, "lines"),
    panel.spacing.y    = unit(0.1, "lines"),
    plot.margin        = margin(4, 4, 2, 4)
  )
 
p_catch
 
# ---- 16d. Summary values for the description ------------------------------
 
n_dist  <- n_distinct(df_map_hcmc$district_blank)
tab_pct <- map_pct |> st_drop_geometry()
 
# sample size behind each panel
bind_rows(
  df_map_pct |> as_tibble() |>
    summarise(n = sum(n), .by = hospital) |> mutate(panel = "Serum samples"),
  ep_win |>
    reframe(df_case_3h |>
              filter(between(.data[[col_date]], start, end)) |>
              drop_na(district) |>
              count(hosp, name = "n") |> rename(hospital = hosp),
            .by = panel)
) |>
  pivot_wider(names_from = panel, values_from = n) |>
  print()
 
# concentration: how few districts make up most of each catchment
tab_pct |>
  summarise(
    top1     = max(pct),
    top1_d   = district[which.max(pct)],
    top3     = sum(sort(pct, decreasing = TRUE)[1:3]),
    n_to_80  = which(cumsum(sort(pct, decreasing = TRUE)) >= 80)[1],
    n_over_1 = sum(pct >= 1),
    .by = c(hospital, panel)
  ) |>
  arrange(hospital, panel) |>
  print(n = Inf)
 
# agreement: Spearman rho and total variation distance (half the sum of
# absolute share differences, so 0 = identical, 100 = disjoint)
wide <- tab_pct |>
  select(hospital, panel, district, pct) |>
  pivot_wider(names_from = panel, values_from = pct, values_fill = 0)
 
agree <- wide |>
  summarise(
    n       = n(),
    rho_ep1 = cor(`Serum samples`, `2018-2020`, method = "spearman"),
    p_ep1   = suppressWarnings(
                cor.test(`Serum samples`, `2018-2020`,
                         method = "spearman", exact = FALSE)$p.value),
    rho_ep2 = cor(`Serum samples`, `2024-2025`, method = "spearman"),
    p_ep2   = suppressWarnings(
                cor.test(`Serum samples`, `2024-2025`,
                         method = "spearman", exact = FALSE)$p.value),
    tvd_ep1 = sum(abs(`Serum samples` - `2018-2020`)) / 2,
    tvd_ep2 = sum(abs(`Serum samples` - `2024-2025`)) / 2,
    .by = hospital
  ) |>
  mutate(across(starts_with("rho"), \(x) round(x, 2)),
         across(starts_with("tvd"), \(x) round(x, 1)),
         across(starts_with("p_"),  \(x) format.pval(x, digits = 2, eps = 0.001)))
 
agree |> print()
 
# largest district-level discrepancies, which drive the TVD
wide |>
  mutate(d_ep1 = `Serum samples` - `2018-2020`,
         d_ep2 = `Serum samples` - `2024-2025`) |>
  slice_max(abs(d_ep2), n = 3, by = hospital) |>
  select(hospital, district, `Serum samples`, `2024-2025`, d_ep2) |>
  arrange(hospital, desc(abs(d_ep2))) |>
  print(n = Inf)
 
# districts each hospital effectively does not reach
tibble(equal_share = mid, n_districts = n_dist) |> print()
 
tab_pct |>
  summarise(n_below_1pct = sum(pct < 1), n_zero = sum(pct == 0),
            .by = c(hospital, panel)) |>
  arrange(hospital, panel) |>
  print(n = Inf)
 
# pairwise overlap between the three sample catchments
smp <- wide |> select(hospital, district, pct = `Serum samples`)
 
crossing(a = levels(smp$hospital), b = levels(smp$hospital)) |>
  filter(a < b) |>
  mutate(overlap = map2_dbl(a, b, \(x, y) {
    pa <- smp |> filter(hospital == x) |> arrange(district) |> pull(pct)
    pb <- smp |> filter(hospital == y) |> arrange(district) |> pull(pct)
    sum(pmin(pa, pb))          # shared share, 0 = disjoint, 100 = identical
  })) |>
  print()
 
# ---- 16e. Panel B: sample share vs admission share ------------------------
# aspect.ratio = 1 makes the panels square while letting them size to the
# available height, so B fills its cell instead of leaving side gaps.
 
p_agree <- ggplot(sc_dat, aes(pct_adm, pct_smp, colour = region)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", colour = "grey50") +
  geom_point(size = 1.1, alpha = 0.85) +
  geom_text(data = lab_pos, aes(x = pct_adm, y = pct_smp, label = lab),
            inherit.aes = FALSE, colour = "black", size = 2.2,
            hjust = 0, vjust = 1, lineheight = 0.9) +
  facet_grid(hospital ~ panel) +
  scale_colour_manual(values = pal_region) +
  {if (use_log) {
     list(scale_x_log10(labels = \(x) paste0(x, "%"),
                        breaks = c(0.1, 1, 10)),
          scale_y_log10(labels = \(x) paste0(x, "%"),
                        breaks = c(0.1, 1, 10)))
   } else {
     list(scale_x_continuous(labels = \(x) paste0(x, "%")),
          scale_y_continuous(labels = \(x) paste0(x, "%")))
   }} +
  coord_cartesian(xlim = ax_lim, ylim = ax_lim) +
  labs(x = "Share of admissions", y = "Share of serum samples",
       colour = "Region") +
  theme_epi() +
  theme(aspect.ratio  = 1,
        text          = element_text(size = 8),
        axis.title    = element_text(size = 8, face = "bold"),
        axis.text     = element_text(size = 6.5),
        strip.text    = element_text(size = 8),
        panel.spacing = unit(0.25, "lines"))
 
# ---- 16f. Combined figure at 8 x 4.5 in -----------------------------------
# All sizes below are chosen for that canvas. Both panels put their legend
# in a bottom band of similar height so the panel regions line up.
 
p_catch_c <- p_catch +
  guides(fill = guide_colourbar(title.position = "top",
                                barwidth  = unit(3.2, "cm"),
                                barheight = unit(0.22, "cm"),
                                ticks.colour = "white",
                                frame.colour = NA)) +
  theme(legend.position    = "bottom",
        legend.title       = element_text(size = 7.5, hjust = 0.5),
        legend.text        = element_text(size = 6.5),
        legend.box.spacing = unit(0, "pt"),
        legend.margin      = margin(0, 0, 0, 0),
        strip.text.x       = element_text(size = 8, margin = margin(b = 6)),
        strip.text.y.left  = element_text(size = 8, margin = margin(r = 3), face = "bold"),
        panel.spacing.x    = unit(0.9, "lines"),
        panel.spacing.y    = unit(0, "lines"),
        plot.margin        = margin(2, 4, 0, 2))
 
p_agree_c <- p_agree +
  guides(colour = guide_legend(nrow = 2, byrow = TRUE,
                               override.aes = list(size = 2))) +
  theme(strip.text.y       = element_blank(),
        legend.position    = "bottom",
        legend.direction   = "horizontal",
        legend.title       = element_text(size = 7.5, face = "bold"),
        legend.text        = element_text(size = 6.5),
        legend.key.size    = unit(0.55, "lines"),
        legend.box.spacing = unit(0, "pt"),
        legend.margin      = margin(0, 0, 0, 0),
        plot.margin        = margin(2, 2, 0, 4))
 
p_comb <- (p_catch_c | p_agree_c) +
  plot_layout(widths = c(1.2, 1)) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag          = element_text(face = "bold", size = 10),
        plot.tag.position = c(0.01, 0.99),
        plot.margin       = margin(t = 5, r = 4, b = 5, l = 2))
 
p_comb
 
ggsave("figs/chap1/fig_catchment_combined.png", p_comb,
       width = 6, height = 3.8, dpi = 300, bg = "white")
```

```{r}
tab_pct <- map_pct |> st_drop_geometry()
tab_pct |>
  summarise(
    top1     = max(pct),
    top1_d   = district[which.max(pct)],
    top3     = sum(sort(pct, decreasing = TRUE)[1:3]),
    n_to_80  = which(cumsum(sort(pct, decreasing = TRUE)) >= 80)[1],
    n_over_1 = sum(pct >= 1),
    .by = c(hospital, panel)
  ) |>
  arrange(hospital, panel) |>
  print(n = Inf)

agree <- wide |>
  summarise(
    n       = n(),
    rho_ep1 = cor(`Serum samples`, `2018-2020`, method = "spearman"),
    p_ep1   = suppressWarnings(
                cor.test(`Serum samples`, `2018-2020`,
                         method = "spearman", exact = FALSE)$p.value),
    rho_ep2 = cor(`Serum samples`, `2024-2025`, method = "spearman"),
    p_ep2   = suppressWarnings(
                cor.test(`Serum samples`, `2024-2025`,
                         method = "spearman", exact = FALSE)$p.value),
    tvd_ep1 = sum(abs(`Serum samples` - `2018-2020`)) / 2,
    tvd_ep2 = sum(abs(`Serum samples` - `2024-2025`)) / 2,
    .by = hospital
  ) |>
  mutate(across(starts_with("rho"), \(x) round(x, 2)),
         across(starts_with("tvd"), \(x) round(x, 1)),
         across(starts_with("p_"),  \(x) format.pval(x, digits = 2, eps = 0.001)))
 
agree |> print()
```


## Sero of infants

```{r}
df_sero_infant |> 
  # 1. Convert age inline
  mutate(age_month = age * 12) |> 
  ggplot(aes(x = age_month, y = titre)) +
  geom_hline(yintercept = 200, linetype = "dashed") +
  # Added alpha to prevent overplotting with your raw points
  geom_point(alpha = 0.4) + 
  # 2. Fit and plot the GLM in one step
  geom_smooth(
    method = "glm",
    method.args = list(family = Gamma(link = "log")),
    level = 0.95,
    colour = "blue" 
  ) +
  # Vaccination milestones
  geom_vline(xintercept = 9, linetype = "dashed", colour = "#D55E00", linewidth = 0.7) +
  geom_vline(xintercept = 12, linetype = "dashed", colour = "#0072B2", linewidth = 0.7) +
  scale_y_continuous(breaks = c(200, 500, 1000, 1500, 2000)) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  labs(
    x = "Age (months)",
    y = "Titer (mIU/mL)"
  ) +
  theme_classic()

# ggsave("figs/infant.png", width = 5, height = 2, dpi = 300)
```

```{r}
# saveRDS(df_ep2, "C:/Users/ongph/github/data/measles/cleaned/measles_hcmc_24-25.rds")
```


