---
title: "Result 1: Serosurveillance"
format: html
editor_options: 
  chunk_output_type: console
execute: 
  eval: false
---

## Constant

Path to data:

```{r}
path2data <- paste0("C:/Users/ongph/github/data/dphil")
```

Files names:

```{r}
# Case notification
cases_file <- "measles_hcmc_18-25.rds"
cases_neg_file <- "measles_hcmc_18-25_neg.rds"

# Serology
sero_file <- "measles_elisa_hcmc_children.rds"

# Vaccination coverage
## HCMC
vax_hcmc_file <- "vaxreg_hcmc_measles.rds"
## Vietnam
vax_vn_file <- "vaxvn_2023.xlsx"

# Population metadata
map_hcmc_file <- "gadm41_hcmc_district.rds"
map_vn_file <- "gadm41_vn.rds"
census_file <- "census2019_hcm.rds"
```

## Packages

Required packages:

```{r, message=FALSE}
required <- c("openxlsx", "janitor", "tidyverse", "scales", "mgcv", "stringi", "sf", "binom", "ggsci", "ggspatial", "patchwork", "zoo", "slider", "cowplot", "ggh4x", "gtable", "grid")

# Installing those that are not installed yet:
to_install <- required[! required %in% installed.packages()[, "Package"]]
if (length(to_install)) install.packages(to_install)

# Loading the packages for interactive use:
invisible(sapply(required, library, character.only = TRUE))

source("code/prop_query.R")
```

## Functions

Clean Vietnamese district and province name

```{r}
clean_district <- function(x) {
  x |>
    stri_trans_general("Latin-ASCII") |> 
    str_to_lower() |>                      
    str_squish() |>                        
    str_replace_all("đ", "d") |>
    str_remove("^(quan |huyen |thanh pho |tp )")
}

clean_province <- function(x) {
  x |>
    stri_trans_general("Latin-ASCII") |> 
    str_to_lower() |>
    str_squish() |>
    str_replace_all("đ", "d")
}

# Merge D2 and Đ into Thu Duc
thuduc_parts  <- c("2", "9", "thu duc")
merge_thuduc <- function(x) if_else(x %in% thuduc_parts, "thu duc", x)
```

Correct seroprevalence by sensitivity and specificity

```{r}
clamp01 <- function(x) pmin(pmax(x, 0), 1)

# Rogan-Gladen correction for test sensitivity and specificity
prev_correct <- function(x) clamp01((x + spec - 1) / (sens + spec - 1))

# Wilson score interval with continuity correction, same as prop.test()
prop_ci <- function(x, n) {
  map2(x, n, \(a, b) as_tibble_row(set_names(prop.test(a, b)$conf.int,
                                             c("ci_lwr", "ci_upr")))) |>
    list_rbind()
}

# df.residual comes from the effective degrees of freedom, so penalised
# smooths are counted correctly
predict_ci <- function(model, data, var, ci = 0.95) {
  tq   <- qt(1 - (1 - ci) / 2, model$df.residual)
  rng  <- range(data[[var]], na.rm = TRUE)
  newd <- tibble(!!var := seq(rng[1], rng[2], length.out = n_grid))
  pr   <- predict(model, newdata = newd, se.fit = TRUE, type = "link")
  inv  <- model$family$linkinv

  newd |>
    mutate(fit_true = prev_correct(inv(pr$fit)),
           lwr_true = prev_correct(inv(pr$fit - tq * pr$se.fit)),
           upr_true = prev_correct(inv(pr$fit + tq * pr$se.fit)))
}
```

Predict with confidence interval

```{r}
sero_ci <- function(df, key) {
  df |>
    summarise(n_sero = n(), x_pos = sum(pos), .by = all_of(key)) |>
    mutate(
      sero    = x_pos / n_sero,
      sero_lo = if_else(x_pos == 0, 0, qbeta(0.025, x_pos, n_sero - x_pos + 1)),
      sero_hi = if_else(x_pos == n_sero, 1, qbeta(0.975, x_pos + 1, n_sero - x_pos))
    ) |>
    rename(group = all_of(key))
}

sero_for <- function(lo, hi) {
  d <- sero_src |> filter(age_y >= lo, age_y < hi)
  bind_rows(
    sero_ci(d, "district") |>
      mutate(group = as.character(group), level = "District"),
    sero_ci(mutate(d, region = as.character(region)), "region") |>
      mutate(level = "Region")
  )
}
```

Turn districts to regions:

```{r}
district_region <- tribble(
  ~district,     ~region,
 
  "binh chanh",  "West",
  "binh tan",    "West",
 
  "12",          "North",
  "hoc mon",     "North",
  "cu chi",      "North",
 
  "thu duc",     "East",
  "2",           "East",
  "9",           "East",
 
  "7",           "South",
  "nha be",      "South",
  "can gio",     "South",
 
  "1",           "Central",
  "3",           "Central",
  "4",           "Central",
  "5",           "Central",
  "6",           "Central",
  "8",           "Central",
  "10",          "Central",
  "11",          "Central",
  "binh thanh",  "Central",
  "go vap",      "Central",
  "phu nhuan",   "Central",
  "tan binh",    "Central",
  "tan phu",     "Central"
)

lv_region <- c("West", "North", "East", "South", "Central")
```

## Data

Map:

```{r}
df_map_hcmc <- file.path(path2data, map_hcmc_file) |> 
  readRDS() |> 
  mutate(district_blank = clean_district(district))

df_map_vn <- file.path(path2data, map_vn_file) |> 
  readRDS() |> 
  mutate(province_blank = clean_province(province))
```

Census:

```{r}
df_census <- file.path(path2data, census_file) |> 
  readRDS()

df_census_age <- df_census |> 
  group_by(age) |>
  summarise(n_census = sum(n), .groups = "drop")

df_census_district <- df_census |> 
  group_by(district) |>
  summarise(n_census = sum(n), .groups = "drop") |> 
  mutate(
    district = clean_district(district)
  )
```

Case report data:

```{r}
df_case <- readRDS(file.path(path2data, cases_file))
```

Serology:

```{r}
df_sero <- file.path(path2data, sero_file) |> 
  readRDS() |> 
  mutate(district = clean_district(district))
```

Vaccination coverage

```{r}
df_vax_hcmc <- file.path(path2data, vax_hcmc_file) |> 
  readRDS() |> 
  mutate(district_blank = clean_district(district))

df_vax_vn <- file.path(path2data, vax_vn_file) |> 
  read.xlsx() |> 
  mutate(province_blank = clean_province(province))
```

## Analysis

## Case data

### Proportion by age and cohort

```{r}
#| fig-width: 8
#| fig-height: 5

week_start   <- 1   # 1 = ISO weeks (Monday), 7 = MMWR (Sunday)
ob_threshold <- 5   # admissions per week
ob_min_run   <- 3   # consecutive weeks at or above the threshold
ob_max_gap   <- 8   # weeks below threshold tolerated within one period
ob_min_cases <- 50  # smaller episodes treated as sporadic

shade_win <- tibble(
  xmin = dmy(c("01-12-2018", "01-09-2024")),
  xmax = dmy(c("31-05-2019", "30-04-2025"))
)
shade_label <- "Outbreak response\nimmunisation"
shade_fill  <- "grey70"
shade_alpha <- 0.35

age_cols <- c(
  "<1" = "#08519C", "1-5" = "#56B4E9", "5-10" = "#FBD87F",
  "10-15" = "#B87400", ">15" = "#009E73"
)

cohort_cols <- c(
  "<2009" = "#999999", "2009-2013" = "#CC79A7", "2014-2018" = "#F0E442",
  "2019-2023" = "#0072B2", "2024" = "#D55E00", "2025" = "#000000"
)

df_week <- df_case |>
  mutate(
    week = floor_date(ngay_nv, "week", week_start = week_start),
    age_group = factor(case_when(
      age <  1  ~ "<1",
      age <  5  ~ "1-5",
      age < 10  ~ "5-10",
      age <= 15 ~ "10-15",
      age >  15 ~ ">15"
    ), levels = names(age_cols)),
    birth_year = year(ngay_sinh),
    cohort = factor(case_when(
      birth_year <  2009 ~ "<2009",
      birth_year <  2014 ~ "2009-2013",
      birth_year <  2019 ~ "2014-2018",
      birth_year <  2024 ~ "2019-2023",
      birth_year == 2024 ~ "2024",
      birth_year == 2025 ~ "2025"
    ), levels = names(cohort_cols))
  ) |>
  drop_na(week)

# Quiet weeks must appear as zero, or a gap reads as continuous transmission
weekly_counts <- function(data) {
  data |>
    count(week, name = "cases") |>
    complete(week = seq(min(week), max(week), by = "week"),
             fill = list(cases = 0L))
}

find_outbreaks <- function(data, threshold = ob_threshold,
                           min_run = ob_min_run, max_gap = ob_max_gap) {
  wc  <- weekly_counts(data)
  r   <- rle(wc$cases >= threshold)
  end <- cumsum(r$lengths)
  hit <- which(r$values & r$lengths >= min_run)

  if (length(hit) == 0) {
    return(tibble(episode = integer(), start = wc$week[0], end = wc$week[0],
                  weeks = integer(), cases = integer()))
  }

  s <- (end - r$lengths + 1L)[hit]
  e <- end[hit]

  tibble(s, e, grp = cumsum(c(1L, (s[-1] - head(e, -1) - 1L) >= max_gap))) |>
    summarise(i1 = min(s), i2 = max(e), .by = grp) |>
    transmute(
      episode = row_number(),
      start   = wc$week[i1],
      end     = wc$week[i2],
      weeks   = i2 - i1 + 1L,
      cases   = map2_int(i1, i2, \(a, b) sum(wc$cases[a:b])),
      peak_week  = map2_vec(i1, i2, \(a, b) wc$week[a:b][which.max(wc$cases[a:b])]),
      peak_cases = map2_int(i1, i2, \(a, b) max(wc$cases[a:b]))
    )
}

outbreak_sensitivity <- function(data, thresholds = c(3, 5, 10, 20, 30)) {
  map(thresholds, \(x) find_outbreaks(data, threshold = x) |>
        mutate(threshold = x, .before = 1)) |>
    list_rbind()
}

episodes_all <- find_outbreaks(df_week)
episodes_all

outbreak_sensitivity(df_week)

episodes <- episodes_all |>
  filter(cases >= ob_min_cases) |>
  mutate(episode = row_number())
episodes

if (nrow(episodes) != 2) {
  stop("Expected 2 outbreak periods, found ", nrow(episodes),
       ". Adjust ob_threshold, ob_max_gap or ob_min_cases.")
}

plot_start <- episodes$start[1]
gap_start  <- episodes$end[1] + 7
gap_end    <- episodes$start[2]
plot_end   <- episodes$end[2]

df_plot <- df_week |>
  filter(between(week, plot_start, plot_end),
         week < gap_start | week >= gap_end) |>
  mutate(period = factor(if_else(week < gap_start, "p1", "p2"), c("p1", "p2")))

df_week |>
  summarise(
    total     = n(),
    plotted   = sum(between(week, plot_start, plot_end) &
                      (week < gap_start | week >= gap_end)),
    before_p1 = sum(week < plot_start),
    in_gap    = sum(week >= gap_start & week < gap_end),
    after_p2  = sum(week > plot_end),
    no_age    = sum(is.na(age_group)),
    no_cohort = sum(is.na(cohort))
  )

# 4 days = half a bar width, so the outermost bars are not cut in half
period_range <- tibble(
  period = factor(c("p1", "p2"), c("p1", "p2")),
  start  = c(plot_start, gap_end),
  end    = c(gap_start - 7, plot_end)
) |>
  mutate(lo = start - 4, hi = end + 4)

anchors_f <- period_range |>
  pivot_longer(c(lo, hi), values_to = "week")

shade_all <- shade_win |>
  mutate(period = factor(if_else(xmin < gap_start, "p1", "p2"), c("p1", "p2")),
         label = shade_label) |>
  left_join(period_range, by = "period")

vline_f <- shade_all |>
  pivot_longer(c(xmin, xmax), values_to = "x") |>
  filter(x > lo, x < hi)

# Unclipped, shading past the panel limits would widen the scale and break
# the proportional panel widths
shade_f <- shade_all |>
  mutate(xmin = pmax(xmin, lo), xmax = pmin(xmax, hi)) |>
  filter(xmax > xmin)

draw_key_shade <- function(data, params, size) {
  grid::grobTree(
    grid::rectGrob(gp = grid::gpar(fill = alpha(shade_fill, shade_alpha), col = NA)),
    grid::segmentsGrob(x0 = c(0.05, 0.95), x1 = c(0.05, 0.95), y0 = 0, y1 = 1,
                       gp = grid::gpar(col = "black", lty = "dashed"))
  )
}

anchors <- geom_blank(data = anchors_f, aes(x = week), inherit.aes = FALSE)

# Only panel A maps fill, so only panel A produces a key
rect_keyed <- geom_rect(
  data = shade_f,
  aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = label),
  inherit.aes = FALSE, alpha = shade_alpha, key_glyph = draw_key_shade
)

rect_plain <- geom_rect(
  data = shade_f, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
  inherit.aes = FALSE, fill = shade_fill, alpha = shade_alpha
)

# space = "free_x" needs facet_grid, it is not available in facet_wrap
layers_fg <- list(
  geom_vline(data = vline_f, aes(xintercept = x), inherit.aes = FALSE,
             linetype = "dashed"),
  facet_grid(cols = vars(period), scales = "free_x", space = "free_x")
)

axis_breaks <- sort(unique(c(period_range$start, period_range$end,
                            shade_win$xmin, shade_win$xmax)))

x_scale <- scale_x_date(
  breaks = \(lim) axis_breaks[axis_breaks >= lim[1] & axis_breaks <= lim[2]],
  labels = \(x) format(round_date(x, "month"), "%b\n%Y"),
  minor_breaks = NULL,
  expand = expansion(mult = 0.01)
)

base_theme <- theme_minimal(base_size = 12) +
  theme(
    panel.spacing.x    = unit(1.5, "lines"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.y = element_line(colour = "grey90", linewidth = 0.4),
    axis.line.x  = element_line(colour = "black", linewidth = 0.6),
    axis.ticks.x = element_line(colour = "black"),
    axis.text    = element_text(colour = "black"),
    axis.title   = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    strip.text   = element_blank(),
    plot.tag     = element_text(face = "bold", size = 13),
    plot.tag.position = "topleft",
    plot.margin  = margin(t = 5, r = 15, b = 5, l = 10)
  )

drop_x <- theme(axis.title.x = element_blank(),
                axis.text.x  = element_blank(),
                axis.ticks.x = element_blank())

prep_stack <- function(d, var) {
  d |>
    drop_na({{ var }}) |>
    count(period, week, grp = {{ var }}) |>
    mutate(prop = n / sum(n), .by = week)
}

stacked_panel <- function(d, cols, legend_title, reverse_stack, tag) {
  ggplot(d, aes(x = week, y = prop)) +
    rect_plain + anchors +
    geom_col(aes(fill = grp), width = 7, colour = "white", linewidth = 0.2,
             position = position_stack(reverse = reverse_stack)) +
    layers_fg +
    scale_fill_manual(values = cols) +
    scale_y_continuous(labels = label_percent(), limits = c(0, 1),
                       breaks = seq(0, 1, by = 0.25), expand = c(0, 0)) +
    x_scale +
    labs(y = "Proportion", fill = legend_title, tag = tag) +
    base_theme
}

pA <- ggplot(count(df_plot, period, week), aes(x = week, y = n)) +
  rect_keyed + anchors +
  geom_col(width = 7, fill = "grey35") +
  layers_fg +
  scale_fill_manual(
    values = setNames(shade_fill, shade_label), name = NULL,
    guide = guide_legend(keywidth = unit(20, "pt"), keyheight = unit(24, "pt"))
  ) +
  scale_y_continuous(labels = label_comma(),
                     expand = expansion(mult = c(0, 0.05))) +
  x_scale +
  labs(y = "Case counts", tag = "A") +
  base_theme + drop_x

pB <- stacked_panel(prep_stack(df_plot, age_group), age_cols, "Age group",
                    reverse_stack = TRUE, tag = "B") +
  guides(fill = guide_legend(reverse = TRUE)) +
  drop_x

pC <- stacked_panel(prep_stack(df_plot, cohort), cohort_cols, "Birth cohort",
                    reverse_stack = FALSE, tag = "C") +
  labs(x = "Admission week")

fig <- pA / pB / pC +
  plot_layout(heights = c(0.7, 1, 1), guides = "collect") &
  theme(legend.position      = "right",
        legend.justification = "top",
        legend.box.just      = "left",
        legend.spacing.y     = unit(12, "pt"))

fig

# ggsave("figs/chap1/case_age_cohort.png", fig,
#        width = 8, height = 5, dpi = 300, bg = "white")
```

### 2024-2025

```{r}
ma_days       <- 7
snapshot_unit <- "quarter"
age_levels    <- names(age_cols)
ori_label     <- "Outbreak response immunisation"

district_pals <- c(
  "West" = "#CC6677",
  "North" = "#88CCEE",
  "East" = "#DDCC77",
  "South" = "#332288",
  "Central" = "#117733"
)

df_ep2 <- df_week |>
  filter(between(week, episodes$start[2], episodes$end[2])) |>
  mutate(date = ngay_nv)

census_dist <- df_census_district |>
  mutate(district = merge_thuduc(district)) |>
  summarise(n_census = sum(n_census), .by = district)

map_dist <- df_map_hcmc |>
  mutate(district = merge_thuduc(district_blank)) |>
  select(district)

date_grid <- seq(min(df_ep2$date), max(df_ep2$date), by = "day")

ori_win <- shade_win |>
  filter(xmax >= min(date_grid), xmin <= max(date_grid)) |>
  mutate(xmin = pmax(xmin, min(date_grid)), xmax = pmin(xmax, max(date_grid)))

ori_bounds <- c(shade_win$xmin, shade_win$xmax)
ori_lines  <- ori_bounds[ori_bounds > min(date_grid) & ori_bounds < max(date_grid)]

# Drawn before the lines so the shading sits behind them
shade_layer <- geom_rect(
  data = ori_win, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
  inherit.aes = FALSE, fill = shade_fill, alpha = shade_alpha
)

shade_layer_keyed <- geom_rect(
  data = mutate(ori_win, label = ori_label),
  aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = label),
  inherit.aes = FALSE, alpha = shade_alpha, key_glyph = draw_key_shade
)

vline_layer <- geom_vline(xintercept = ori_lines, linetype = "dashed")

axis_breaks_ep2 <- sort(unique(c(range(date_grid), ori_lines)))

x_scale_ep2 <- scale_x_date(
  breaks = \(lim) axis_breaks_ep2[axis_breaks_ep2 >= lim[1] &
                                    axis_breaks_ep2 <= lim[2]],
  labels = \(x) format(round_date(x, "month"), "%b\n%Y"),
  minor_breaks = NULL,
  expand = expansion(mult = c(0.02, 0.05))
)

# Days with no cases anywhere are absent from count(); without completing them
# the cumulative sums and moving averages skip those days entirely
daily_dist <- df_ep2 |>
  drop_na(district) |>
  count(district, date) |>
  complete(district = census_dist$district, date = date_grid, fill = list(n = 0))

panel_theme <- theme_classic() +
  theme(legend.position = "bottom",
        plot.tag = element_text(face = "bold", size = 14))

snapshot_dist <- daily_dist |>
  mutate(bin = floor_date(date, snapshot_unit)) |>
  summarise(n = sum(n), .by = c(district, bin)) |>
  arrange(district, bin) |>
  mutate(cum_n = cumsum(n), .by = district) |>
  left_join(census_dist, by = "district") |>
  mutate(
    cum_inc = na_if(1e5 * cum_n / n_census, 0),
    bin_end = pmin(ceiling_date(bin, snapshot_unit) - 1, max(date_grid)),
    yr = factor(year(bin_end)),
    # Levels from month.abb, not order of appearance: Jun occurs in both years
    mo = factor(format(bin_end, "%b"), levels = month.abb)
  )

map_snapshot <- map_dist |>
  left_join(snapshot_dist, by = "district")

p_arrow <- ggplot() +
  annotation_north_arrow(
    location = "tr", which_north = "grid",
    height = unit(7, "mm"), width = unit(6, "mm"),
    style = north_arrow_fancy_orienteering(text_size = 6)
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
  theme_void() +
  theme(plot.margin = margin(4, 4, 4, 4))

pA <- ggplot(map_snapshot) +
  geom_sf(aes(fill = cum_inc), colour = "grey40", linewidth = 0.15) +
  scale_fill_gsea(
    rescaler = \(x, to = c(0, 1), from = NULL) rescale(sqrt(x)),
    na.value = "white",
    guide = guide_colourbar(
      title.position = "top", title.hjust = 0.5,
      barwidth = unit(55, "mm"), barheight = unit(3, "mm"),
      frame.colour = "black", frame.linewidth = 0.3, ticks.colour = "black"
    )
  ) +
  # Outer strip spans the panels sharing a year, inner strip gives the month
  facet_nested_wrap(
    vars(yr, mo), nrow = 1,
    nest_line = element_line(colour = "grey30", linewidth = 0.4),
    strip = strip_nested(
      background_x = list(
        element_blank(),
        element_rect(fill = "grey93", colour = "grey40", linewidth = 0.3)
      ),
      text_x = list(
        element_text(hjust = 0, face = "bold", colour = "black",
                     margin = margin(t = 2, b = 3)),
        element_text(hjust = 0.5, colour = "black",
                     margin = margin(t = 2, b = 3))
      ),
      by_layer_x = TRUE
    )
  ) +
  labs(fill = "Cumulative incidence rate (per 100,000 population)", tag = "A") +
  theme_bw(base_size = 11) +
  theme(
    legend.position = "bottom",
    legend.justification = "left",
    legend.margin = margin(t = 4, b = 0),
    legend.title = element_text(hjust = 0),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.ticks.length = unit(0, "pt"),
    panel.grid = element_blank(),
    panel.border = element_rect(colour = "grey40", linewidth = 0.3, fill = NA),
    panel.spacing.x = unit(3, "pt"),
    plot.tag = element_text(face = "bold", size = 14),
    plot.margin = margin(2, 2, 2, 2)
  ) +
  inset_element(p_arrow, left = 0.88, bottom = 0.06, right = 0.99, top = 0.30,
                align_to = "full", clip = FALSE)

# pA
```

```{r}
# rate = cases on day t / (population - cumulative cases up to day t-1).
# A daily hazard among those not yet recorded as cases, not a cumulative
# attack rate; name it accordingly in the caption.
add_attack_rate <- function(d, group, n_days = ma_days) {
  d |>
    arrange({{ group }}, date) |>
    mutate(
      at_risk = n_census - lag(cumsum(n), default = 0),
      rate    = n / at_risk,
      rate_ma = slide_mean(rate, before = n_days - 1, complete = TRUE),
      .by = {{ group }}
    )
}

rate_dist <- daily_dist |>
  left_join(census_dist, by = "district") |>
  add_attack_rate(district) |>
  left_join(district_region, by = "district") |>
  mutate(region = factor(region, levels = lv_region))

# CHECK: districts with cases but no census population drop out of the panel.
# Should be empty. Districts 2 or 9 here mean merge_thuduc() was not applied.
rate_dist |> filter(is.na(n_census)) |> distinct(district)

# log() of a zero moving average returns -Inf, which ggplot drops with a
# warning. Count them so the gaps in the lines are expected.
cat("District days with a zero moving average:",
    sum(rate_dist$rate_ma == 0, na.rm = TRUE), "\n")

# Layer order sets which line is drawn on top: faded regions first, then West,
# then East last so no other line covers it
pB <- rate_dist |>
  ggplot(aes(x = date, y = log(rate_ma * 1e5), group = district,
             colour = region)) +
  shade_layer +
  geom_line(data = \(d) filter(d, !region %in% c("West", "East")), alpha = 0.3) +
  geom_line(data = \(d) filter(d, region == "West")) +
  geom_line(data = \(d) filter(d, region == "East")) +
  vline_layer +
  scale_colour_manual(values = district_pals) +
  # legend keys at full colour, not the faded version
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  x_scale_ep2 +
  labs(x = NULL, y = "Log attack rate", colour = "Region", tag = "B") +
  panel_theme

# pB
```

```{r}
# Same case_when as df_week, so denominator bands match numerator bands. A
# parallel cut() with right = TRUE would shift every boundary by one year.
# Census codes age 1 as under 1, so census age n is completed age n - 1.
age_cols <- set_names(viridisLite::plasma(5, begin = 0, end = 0.62),
                      c("<1", "1-5", "5-10", "10-15", ">15"))

census_age <- df_census |>
  mutate(
    age_years = age - 1,
    age_group = factor(case_when(
      age_years <  1  ~ "<1",
      age_years <  5  ~ "1-5",
      age_years < 10  ~ "5-10",
      age_years <= 15 ~ "10-15",
      age_years >  15 ~ ">15"
    ), levels = age_levels)
  ) |>
  summarise(n_census = sum(n), .by = age_group)

census_age

rate_age <- df_ep2 |>
  drop_na(age_group) |>
  count(age_group, date) |>
  complete(age_group, date = date_grid, fill = list(n = 0)) |>
  left_join(census_age, by = "age_group") |>
  add_attack_rate(age_group)

# fct_rev draws <1 last, so the darkest line sits on top
pC <- rate_age |>
  mutate(age_group = fct_rev(age_group)) |>
  ggplot(aes(x = date, y = log(rate_ma * 1e5), colour = age_group)) +
  shade_layer_keyed +
  geom_line() +
  vline_layer +
  scale_colour_manual(values = age_cols) +
  scale_fill_manual(
    values = setNames(shade_fill, ori_label), name = NULL,
    guide = guide_legend(order = 2, keywidth = unit(16, "pt"),
                         keyheight = unit(16, "pt"))
  ) +
  x_scale_ep2 +
  guides(colour = guide_legend(reverse = TRUE, order = 1)) +
  labs(x = "Date of admission", y = "Log attack rate", colour = "Age (years)",
       tag = "C") +
  panel_theme +
  theme(legend.box = "vertical")

# pC
```

```{r}
#| fig-width: 5
#| fig-height: 7.5

fig_ep2 <- plot_grid(pA, pB, pC, ncol = 1, rel_heights = c(1, 1, 1.3))

fig_ep2

ggsave("figs/chap1/ep2_spatial_attack.png", fig_ep2,
       width = 5, height = 7.5, dpi = 500, bg = "white")
```

### Values cases

```{r}
options(tibble.print_max = Inf)

pc <- \(x) sprintf("%.1f", 100 * x)

# Case totals are quoted in the text, so settle this first. Keep dedupe FALSE
# if the figure was built from the un-deduplicated frame, so text and figure
# agree; if TRUE, the figure must be rebuilt too.
df_week |>
  summarise(rows = n(), ids = n_distinct(id), repeated_ids = sum(duplicated(id)))

dedupe <- FALSE
cases  <- if (dedupe) distinct(df_week, id, .keep_all = TRUE) else df_week

ep_win <- episodes |>
  select(start, end, weeks) |>
  mutate(episode = str_c(year(start), "-", year(end)), .before = 1)

ep_win |> mutate(months = weeks * 7 / 30.44)

ep_cases <- cases |>
  inner_join(select(ep_win, episode, start, end),
             by = join_by(between(week, start, end))) |>
  mutate(month = floor_date(week, "month"))

ep_totals <- ep_cases |> summarise(n = n(), .by = c(episode, start, end))

ep_totals

# weekly peak: date and size
ep_cases |>
  count(episode, week) |>
  slice_max(n, n = 1, by = episode, with_ties = FALSE)

# cases between the two outbreaks
cases |>
  filter(week > ep_win$end[1], week < ep_win$start[2]) |>
  summarise(n = n(), first = min(week), last = max(week),
            weeks_with_cases = n_distinct(week))

# <5 = "<1" + "1-5"; <10 adds "5-10"
age_mo <- ep_cases |>
  drop_na(age_group) |>
  summarise(n_tot = n(),
            n_u1  = sum(age_group == "<1"),
            n_u5  = sum(age_group %in% c("<1", "1-5")),
            n_u10 = sum(age_group %in% c("<1", "1-5", "5-10")),
            .by = c(episode, month)) |>
  mutate(pct_u1 = n_u1 / n_tot, pct_u5 = n_u5 / n_tot, pct_u10 = n_u10 / n_tot) |>
  arrange(episode, month)

age_mo

# the specific window in your draft: May to July 2024
age_mo |>
  filter(between(month, ymd("2024-05-01"), ymd("2024-07-01"))) |>
  summarise(n_tot = sum(n_tot),
            pct_u5  = sum(n_u5) / sum(n_tot),
            pct_u10 = sum(n_u10) / sum(n_tot))

# first three months of each outbreak, for the parallel claim
age_mo |>
  slice_head(n = 3, by = episode) |>
  summarise(n_tot = sum(n_tot),
            pct_u5  = sum(n_u5) / sum(n_tot),
            pct_u10 = sum(n_u10) / sum(n_tot),
            .by = episode)

# full age breakdown over the same three months
ep_cases |>
  filter(month <= min(month) %m+% months(2), .by = episode) |>
  count(episode, age_group) |>
  mutate(pct = n / sum(n), .by = episode)

coh_mo <- ep_cases |>
  drop_na(cohort) |>
  count(episode, month, cohort) |>
  mutate(n_tot = sum(n), pct = n / sum(n), .by = c(episode, month)) |>
  arrange(episode, month, desc(pct))

coh_mo |>
  mutate(lab = sprintf("%s%% (%d/%d)", pc(pct), n, n_tot)) |>
  select(episode, month, cohort, lab)

# leading cohort each month, which is the trajectory your text describes
coh_mo |>
  slice_max(pct, n = 1, by = c(episode, month), with_ties = FALSE) |>
  mutate(lab = sprintf("%s: %s%% (%d/%d)", cohort, pc(pct), n, n_tot)) |>
  select(episode, month, n_tot, lab)

shade_win |>
  mutate(days = as.numeric(xmax - xmin), months = days / 30.44)

# n and n_tot are both returned: a rising share can occur while counts fall
coh_track <- function(coh, from, to) {
  coh_mo |>
    filter(cohort == coh, between(month, ymd(from), ymd(to))) |>
    mutate(lab = sprintf("%s: %s%% (%d/%d)",
                         format(month, "%b %Y"), pc(pct), n, n_tot)) |>
    select(month, n, n_tot, pct, lab)
}

coh_track("2019-2023", "2019-06-01", "2020-03-31")
coh_track("2019-2023", "2024-05-01", "2024-08-31")
coh_track("2024",      "2024-09-01", "2025-12-31")
coh_track("2025",      "2025-01-01", "2025-12-31")

ep_totals |>
  mutate(txt = sprintf("%s outbreak: %s to %s, %d cases", episode,
                       format(start, "%B %Y"), format(end, "%B %Y"), n)) |>
  pull(txt) |>
  writeLines()

age_mo |>
  filter(month %in% ymd(c("2018-09-01", "2024-05-01", "2024-06-01", "2024-07-01"))) |>
  mutate(txt = sprintf("%s: under 5 %s%%, under 10 %s%% (n = %d)",
                       format(month, "%b %Y"), pc(pct_u5), pc(pct_u10), n_tot)) |>
  pull(txt) |>
  writeLines()
```

### Values cases 2

```{r}
# Panels B and C plot log(rate_ma * 1e5), a daily hazard among those not yet
# recorded as cases. Everything below back-transforms to a rate per 100,000
# per day, which is the quotable form.
r100k <- \(x) x * 1e5

ori_start <- ori_win$xmin[1]
ori_end   <- ori_win$xmax[1]
after_day <- min(ori_end + 90, max(date_grid))

# ori_end is clipped to the data, so check it against data_end before reading
# anything split by phase
tibble(ori_start, ori_end, data_end = max(date_grid), after_day,
       ori_days = as.numeric(ori_end - ori_start))

add_phase <- function(d) {
  mutate(d, phase = factor(
    case_when(date <  ori_start ~ "Before ORI",
              date <= ori_end   ~ "During ORI",
              .default          = "After ORI"),
    levels = c("Before ORI", "During ORI", "After ORI")
  ))
}

snap_tab <- snapshot_dist |>
  select(district, bin_end, cum_n, n_census, cum_inc)

# city-wide cumulative incidence at each snapshot
snap_tab |>
  summarise(cum_n = sum(cum_n), n_census = sum(n_census), .by = bin_end) |>
  mutate(cum_inc = 1e5 * cum_n / n_census) |>
  arrange(bin_end)

# highest and lowest district at each snapshot, and the ratio between them
snap_tab |>
  drop_na(cum_inc) |>
  summarise(hi_d = district[which.max(cum_inc)], hi = max(cum_inc),
            lo_d = district[which.min(cum_inc)], lo = min(cum_inc),
            ratio = hi / lo, n_dist = n(),
            .by = bin_end) |>
  arrange(bin_end)

snap_tab |> filter(bin_end == max(bin_end)) |> arrange(desc(cum_inc))
snap_tab |> filter(bin_end == min(bin_end), cum_n > 0) |> arrange(desc(cum_inc))

peak_dist <- rate_dist |>
  filter(rate_ma > 0) |>
  slice_max(rate_ma, n = 1, by = district, with_ties = FALSE) |>
  transmute(region, district, date, peak_per100k = r100k(rate_ma))

peak_dist |> arrange(desc(peak_per100k))

# when regions reached their peak, to check for a spatial wave
peak_dist |>
  summarise(n_dist = n(), first_peak = min(date), last_peak = max(date),
            med_peak = median(date), .by = region) |>
  arrange(med_peak)

# date each district first exceeded 1 per 100,000 per day
rate_dist |>
  filter(rate_ma >= 1 / 1e5) |>
  slice_min(date, n = 1, by = district, with_ties = FALSE) |>
  transmute(region, district, first_crossing = date) |>
  arrange(first_crossing)

# incidence before, during and after the ORI window, pooled by region
rate_dist |>
  add_phase() |>
  summarise(cases = sum(n), pop = first(n_census), .by = c(region, district, phase)) |>
  summarise(cases = sum(cases), pop = sum(pop), .by = c(region, phase)) |>
  mutate(per100k = 1e5 * cases / pop) |>
  pivot_wider(names_from = phase, values_from = c(cases, per100k))

peak_age <- rate_age |>
  filter(rate_ma > 0) |>
  slice_max(rate_ma, n = 1, by = age_group, with_ties = FALSE) |>
  transmute(age_group, date, peak_per100k = r100k(rate_ma), n_census)

# ratio_to_top compares each group with whichever group peaked highest
peak_age |>
  mutate(ratio_to_top = max(peak_per100k) / peak_per100k) |>
  arrange(desc(peak_per100k))

# same-day comparison at the city-wide peak, which is the cleaner statement
peak_day <- rate_age |>
  summarise(tot = sum(n), .by = date) |>
  slice_max(tot, n = 1, with_ties = FALSE) |>
  pull(date)

peak_day

rate_age |>
  filter(date == peak_day) |>
  transmute(age_group, per100k = r100k(rate_ma),
            ratio_to_top = max(rate_ma, na.rm = TRUE) / rate_ma) |>
  arrange(desc(per100k))

rate_age |>
  add_phase() |>
  summarise(cases = sum(n), pop = first(n_census), .by = c(age_group, phase)) |>
  mutate(per100k = 1e5 * cases / pop) |>
  pivot_wider(names_from = phase, values_from = c(cases, per100k))

age_inc <- rate_age |>
  summarise(cases = sum(n), pop = first(n_census), .by = age_group) |>
  mutate(per100k = 1e5 * cases / pop, pct_of_cases = 100 * cases / sum(cases))

age_inc |> arrange(desc(per100k))

# hazard at ORI end vs 90 days later
rate_age |>
  filter(date %in% c(ori_end, after_day)) |>
  transmute(age_group, when = if_else(date == ori_end, "at_end", "later"),
            per100k = r100k(rate_ma)) |>
  pivot_wider(names_from = when, values_from = per100k) |>
  mutate(fold_drop = at_end / later)

# Cumulative incidence recomputed from summed cases and population, not from
# averaged district rates
dist_inc <- rate_dist |>
  summarise(cases = sum(n), pop = first(n_census), .by = c(region, district)) |>
  mutate(per100k = 1e5 * cases / pop)

dist_inc |>
  summarise(n_dist = n(), cases = sum(cases), pop = sum(pop), .by = region) |>
  mutate(per100k = 1e5 * cases / pop) |>
  arrange(desc(per100k))

# by district, so you can see whether West leads as a block or through one or
# two districts
dist_inc |> arrange(desc(per100k))

# the "ranging from ... to ..." you want to quote
rate_dist |>
  filter(rate_ma > 0) |>
  summarise(min_per100k = r100k(min(rate_ma)),
            med_per100k = r100k(median(rate_ma)),
            max_per100k = r100k(max(rate_ma)),
            .by = region) |>
  arrange(desc(max_per100k))

reg_daily <- rate_dist |>
  summarise(n = sum(n), n_census = sum(n_census), .by = c(region, date)) |>
  add_attack_rate(region)

# share of days on which each region held the highest hazard
reg_daily |>
  filter(rate_ma > 0) |>
  mutate(top = rate_ma == max(rate_ma), .by = date) |>
  summarise(days_top = sum(top), days = n(), .by = region) |>
  mutate(pct_top = 100 * days_top / days) |>
  arrange(desc(pct_top))

# monthly regional hazard, to see whether the lead holds over time
reg_daily |>
  drop_na(rate_ma) |>
  mutate(mo = floor_date(date, "month")) |>
  summarise(med_per100k = r100k(median(rate_ma)), .by = c(region, mo)) |>
  pivot_wider(names_from = region, values_from = med_per100k) |>
  arrange(mo)

snap_tab |>
  filter(bin_end == max(bin_end)) |>
  slice_max(cum_inc, n = 3) |>
  mutate(txt = sprintf("%s: %.0f per 100,000 (%d cases)", district, cum_inc, cum_n)) |>
  pull(txt) |>
  writeLines()

age_inc |>
  mutate(txt = sprintf("%s: %.0f per 100,000 over the episode (%d cases)",
                       age_group, per100k, cases)) |>
  pull(txt) |>
  writeLines()

rate_dist |>
  filter(date < ori_start, rate_ma > 0) |>
  summarise(min_per100k = r100k(min(rate_ma)),
            max_per100k = r100k(max(rate_ma)),
            n_days = n_distinct(date),
            .by = region) |>
  mutate(txt = sprintf("%s: %.2f to %.2f per 100,000 per day (%d days)",
                       region, min_per100k, max_per100k, n_days)) |>
  arrange(desc(max_per100k)) |>
  pull(txt) |>
  writeLines()
```

## Serology

Children of vaccination age in 3 hospitals, sampled 01-08-2022 to 30-04-2024

```{r}
hosp_lab <- c("Children's hospital 1"    = "CH1",
              "Children's hospital 2"    = "CH2",
              "City Children's Hospital" = "CCH")

df_sero_3ch <- df_sero |>
  filter(hospital %in% names(hosp_lab),
         between(date_collection, dmy("01-08-2022"), dmy("30-04-2024"))) |>
  mutate(hospital = factor(hosp_lab[as.character(hospital)], levels = hosp_lab))

df_sero_3ch_mod <- df_sero_3ch |>
  filter(exact_age >= 9 / 12) |>
  mutate(age_5y = factor(if_else(exact_age < 5, "<5", ">=5"), c("<5", ">=5")),
         age_ceiling = ceiling(exact_age)) |>
  left_join(df_census_age, join_by(age_ceiling == age)) |>
  mutate(weight = n_census / mean(n_census))
```

### Hospital coverage

```{r}
setdiff(df_sero_3ch$district, df_map_hcmc$district_blank)   # want character(0)

brks_age <- c(0, 9/12, 1, 2, 3, 4, 5, 10, 15)
age_labs <- c("<9m", "9-11m", "1y", "2y", "3y", "4y", "5-9y", "10-15y")

df_age <- df_sero_3ch |>
  mutate(
    age_band = cut(exact_age, brks_age, age_labs,
                   right = FALSE, include.lowest = TRUE),
    period_lab = factor(format(period_collect, "%b %Y"),
                        levels = format(sort(unique(period_collect)), "%b %Y"))
  ) |>
  count(hospital, period_lab, age_band) |>
  complete(nesting(hospital, period_lab), age_band, fill = list(n = 0))

xmax <- max(df_age$n)

plot_hosp <- function(d, x_axis = FALSE, y_lab = FALSE) {
  ggplot(d, aes(x = n, y = age_band, fill = age_band == "<9m")) +
    geom_col() +
    facet_grid(hospital ~ period_lab, switch = "y") +
    scale_fill_manual(values = c(`FALSE` = "grey40", `TRUE` = "#D55E00"),
                      guide = "none") +
    scale_x_continuous(
      limits = c(0, xmax),
      # drop the top break, its label would sit on the panel edge
      breaks = \(l) { b <- breaks_pretty(3)(l); b[b <= 0.85 * l[2]] },
      expand = expansion(mult = c(0, 0.08))
    ) +
    labs(x = if (x_axis) "Number of samples" else NULL, y = "Age band") +
    theme_bw() +
    theme(panel.grid.minor = element_blank(),
          strip.placement = "outside",
          # NA keeps the space reserved, so only one row shows the title
          axis.title.y = element_text(colour = if (y_lab) "black" else NA),
          axis.text.x  = if (x_axis) element_text() else element_blank(),
          axis.ticks.x = if (x_axis) element_line() else element_blank())
}

p_list <- df_age |> group_split(hospital) |> map(droplevels)
n_h    <- length(p_list)

grobs <- p_list |>
  imap(\(d, i) plot_hosp(d, x_axis = i == n_h, y_lab = i == ceiling(n_h / 2))) |>
  map(ggplotGrob)

# Hospitals have different numbers of periods, so narrower gtables get extra
# panel and spacing columns until all of them are the same width
pan_l <- \(g) g$layout$l[startsWith(g$layout$name, "panel")]
g_ref <- grobs[[which.max(map_int(grobs, \(g) length(pan_l(g))))]]
gap   <- g_ref$widths[max(pan_l(g_ref)) - 1]

pad_to <- function(g, target) {
  need <- target - ncol(g)
  if (need <= 0) return(g)
  pos <- max(pan_l(g))
  w   <- rep(list(gap, unit(1, "null")), length.out = need)
  if (need %% 2 == 1) w[[need]] <- unit(0, "cm")
  for (i in seq_len(need)) {
    g   <- gtable_add_cols(g, w[[i]], pos)
    pos <- pos + 1
  }
  g
}

grobs       <- map(grobs, pad_to, target = max(map_int(grobs, ncol)))
grobs[-n_h] <- map(grobs[-n_h], \(g) gtable_add_rows(g, unit(4, "mm")))

p_age <- reduce(grobs, \(a, b) rbind(a, b, size = "max"))

# Map margins matched to the space above the first and below the last panel
cm_h  <- \(u) convertHeight(u, "cm", valueOnly = TRUE)
pan_t <- \(g) g$layout$t[startsWith(g$layout$name, "panel")][1]
m_top <- cm_h(sum(grobs[[1]]$heights[seq_len(pan_t(grobs[[1]]) - 1)]))
m_bot <- cm_h(sum(grobs[[n_h]]$heights[-seq_len(pan_t(grobs[[n_h]]))]))

pct_min <- 1                                             # below this % draws grey
mid     <- 100 / n_distinct(df_map_hcmc$district_blank)  # equal-share reference

df_map_pct <- df_sero_3ch |>
  count(hospital, district) |>
  complete(hospital, district = df_map_hcmc$district_blank, fill = list(n = 0)) |>
  mutate(pct = 100 * n / sum(n), .by = hospital) |>
  mutate(lfc = if_else(pct >= pct_min, log2(pct / mid), NA)) |>
  left_join(select(df_map_hcmc, district_blank),
            join_by(district == district_blank)) |>
  st_as_sf()

lo <- log2(pct_min / mid)                # bar starts exactly at pct_min
hi <- max(df_map_pct$lfc, na.rm = TRUE)

brks_pct <- c(1, 2, 5, 10, 20, 40)
brks_pct <- brks_pct[brks_pct >= pct_min & log2(brks_pct / mid) <= hi]

p_map_col <- df_map_pct |>
  ggplot() +
  geom_sf(aes(fill = lfc), colour = "grey40", linewidth = 0.15) +
  facet_wrap(~ hospital, ncol = 1) +
  coord_sf(expand = FALSE) +
  scale_fill_gradientn(
    colours  = pal_gsea()(12),
    limits   = c(lo, hi),
    na.value = "grey88",
    rescaler = \(x, to = c(0, 1), from = range(x, na.rm = TRUE))
                 rescale_mid(x, to, from, mid = 0),
    breaks   = log2(brks_pct / mid),
    labels   = paste0(brks_pct, "%"),
    name     = "% of samples",
    guide    = guide_colourbar(
      theme = theme(legend.key.height = unit(2.6, "cm"),
                    legend.key.width  = unit(0.35, "cm"))
    )
  ) +
  theme_void() +
  theme(strip.text      = element_blank(),
        legend.position = "right",
        legend.margin   = margin(l = 0.25, unit = "cm"),
        legend.title    = element_text(size = 10),
        legend.text     = element_text(size = 9),
        panel.spacing.y = unit(0, "cm"),
        plot.margin     = margin(t = m_top, b = m_bot, l = 4, r = 0, unit = "cm"))

# p_age is a raw gtable so it needs wrap_elements; the maps stay a ggplot so
# patchwork can size them
p_all <- wrap_elements(full = p_age) + p_map_col +
  plot_layout(widths = c(2, 1)) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(face = "bold", size = 14),
        plot.tag.position = c(0, 1),
        plot.margin = margin(t = 10, r = 4, b = 4, l = 4))

p_all

# ggsave("figs/chap1/age_and_map_combined.png", p_all,
#        width = 9, height = 5, dpi = 300, bg = "white")
```

### GAM model

```{r}
sens   <- 0.99
spec   <- 0.95
critp  <- 0.95
n_grid <- 200      # points per fitted curve
w_date <- 1.4      # width of the time panel
pad_v  <- 1        # pt of padding above/below each row
pad_s  <- 2        # pt of padding inside facet strips

# Collection dates closer than this share a column across hospitals.
# 45 days merges Mar with Apr and Aug with Sep, but keeps Apr and Jun apart.
slot_gap <- 45

col_age  <- c("<5" = "#00BA38", ">=5" = "#619CFF")
lab_all  <- "All periods"
lab_time <- "Over time"
lab_data <- "Data mean (95% CI)"
lab_gam  <- "GAM fit"

per_lvl <- df_sero_3ch_mod |>
  distinct(period_collect) |>
  arrange(period_collect) |>
  pull(period_collect) |>
  format("%b %Y")

fct_period <- function(x) factor(as.character(x), levels = c(lab_all, per_lvl))

df_mod <- df_sero_3ch_mod |>
  mutate(pos = as.integer(titer > 200),
         period_lab = fct_period(format(period_collect, "%b %Y"))) |>
  drop_na(pos, exact_age, date_collection)

slot_name <- function(d) {
  a <- min(d); b <- max(d)
  case_when(a == b             ~ format(a, "%b %Y"),
            year(a) == year(b) ~ str_c(format(a, "%b"), "-", format(b, "%b %Y")),
            .default           = str_c(format(a, "%b %Y"), "-", format(b, "%b %Y")))
}

# Shared columns: a new one starts when the gap exceeds slot_gap
slots <- df_mod |>
  distinct(period_collect) |>
  arrange(period_collect) |>
  mutate(gap = as.numeric(period_collect - lag(period_collect,
                                               default = first(period_collect))),
         slot_id = cumsum(gap > slot_gap)) |>
  mutate(slot = slot_name(period_collect), .by = slot_id)

slot_lvl <- slots |> distinct(slot_id, slot) |> arrange(slot_id) |> pull(slot)
fct_slot <- function(x) factor(as.character(x), levels = c(lab_all, slot_lvl))

df_mod <- df_mod |>
  left_join(select(slots, period_collect, slot), by = "period_collect") |>
  mutate(slot = fct_slot(slot))

slot_tab <- df_mod |> distinct(hospital, slot, period_lab) |> arrange(hospital, slot)

# Two periods from one hospital in one column would overplot, so stop here
if (any(duplicated(select(slot_tab, hospital, slot)))) {
  stop("A hospital has two periods in one column. Lower slot_gap.")
}

# Check the alignment before trusting the figure
slot_tab |> pivot_wider(names_from = slot, values_from = period_lab)

hosp     <- levels(df_mod$hospital)
mid_h    <- hosp[ceiling(length(hosp) / 2)]
end_h    <- hosp[length(hosp)]
x_age    <- c(0, ceiling(max(df_mod$exact_age, na.rm = TRUE)))
slot_all <- levels(df_mod$slot)

# NOTE: weighted here, unweighted in fit_age(). Decide which before quoting.
pred_date <- df_mod |>
  nest(.by = c(hospital, age_5y)) |>
  mutate(pred = map(data, \(d) {
    gam(pos ~ s(as.numeric(date_collection)), family = binomial,
        weights = weight, data = d, method = "REML") |>
      predict_ci(d, "date_collection")
  })) |>
  select(-data) |>
  unnest(pred)

agg_sero <- df_mod |>
  summarise(npos = sum(pos), ntot = n(),
            .by = c(hospital, age_5y, period_collect)) |>
  mutate(prop_ci(npos, ntot)) |>
  mutate(prev_true = prev_correct(npos / ntot),
         lwr_true  = prev_correct(ci_lwr),
         upr_true  = prev_correct(ci_upr))

fit_age <- function(d) {
  nu <- n_distinct(d$exact_age)
  if (nrow(d) < 10 || nu < 4) return(NULL)      # too thin for a smooth
  gam(pos ~ s(exact_age, k = max(3, min(10, nu - 1))),
      family = binomial, data = d, method = "REML") |>
    predict_ci(d, "exact_age")
}

fit_age_safe <- possibly(fit_age, otherwise = NULL)

# slot places the panel, strip labels it with the hospital's own date
pred_age <- bind_rows(
  df_mod |>
    nest(.by = hospital) |>
    mutate(pred = map(data, fit_age_safe), slot = lab_all, strip = lab_all),
  df_mod |>
    nest(.by = c(hospital, slot)) |>
    mutate(pred = map(data, fit_age_safe)) |>
    left_join(slot_tab, by = join_by(hospital, slot)) |>
    mutate(strip = as.character(period_lab))
) |>
  select(hospital, slot, strip, pred) |>
  unnest(pred) |>
  mutate(slot = fct_slot(slot),
         agegr = factor(if_else(exact_age < 5, "<5", ">=5"), names(col_age)))

# Rug data: each observation once in its own column, once in the pooled panel
rug_age <- df_mod |>
  select(hospital, exact_age, pos, slot, period_lab) |>
  mutate(strip = as.character(period_lab))

rug_age <- bind_rows(rug_age,
                     mutate(rug_age, slot = fct_slot(lab_all), strip = lab_all))

theme_sero <- theme_light() +
  theme(
    strip.placement  = "outside",
    strip.clip       = "off",
    strip.text       = element_text(colour = "grey30",
                                    margin = margin(pad_s, pad_s, pad_s, pad_s)),
    strip.background = element_rect(fill = NA, colour = NA),
    panel.spacing    = unit(0, "cm"),
    panel.border     = element_rect(colour = NA),
    panel.grid       = element_blank(),
    axis.line        = element_line(colour = "black")
  )

# limits pin both keys, so a single-age-group panel cannot produce a second
# legend when patchwork collects them
layers_sero <- list(
  geom_hline(yintercept = critp, colour = "darkred", linetype = "dashed"),
  scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(-0.1, 1.1),
                     labels = label_percent(suffix = "")),
  scale_colour_manual(values = col_age, limits = names(col_age),
                      name = "Age group"),
  scale_fill_manual(values = col_age, limits = names(col_age), guide = "none"),
  scale_linetype_manual(values = set_names(1, lab_gam), limits = lab_gam,
                        name = NULL),
  guides(colour   = guide_legend(order = 1, override.aes = list(linetype = 1)),
         shape    = guide_legend(order = 2,
                                 override.aes = list(colour = "grey30", alpha = 1)),
         linetype = guide_legend(order = 3, override.aes = list(colour = "grey30"))),
  theme_sero
)

p_date <- map(hosp, \(h) {
  d_pt <- filter(df_mod,    hospital == h) |> mutate(facet = lab_time)
  d_pr <- filter(pred_date, hospital == h) |> mutate(facet = lab_time)
  d_ag <- filter(agg_sero,  hospital == h) |> mutate(facet = lab_time)

  ym <- tibble(d = sort(unique(d_pt$period_collect))) |>
    mutate(txt = if_else(!duplicated(year(d)), format(d, "%b\n%Y"), format(d, "%b")))

  # Breaks sit at the start of each period, which can fall before the first
  # sample date, so span the limits over both or the label gets dropped
  x_rng <- range(c(ym$d, d_pt$date_collection, d_ag$period_collect + days(15)))

  ggplot() +
    geom_point(data = d_pt, aes(date_collection, pos), shape = 124, alpha = 0.3) +
    geom_ribbon(data = d_pr,
                aes(date_collection, ymin = lwr_true, ymax = upr_true, fill = age_5y),
                alpha = 0.15, show.legend = FALSE) +
    geom_line(data = d_pr,
              aes(date_collection, fit_true, colour = age_5y, linetype = lab_gam)) +
    geom_pointrange(data = d_ag,
                    aes(period_collect + days(15), prev_true,
                        ymin = lwr_true, ymax = upr_true,
                        colour = age_5y, shape = lab_data),
                    alpha = 0.6,
                    show.legend = c(colour = FALSE, fill = FALSE, shape = TRUE)) +
    facet_grid(hospital ~ facet, switch = "y") +
    scale_x_date(breaks = ym$d, labels = ym$txt, limits = x_rng,
                 expand = expansion(mult = 0.04)) +
    scale_shape_manual(values = set_names(16, lab_data), name = NULL) +
    labs(x = NULL, y = "Seroprevalence (%)") +
    layers_sero +
    # NA keeps the horizontal space, so panels start at the same x in every row
    theme(axis.title.y = element_text(colour = if (h == mid_h) "black" else NA))
}) |>
  set_names(hosp)

# The x title goes on one panel of the bottom row: repeated under every column
# it would be noise, and patchwork sizes the title row across the whole row
mid_slot <- rug_age |>
  filter(hospital == end_h) |>
  distinct(slot) |>
  arrange(slot) |>
  slice(ceiling(n() / 2)) |>
  pull(slot) |>
  as.character()

# A hospital with no sample in a column gets a spacer, which draws nothing
panel_age <- function(h, s) {
  d_pt <- filter(rug_age, hospital == h, slot == s)
  if (nrow(d_pt) == 0) return(plot_spacer())
  d_pr <- filter(pred_age, hospital == h, slot == s)

  ggplot() +
    geom_point(data = d_pt, aes(exact_age, pos), shape = 124, alpha = 0.3) +
    geom_ribbon(data = d_pr,
                aes(exact_age, ymin = lwr_true, ymax = upr_true, fill = agegr),
                alpha = 0.15, show.legend = FALSE) +
    geom_line(data = d_pr,
              aes(exact_age, fit_true, colour = agegr, linetype = lab_gam,
                  group = 1)) +
    facet_wrap(~strip) +
    # top break dropped, its label would collide with the next panel's zero
    scale_x_continuous(breaks = seq(0, x_age[2] - 1, 5), limits = x_age,
                       expand = expansion(mult = 0.02)) +
    labs(x = if (h == end_h && s == mid_slot) "Age (years)" else NULL, y = NULL) +
    layers_sero +
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
}

fig <- map(hosp, \(h) {
  wrap_plots(c(list(p_date[[h]]), map(slot_all, \(s) panel_age(h, s))),
             nrow = 1, widths = c(w_date, rep(1, length(slot_all))))
}) |>
  wrap_plots(ncol = 1, guides = "collect") &
  theme(legend.position = "bottom",
        plot.margin     = margin(t = pad_v, r = 2, b = pad_v, l = 2))

fig

# ggsave("figs/chap1/fig_seroprev_by_hospital_period.png", fig,
#        width = 11, height = 4.5, dpi = 300)
```

### Sero values

```{r}
# All *_true columns are corrected for assay sensitivity and specificity;
# quote those, not the raw values.
options(tibble.print_max = Inf)

pc  <- \(x) sprintf("%.1f", 100 * x)
fmt <- \(d) sprintf("%s%% (%s to %s), %d/%d",
                    pc(d$prev_true), pc(d$lwr_true), pc(d$upr_true), d$npos, d$ntot)

add_prev <- function(d) {
  d |>
    mutate(prop_ci(npos, ntot)) |>
    mutate(prev_true = prev_correct(npos / ntot),
           lwr_true  = prev_correct(ci_lwr),
           upr_true  = prev_correct(ci_upr)) |>
    mutate(lab = fmt(pick(everything())))
}

n_cell <- df_mod |> summarise(npos = sum(pos), ntot = n(), .by = c(hospital, age_5y))

n_cell |> pivot_wider(names_from = age_5y, values_from = c(ntot, npos))

df_mod |> summarise(n = n(), min_age = min(exact_age), max_age = max(exact_age))

# observed seroprevalence per period
agg_sero |>
  mutate(lab = fmt(pick(everything()))) |>
  select(hospital, age_5y, period_collect, ntot, prev_true, lab) |>
  arrange(hospital, age_5y, period_collect)

# pooled across periods, per hospital and age group
n_cell |> add_prev() |> select(hospital, age_5y, lab)

# range across periods: how much the point estimates moved
agg_sero |>
  summarise(n_periods = n(), min = min(prev_true), max = max(prev_true),
            range_pp = 100 * (max - min),
            .by = c(hospital, age_5y))

# how many period estimates sit below the critical threshold
agg_sero |>
  summarise(n = n(),
            n_below = sum(prev_true < critp),
            n_ci_below = sum(upr_true < critp),      # CI entirely below
            .by = c(hospital, age_5y))

# Pooled fit only, since the per-period curves are too thin to quote
pooled_age <- pred_age |> filter(slot == lab_all)

# nearest grid point to each reference age
pooled_age |>
  cross_join(tibble(ref = c(1, 2, 3, 5, 10, 15))) |>
  slice_min(abs(exact_age - ref), n = 1, by = c(hospital, ref), with_ties = FALSE) |>
  mutate(lab = sprintf("%s%% (%s to %s)", pc(fit_true), pc(lwr_true), pc(upr_true))) |>
  select(hospital, ref, age = exact_age, fit_true, lab) |>
  arrange(hospital, ref)

# minimum of the pooled curve and where it occurs: the susceptibility trough
trough <- pooled_age |>
  slice_min(fit_true, n = 1, by = hospital, with_ties = FALSE) |>
  mutate(lab = sprintf("%s%% (%s to %s) at age %.1f y",
                       pc(fit_true), pc(lwr_true), pc(upr_true), exact_age))

trough |> select(hospital, exact_age, fit_true, lab)

# age range over which the fitted curve sits below the critical threshold,
# and where the upper CI is entirely below it, which is the stronger claim
pooled_age |>
  summarise(fit_from = min(exact_age[fit_true < critp]),
            fit_to   = max(exact_age[fit_true < critp]),
            ci_from  = min(exact_age[upr_true < critp]),
            ci_to    = max(exact_age[upr_true < critp]),
            .by = hospital)

# fitted time curve: start, end, and change
pred_date |>
  slice(c(which.min(date_collection), which.max(date_collection)),
        .by = c(hospital, age_5y)) |>
  mutate(point = if_else(date_collection == min(date_collection), "start", "end"),
         .by = c(hospital, age_5y)) |>
  select(hospital, age_5y, point, fit_true, lwr_true, upr_true) |>
  pivot_wider(names_from = point, values_from = c(fit_true, lwr_true, upr_true)) |>
  mutate(change_pp = 100 * (fit_true_end - fit_true_start))

# under-5 versus 5-and-over, the headline contrast
df_mod |>
  summarise(npos = sum(pos), ntot = n(), .by = age_5y) |>
  add_prev() |>
  select(age_5y, lab)

trough |>
  mutate(txt = sprintf("%s: fitted seroprevalence reaches a minimum of %s at %s",
                       hospital, str_remove(lab, " at age.*"),
                       sprintf("age %.1f years", exact_age))) |>
  pull(txt) |>
  writeLines()
```

## Vaccination coverage

Children born since 01-01-2017

```{r}
df_vax_hcmc |> filter(dob >= dmy("01-01-2017")) |> nrow()
```

## Sero vs coverage

```{r}
set.seed(2026)
options(tibble.print_max = Inf)

study_start <- as.Date("2022-09-05")
study_end   <- as.Date("2024-04-30")

# Comparison window for the scatter. Set these to the sero collection window
# (min/max of df_mod$date_collection) if you want x and y over the same months.
# Over the full period MCV2 coverage rises about 23 points, so a min-max bar
# spans most of the axis and the x position stops meaning anything.
win_start <- study_start
win_end   <- study_end

m_draws     <- 1000L   # Monte Carlo draws for VE parameters
n_min       <- 5       # minimum sera per region to plot
diag_expand <- 0.35    # fraction of the x range the 1:1 line should cross

wk1    <- floor_date(study_start, "week", week_start = 1)
weeks  <- seq(wk1, study_end, by = "7 days")
n_week <- length(weeks)

# index of the first week whose Monday falls on or after date d
wk_ge <- function(d) (as.integer(unclass(d) - unclass(wk1)) + 6L) %/% 7L + 1L

# calendar-month arithmetic on distinct values only, then matched back
add_months_fast <- function(d, n) {
  u <- unique(d)
  (u %m+% months(n))[match(d, u)]
}

metrics <- tibble(
  dose   = c("Dose 1", "Dose 2", "EVC"),
  metric = c("MCV1 coverage", "MCV2 coverage", "Estimated immunity"),
  short  = c("MCV1", "MCV2", "EVC"),
  lty    = c("solid", "dashed", "dotted"),
  age_lo = c(9, 18, 9) / 12,     # years, inclusive
  age_hi = 5                     # years, exclusive
)

metric_lvl <- metrics$metric
metric_of  <- deframe(select(metrics, dose, metric))

# Same lookup as every other figure in the chapter, so the city is split the
# same way. Anything unmatched becomes NA rather than landing in Central.
norm_dist <- \(x) str_squish(str_to_lower(x))

region_of <- district_region |>
  mutate(district = norm_dist(district)) |>
  deframe()

assign_region <- function(x) factor(unname(region_of[norm_dist(x)]),
                                    levels = lv_region)

# Belongs in the shared setup next to district_region: these must match the
# outbreak figure or the same region changes colour between figures
pal_region <- c(West = "#CC6677", North = "#88CCEE", East = "#DDCC77",
                South = "#332288", Central = "#117733")

child <- df_vax_hcmc |>
  as_tibble() |>
  transmute(
    district = norm_dist(district_blank),
    region   = assign_region(district_blank),
    dob = as.Date(dob),
    d1  = as.Date(date_m1),
    d2  = as.Date(date_m2),
    t9  = add_months_fast(dob,  9L),     # eligible for dose 1
    t18 = add_months_fast(dob, 18L),     # eligible for dose 2
    t60 = add_months_fast(dob, 60L)      # turns 5, leaves the denominator
  ) |>
  filter(dob <= study_end, t60 > wk1, !is.na(district), district != "")

# CHECK: districts with no region are dropped from every panel. Read this.
child |> filter(is.na(region)) |> count(district, sort = TRUE)
child |> count(region, district, sort = TRUE)

# EVC denominator = the dose 1 population (9-59 mo). Waning is timed from the
# dose date; a child with only a dose 2 record is treated as single-dose.
a_den      <- wk_ge(child$t9)
b_exit     <- wk_ge(child$t60) - 1L
t12        <- add_months_fast(child$dob, 12L)
first_date <- coalesce(child$d1, child$d2)
s_first    <- wk_ge(first_date)
s_second   <- if_else(is.na(child$d1), NA_integer_, wk_ge(child$d2))
band_first <- if_else(first_date < t12, 1L, 2L)   # 1 = 9-11 mo, 2 = >= 12 mo
b_one      <- if_else(is.na(s_second), b_exit, pmin(b_exit, s_second - 1L))

sum(is.na(child$d1) & !is.na(child$d2))   # orphan dose 2 records

# Normal on the logit scale for the VE proportions (keeps draws inside (0,1),
# which matters at 99.6%), log scale for the reversion rate (keeps it
# positive). VE depends on dose count, not a product across doses, because the
# MCV2 estimate is a two-dose effectiveness.
# NOTE the 9-11 mo CI is asymmetric on the logit scale, so centring on the best
# estimate gives a sampled interval of about (66%, 94%) against the reported
# (72%, 95%): wider and lower, so conservative.
ci_sd <- function(lo, hi, trans) (trans(hi) - trans(lo)) / 3.92

draws <- tibble(
  ve1_L = plogis(rnorm(m_draws, qlogis(0.840), ci_sd(0.720, 0.950, qlogis))),
  ve1_H = plogis(rnorm(m_draws, qlogis(0.925), ci_sd(0.848, 0.970, qlogis))),
  ve2   = plogis(rnorm(m_draws, qlogis(0.996), ci_sd(0.993, 0.998, qlogis))),
  p_rev = exp(rnorm(m_draws, log(0.0022), ci_sd(0.0017, 0.0028, log)))
)

q_wk <- -log1p(-draws$p_rev) * 7 / 365.25   # weekly decay from annual reversion

draws |>                                    # sampled vs reported intervals
  summarise(across(everything(),
                   list(med = median,
                        lo = \(x) quantile(x, 0.025),
                        hi = \(x) quantile(x, 0.975))))

# Each child contributes one interval, so counting is a +1/-1 event table plus
# a cumsum: O(children), not O(children x weeks). key adds a dose-week
# dimension, which is what lets
#   sum_i VE_i(w) = mult * exp(-q w) * sum_s N[w, s] exp(q s)
# be evaluated with two small matrix products per term instead of m_draws
# passes over the registry.
count_intervals <- function(a, b, g, ng, key = rep(1L, length(a))) {
  a  <- pmax(a, 1L)
  b  <- pmin(b, n_week)
  ok <- !is.na(a) & !is.na(b) & !is.na(g) & !is.na(key) & b >= a
  if (!any(ok)) return(list(n = matrix(0, n_week, ng), z = 0L, nk = 1L))
  a <- a[ok]; b <- b[ok]; g <- g[ok]; key <- key[ok]

  z   <- min(key):max(key)
  nk  <- length(z)
  nb  <- ng * nk * (n_week + 1L)
  off <- ((g - 1L) * nk + (key - z[1L])) * (n_week + 1L)
  d   <- tabulate(off + a, nb) - tabulate(off + b + 1L, nb)

  n <- matrix(d, nrow = n_week + 1L) |> apply(2, cumsum)
  list(n = n[seq_len(n_week), , drop = FALSE], z = z, nk = nk)
}

cover <- function(a, b, g, ng) count_intervals(a, b, g, ng)$n

term_matrices <- function(a, b, g, key, mult, ng) {
  kc <- count_intervals(a, b, g, ng, key)
  Ez <- exp(outer(kc$z, q_wk))                 # n_key  x M
  Ew <- exp(outer(seq_len(n_week), -q_wk))     # n_week x M
  map(seq_len(ng), \(gg) {
    cols <- ((gg - 1L) * kc$nk + 1L):(gg * kc$nk)
    (kc$n[, cols, drop = FALSE] %*% Ez) * Ew * rep(mult, each = n_week)
  })
}

build_series <- function(grp) {
  lev <- levels(droplevels(grp))
  ng  <- length(lev)
  gi  <- as.integer(factor(as.character(grp), levels = lev))

  mk <- function(entry, exit, vax, dose_lab) {
    b   <- wk_ge(exit) - 1L
    a_d <- wk_ge(entry)
    a_n <- pmax(a_d, wk_ge(vax))      # doses before entry count from entry
    tibble(
      week    = rep(weeks, times = ng),
      region  = rep(lev, each = n_week),
      dose    = dose_lab,
      n_denom = as.vector(cover(a_d, b, gi, ng)),
      n_vax   = as.vector(cover(a_n, b, gi, ng))
    )
  }

  cov_df <- bind_rows(
    mk(child$t9,  child$t60, child$d1, "Dose 1"),
    mk(child$t18, child$t60, child$d2, "Dose 2")
  ) |>
    mutate(region = factor(region, levels = lev),
           coverage = if_else(n_denom > 0, n_vax / n_denom, NA))

  numM <- replicate(ng, matrix(0, n_week, m_draws), simplify = FALSE)
  for (bb in 1:2) {                                   # single-dose periods
    sel <- which(!is.na(s_first) & band_first == bb)
    if (length(sel)) numM <- Map(`+`, numM, term_matrices(
      pmax(a_den, s_first)[sel], b_one[sel], gi[sel], s_first[sel],
      if (bb == 1L) draws$ve1_L else draws$ve1_H, ng))
  }
  sel <- which(!is.na(s_second))                      # two-dose periods
  if (length(sel)) numM <- Map(`+`, numM, term_matrices(
    pmax(a_den, s_second)[sel], b_exit[sel], gi[sel], s_second[sel],
    draws$ve2, ng))

  denE <- cover(a_den, b_exit, gi, ng)

  evc_df <- map2(seq_len(ng), lev, \(gg, nm) {
    E  <- numM[[gg]] / na_if(denE[, gg], 0L)
    qs <- apply(E, 1, quantile, probs = c(0.5, 0.025, 0.975), na.rm = TRUE)
    tibble(week = weeks, region = nm,
           evc = qs[1, ], evc_lo = qs[2, ], evc_hi = qs[3, ])
  }) |>
    list_rbind() |>
    mutate(region = factor(region, levels = lev))

  idx <- which(weeks >= win_start & weeks <= win_end)

  evc_win <- map2(seq_len(ng), lev, \(gg, nm) {
    dsum <- sum(denE[idx, gg])
    pa   <- if (dsum > 0) colSums(numM[[gg]][idx, , drop = FALSE]) / dsum
            else rep(NA_real_, m_draws)
    w <- evc_df$evc[evc_df$region == nm & between(evc_df$week, win_start, win_end)]
    tibble(region = nm,
           x     = median(pa),
           xu_lo = unname(quantile(pa, 0.025, na.rm = TRUE)),
           xu_hi = unname(quantile(pa, 0.975, na.rm = TRUE)),
           x_lo  = min(w, na.rm = TRUE),
           x_hi  = max(w, na.rm = TRUE))
  }) |>
    list_rbind()

  cov_win <- cov_df |>
    filter(between(week, win_start, win_end), n_denom > 0) |>
    summarise(x = sum(n_vax) / sum(n_denom),
              x_lo = min(coverage), x_hi = max(coverage),
              .by = c(region, dose)) |>
    mutate(xu_lo = NA_real_, xu_hi = NA_real_)

  list(cov = cov_df, evc = evc_df, cov_win = cov_win, evc_win = evc_win)
}

ser <- build_series(child$region)

# Dose 1 is near-flat across regions, so dose 2 separates them. Pooled over the
# period, so the order does not hinge on one week.
rank_order <- ser$cov |>
  filter(dose == "Dose 2") |>
  summarise(m = sum(n_vax) / sum(n_denom), .by = region) |>
  arrange(desc(m)) |>
  pull(region) |>
  as.character()

rank_order

ser$cov |> filter(coverage > 1) |> nrow()                  # should be 0
ser$evc |> summarise(lo = min(evc_lo), hi = max(evc_hi))   # in [0, 1]

# MCV1 and EVC use a 9-59 month denominator; MCV2 uses 18-59 months. Sera are
# restricted to the matching window for each panel so x and y describe the same
# population.
sero_src <- df_mod |>
  as_tibble() |>
  mutate(district = norm_dist(district), region = assign_region(district)) |>
  filter(!is.na(pos), !is.na(exact_age), !is.na(district), district != "")

sero_src |> summarise(n = n(), min = min(exact_age), med = median(exact_age),
                      max = max(exact_age))

# CHECK: unmatched district strings drop silently
setdiff(unique(sero_src$district), unique(child$district))

# Clopper-Pearson. prop_ci() in the seroprevalence figure gives Wilson: pick one
sero_ci <- function(d) {
  d |>
    summarise(n_sero = n(), x_pos = sum(pos), .by = region) |>
    mutate(region  = as.character(region),
           sero    = x_pos / n_sero,
           sero_lo = if_else(x_pos == 0, 0,
                             qbeta(0.025, x_pos, n_sero - x_pos + 1)),
           sero_hi = if_else(x_pos == n_sero, 1,
                             qbeta(0.975, x_pos + 1, n_sero - x_pos)))
}

sero_all <- metrics |>
  transmute(metric,
            res = map2(age_lo, age_hi, \(lo, hi)
                       sero_src |> filter(exact_age >= lo, exact_age < hi) |>
                         sero_ci())) |>
  unnest(res)

# how many sera survive each age window, and how many regions clear n_min
sero_all |> summarise(n_sera = sum(n_sero), n_groups = sum(n_sero >= n_min),
                      .by = metric)

scat <- bind_rows(
  ser$cov_win |> mutate(metric = unname(metric_of[as.character(dose)]),
                        .keep = "unused"),
  ser$evc_win |> mutate(metric = metric_of[["EVC"]])
) |>
  mutate(region = as.character(region)) |>
  inner_join(sero_all, by = join_by(metric, region)) |>
  filter(n_sero >= n_min) |>
  mutate(metric = factor(metric, levels = metric_lvl),
         region = factor(region, levels = rank_order))

# the gap that dominates panel C: seroprevalence minus x
scat |> summarise(med_gap = median(sero - x), lo = min(sero - x),
                  hi = max(sero - x), .by = metric)

theme_epi <- theme_bw() +
  theme(
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background  = element_rect(fill = "white", colour = NA),
    panel.border     = element_rect(fill = NA, colour = "black"),
    panel.grid       = element_blank(),
    axis.line   = element_blank(),
    axis.ticks  = element_line(colour = "black"),
    axis.text   = element_text(colour = "black"),
    axis.text.x = element_text(vjust = 1),
    axis.title  = element_text(face = "bold", colour = "black"),
    strip.background = element_blank(),
    strip.text       = element_text(colour = "black"),
    legend.position   = "bottom",
    legend.title      = element_text(face = "bold", colour = "black"),
    legend.text       = element_text(colour = "black"),
    legend.key        = element_blank(),
    legend.background = element_blank()
  )

make_x_breaks <- function(rng, by = "3 months", min_gap_days = 70) {
  inner <- seq(floor_date(rng[1], "month"), rng[2], by = by)
  keep  <- as.numeric(inner - rng[1]) >= min_gap_days &
           as.numeric(rng[2] - inner)  >= min_gap_days
  sort(unique(c(rng[1], inner[keep], rng[2])))
}

x_breaks <- make_x_breaks(range(weeks))

# month on line 1, year on line 2 for the first break of each year
label_month_year <- function(x) {
  y   <- format(x, "%Y")
  out <- if_else(!duplicated(y), str_c(format(x, "%b"), "\n", y),
                 str_c(format(x, "%b"), "\n "))
  replace(out, is.na(x), NA_character_)
}

lines_region <- bind_rows(
  ser$cov |> transmute(week, region,
                       series = unname(metric_of[as.character(dose)]),
                       value = coverage),
  ser$evc |> transmute(week, region, series = metric_of[["EVC"]], value = evc)
) |>
  mutate(series = factor(series, levels = metric_lvl),
         region = factor(as.character(region), levels = rank_order))

rib_region <- ser$evc |>
  mutate(region = factor(as.character(region), levels = rank_order))

p_cov_evc <- ggplot(lines_region,
                    aes(week, value, colour = region, linetype = series,
                        group = interaction(region, series))) +
  geom_ribbon(data = rib_region, inherit.aes = FALSE,
              aes(week, ymin = evc_lo, ymax = evc_hi, fill = region,
                  group = region),
              alpha = 0.15, colour = NA, show.legend = FALSE) +
  geom_line() +
  scale_colour_manual(values = pal_region) +
  scale_fill_manual(values = pal_region) +
  scale_linetype_manual(values = deframe(select(metrics, metric, lty)),
                        labels = deframe(select(metrics, metric, short))) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_x_date(breaks = x_breaks, labels = label_month_year,
               expand = expansion(mult = 0.03)) +
  labs(x = NULL, y = "Coverage", colour = "Region", linetype = "Series") +
  theme_epi

# Horizontal bar = range over the window, NOT a confidence interval.
# Faded bar on the EVC panel = 95% VE-parameter interval on the period mean.
# Dotted diagonal = 1:1. It only enters a panel where the x and y ranges
# overlap, which they do not here, so the y limit is extended just enough for
# the line to cross the top-left corner.
diag_lim <- scat |>
  summarise(x  = min(c(x_lo, xu_lo), na.rm = TRUE),
            x2 = max(c(x_hi, xu_hi), na.rm = TRUE),
            y1 = min(sero_lo), y2 = max(sero_hi),
            .by = metric) |>
  mutate(span = x2 - x,
         lo = pmax(0, pmin(y1, x2 - diag_expand * span)),
         hi = pmin(1, pmax(y2, x + diag_expand * span))) |>
  select(metric, x, lo, hi) |>
  pivot_longer(c(lo, hi), values_to = "sero")

p_sero <- ggplot(scat, aes(x, sero, colour = region)) +
  (if (diag_expand > 0) geom_blank(data = diag_lim, inherit.aes = FALSE,
                                   aes(x = x, y = sero))) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", colour = "grey50") +
  geom_segment(aes(x = xu_lo, xend = xu_hi, y = sero, yend = sero),
               alpha = 0.35, na.rm = TRUE) +
  geom_segment(aes(x = x_lo, xend = x_hi, y = sero, yend = sero)) +
  geom_segment(aes(x = x, xend = x, y = sero_lo, yend = sero_hi)) +
  facet_wrap(~metric, nrow = 1, scales = "free") +
  scale_colour_manual(values = pal_region) +
  scale_x_continuous(labels = label_percent(accuracy = 1)) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  guides(colour = "none") +          # Region legend collected from panel A
  labs(x = NULL, y = "Seroprevalence") +
  theme_epi

province_south_raw <- c("Bình Dương", "Bình Phước", "Tây Ninh",
                        "Bà Rịa Vũng Tàu", "Đồng Nai", "Long An", "Tiền Giang",
                        "Bến Tre", "Vĩnh Long", "Trà Vinh", "Đồng Tháp",
                        "Hậu Giang", "An Giang", "Kiên Giang", "Bạc Liêu",
                        "Sóc Trăng", "Cà Mau", "Hồ Chí Minh", "Cần Thơ")

# same cleaning as applied to the dataframe
province_south <- province_south_raw |>
  stri_trans_general("Latin-ASCII") |>
  str_to_lower() |>
  str_squish() |>
  str_replace_all("đ", "d")

map_south <- df_map_vn |>
  left_join(df_vax_vn, by = "province_blank") |>
  filter(province_blank %in% province_south)

# CHECK: provinces failing the name match drop silently and draw grey
setdiff(province_south, map_south$province_blank)
map_south |> st_drop_geometry() |> filter(is.na(measles_pct)) |> pull(province_blank)

hcmc_poly <- filter(map_south, province_blank == "ho chi minh")

p_map_south <- ggplot() +
  geom_sf(data = map_south, aes(fill = measles_pct), colour = "white",
          linewidth = 0.2) +
  geom_sf(data = hcmc_poly, fill = NA, colour = "white", linewidth = 1.5) +
  geom_sf(data = hcmc_poly, fill = NA, colour = "grey10", linewidth = 0.6) +
  scale_fill_gsea(reverse = TRUE,
                  labels = percent_format(scale = 1, accuracy = 1),
                  na.value = "grey90",
                  breaks = pretty_breaks(4),
                  guide = guide_colourbar(barwidth = unit(3.5, "cm"),
                                          barheight = unit(0.3, "cm"),
                                          ticks.colour = "white",
                                          frame.colour = NA)) +
  labs(fill = NULL) +
  theme_void(base_size = 11) +
  theme(legend.position      = "top",
        legend.direction     = "horizontal",
        legend.justification = "center",
        legend.box.spacing   = unit(3, "pt"),
        legend.margin        = margin(0, 0, 0, 0),
        legend.text          = element_text(size = 8, colour = "black"),
        plot.margin          = margin(2, 2, 2, 2))

# Key map for the legend: same lookup and same palette as panel A, so the
# colours cannot drift apart. Area is not population: Can Gio is a third of
# the city's land and a small share of its children.
map_region <- df_map_hcmc |> mutate(region = assign_region(district_blank))

# CHECK: districts with no region draw grey here and are absent from A and C
map_region |> st_drop_geometry() |> filter(is.na(region)) |> pull(district_blank)

p_region_map <- ggplot(map_region) +
  geom_sf(aes(fill = region), colour = "white", linewidth = 0.1) +
  scale_fill_manual(values = pal_region, na.value = "grey85", guide = "none") +
  coord_sf(expand = FALSE) +
  theme_void() +
  theme(plot.margin = margin(2, 2, 4, 20))

# No guide collection anywhere. The two legends are extracted from panel A as
# separate objects so the key map can sit between them, and each goes in its
# own grid cell where patchwork cannot reposition it.
p_leg_base <- p_cov_evc +
  theme(legend.position  = "right",
        legend.direction = "vertical",
        legend.margin    = margin(0, 0, 0, 0),
        legend.title     = element_text(size = 10, face = "bold"),
        legend.text      = element_text(size = 9),
        legend.key.width = unit(1.2, "lines"))

# If either comes out blank, cowplot::get_legend() has hit the ggplot2 3.5
# guide change: use get_plot_component(p, "guide-box-right", return_all
# = TRUE)[[1]] instead
leg_region <- cowplot::get_legend(
  p_leg_base + guides(linetype = "none",
                      colour = guide_legend(ncol = 1,
                                            override.aes = list(linetype = "solid")))
)

leg_series <- cowplot::get_legend(
  p_leg_base + guides(colour = "none",
                      linetype = guide_legend(ncol = 1,
                                              override.aes = list(colour = "grey30")))
)

# Each legend cell takes the legend's own height, so the map gets the rest
leg_h <- function(g) {
  h <- tryCatch(convertHeight(sum(g$heights), "cm", valueOnly = TRUE),
                error = \(e) NA_real_)
  if (is.na(h)) 1.5 else h + 0.15
}

as_cell <- function(g) {
  wrap_elements(full = grid::grobTree(
    g, vp = grid::viewport(y = 1, just = "top", height = grid::unit(1, "npc"))
  )) + theme(plot.margin = margin(0, 0, 0, 20))
}

p_key <- wrap_plots(
  as_cell(leg_series),
  as_cell(leg_region),
  p_region_map,
  ncol = 1,
  heights = unit(c(leg_h(leg_series), leg_h(leg_region), 1),
                 c("cm", "cm", "null"))
)

# B is wrapped so patchwork does not align A's panel top with B's panel top,
# which is what left the gap above A when B's colourbar sat on top
design <- "
AABL
CCCL
"

p_all <- (p_cov_evc + theme(legend.position = "none")) +
  wrap_elements(full = p_map_south) +
  p_sero +
  wrap_elements(full = p_key) +
  plot_layout(design = design, widths = c(1, 1, 1.05, 0.7),
              heights = c(1, 1)) +
  plot_annotation(tag_levels = list(c("A", "B", "C", ""))) &
  theme(plot.tag = element_text(face = "bold"))

p_all

ggsave("figs/chap1/fig_cov_evc_sero.png", p_all, width = 8, height = 4.7, dpi = 300)
```

## Values

```{r}
# ===========================================================================
# SUMMARY VALUES FOR THE PANEL A DESCRIPTION
# ===========================================================================
 
pp  <- \(x) sprintf("%.1f", 100 * x)      # proportion as a percentage
ppd <- \(x) sprintf("%+.1f", 100 * x)     # change in percentage points
 
wk_first <- min(lines_region$week)
wk_last  <- max(lines_region$week)
 
c(as.character(wk_first), as.character(wk_last))   # the period being described
 
# ---- 1. Start, end, and change for every region and series ----------------
 
ends <- lines_region |>
  filter(week %in% c(wk_first, wk_last)) |>
  mutate(point = if_else(week == wk_first, "start", "end")) |>
  select(series, region, point, value) |>
  pivot_wider(names_from = point, values_from = value) |>
  mutate(change_pp = 100 * (end - start)) |>
  arrange(series, desc(end))
 
ends |> print(n = Inf)
 
# ---- 2. Leading and trailing region, and the spread between them ----------
 
extremes <- lines_region |>
  filter(week %in% c(wk_first, wk_last)) |>
  mutate(point = factor(if_else(week == wk_first, "Start", "End"),
                        levels = c("Start", "End"))) |>
  summarise(
    highest   = as.character(region[which.max(value)]),
    high_val  = max(value),
    lowest    = as.character(region[which.min(value)]),
    low_val   = min(value),
    spread_pp = 100 * (max(value) - min(value)),
    .by = c(series, point)
  ) |>
  arrange(series, point)
 
extremes |> print(n = Inf)
 
# did the ranking of regions change between start and end?
lines_region |>
  filter(week %in% c(wk_first, wk_last)) |>
  mutate(point = if_else(week == wk_first, "start", "end"),
         rk    = rank(-value), .by = c(series, week)) |>
  select(series, region, point, rk) |>
  pivot_wider(names_from = point, values_from = rk) |>
  filter(start != end) |>
  print(n = Inf)
 
# ---- 3. EVC with its uncertainty interval --------------------------------
 
evc_ci <- rib_region |>
  filter(week %in% c(wk_first, wk_last)) |>
  left_join(lines_region |>
              filter(series == x_labels[["EVC"]]) |>
              select(week, group, evc = value),
            by = c("week", "group")) |>
  mutate(point    = factor(if_else(week == wk_first, "Start", "End"),
                           levels = c("Start", "End")),
         width_pp = 100 * (evc_hi - evc_lo),
         lab      = sprintf("%s%% (%s to %s%%)", pp(evc), pp(evc_lo), pp(evc_hi))) |>
  select(region, point, evc, evc_lo, evc_hi, width_pp, lab) |>
  arrange(point, desc(evc))
 
evc_ci |> print(n = Inf)
 
# band width across the whole period: all regions share the same draws, so
# this is VE-parameter uncertainty, not between-region precision
rib_region |>
  summarise(width_min = 100 * min(evc_hi - evc_lo),
            width_med = 100 * median(evc_hi - evc_lo),
            width_max = 100 * max(evc_hi - evc_lo))
 
# ---- 4. Gap between MCV1 coverage and estimated immunity -----------------
# The vertical distance between the two lines is primary vaccine failure
# plus waning, expressed in percentage points.
 
gap <- lines_region |>
  filter(series %in% unname(x_labels[c("Dose 1", "EVC")])) |>
  select(week, region, series, value) |>
  pivot_wider(names_from = series, values_from = value) |>
  rename(mcv1 = !!x_labels[["Dose 1"]], evc = !!x_labels[["EVC"]]) |>
  mutate(gap_pp = 100 * (mcv1 - evc))
 
gap |>
  summarise(gap_start = gap_pp[which.min(week)],
            gap_end   = gap_pp[which.max(week)],
            gap_min   = min(gap_pp),
            gap_max   = max(gap_pp),
            .by = region) |>
  print(n = Inf)
 
# ---- 5. Largest week-on-week step, e.g. the early-2024 jump --------------
 
lines_region |>
  arrange(region, series, week) |>
  mutate(step_pp = 100 * (value - lag(value)), .by = c(region, series)) |>
  slice_max(step_pp, n = 1, by = c(region, series)) |>
  select(series, region, week, step_pp) |>
  arrange(series, desc(step_pp)) |>
  print(n = Inf)
 
# ---- 6. Sentence-ready strings -------------------------------------------
 
ends |>
  mutate(txt = sprintf("%s, %s: %s%% to %s%% (%s pp)",
                       series, region, pp(start), pp(end), ppd(end - start))) |>
  pull(txt) |>
  writeLines()
 
evc_ci |>
  filter(point == "End") |>
  mutate(txt = sprintf("%s: EVC %s", region, lab)) |>
  pull(txt) |>
  writeLines()
```

## Values sero vs cov

```{r}
# ===========================================================================
# SUMMARY VALUES FOR THE PANEL C DESCRIPTION
# ===========================================================================
 
pp <- \(x) sprintf("%.1f", 100 * x)
 
# ---- 1. Seroprevalence by region, with sample sizes -----------------------
# Same sera for MCV1 and EVC (9-59 mo); MCV2 uses 18-59 mo, so its
# seroprevalence and n differ.
 
scat |>
  select(metric, region, x_pos, n_sero, sero, sero_lo, sero_hi) |>
  mutate(lab = sprintf("%s%% (%s to %s%%), %d/%d",
                       pp(sero), pp(sero_lo), pp(sero_hi), x_pos, n_sero)) |>
  arrange(metric, desc(sero)) |>
  print(n = Inf)
 
# range across regions, per age window
scat |>
  summarise(n_regions = n(),
            n_sera    = sum(n_sero),
            sero_min  = min(sero), sero_max = max(sero),
            reg_min   = as.character(region[which.min(sero)]),
            reg_max   = as.character(region[which.max(sero)]),
            .by = metric) |>
  print(n = Inf)
 
# widest and narrowest confidence intervals, which drive how much the
# scatter can be read at all
scat |>
  mutate(ci_pp = 100 * (sero_hi - sero_lo)) |>
  slice_max(ci_pp, n = 1, by = metric) |>
  select(metric, region, n_sero, sero, ci_pp) |>
  print(n = Inf)
 
# ---- 2. The x metrics: period value and temporal range -------------------
 
scat |>
  mutate(lab = sprintf("%s%% (range %s to %s%%)", pp(x), pp(x_lo), pp(x_hi))) |>
  select(metric, region, x, x_lo, x_hi, lab) |>
  arrange(metric, desc(x)) |>
  print(n = Inf)
 
scat |>
  summarise(x_min = min(x), x_max = max(x),
            spread_pp = 100 * (max(x) - min(x)),
            reg_min = as.character(region[which.min(x)]),
            reg_max = as.character(region[which.max(x)]),
            .by = metric) |>
  print(n = Inf)
 
# EVC parameter interval, which only the EVC panel carries
scat |>
  filter(!is.na(xu_lo)) |>
  mutate(u_pp = 100 * (xu_hi - xu_lo)) |>
  select(metric, region, x, xu_lo, xu_hi, u_pp) |>
  print(n = Inf)
 
# ---- 3. The gap: how far every point sits below the 1:1 line -------------
# This is the dominant feature of panel C and the number to lead with.
 
scat |>
  mutate(gap_pp = 100 * (x - sero)) |>
  summarise(gap_med = median(gap_pp),
            gap_min = min(gap_pp), gap_max = max(gap_pp),
            n_above = sum(gap_pp < 0),          # regions above the 1:1 line
            .by = metric) |>
  print(n = Inf)
 
# per region, for the metric-specific sentences
scat |>
  mutate(gap_pp = 100 * (x - sero)) |>
  select(metric, region, x, sero, gap_pp) |>
  arrange(metric, desc(gap_pp)) |>
  print(n = Inf)
 
# does any region's seroprevalence CI reach its x value?
scat |>
  filter(sero_hi >= x) |>
  select(metric, region, x, sero, sero_hi) |>
  print(n = Inf)
 
# ---- 4. Do the two axes rank the regions the same way? -------------------
# Ranking agreement is the descriptive version of the correlation, and is
# more honest than a p-value at n = 5.
 
scat |>
  mutate(rk_x = rank(-x), rk_y = rank(-sero), .by = metric) |>
  select(metric, region, rk_x, rk_y) |>
  arrange(metric, rk_x) |>
  print(n = Inf)
 
scat |>
  summarise(rho = cor(x, sero, method = "spearman"),
            n_concordant = sum(rank(-x) == rank(-sero)),
            .by = metric) |>
  print(n = Inf)
 
# ---- 5. Sentence-ready strings -------------------------------------------
 
scat |>
  filter(metric == x_labels[["EVC"]]) |>
  mutate(txt = sprintf("%s: EVC %s%%, seroprevalence %s%% (%s to %s%%), gap %.1f pp",
                       region, pp(x), pp(sero), pp(sero_lo), pp(sero_hi),
                       100 * (x - sero))) |>
  arrange(desc(x)) |>
  pull(txt) |>
  writeLines()
```


## Sero vs cum inc

```{r}
sero_age_lo <- 9 / 12      # years; matches the EVC denominator
sero_age_hi <- 5
snap_week   <- max(weeks)  # last study week; coverage and EVC snapshot
n_min_round <- 5           # minimum sera per region within the last round
common_only <- TRUE

bin_all  <- as.Date(c("2024-07-01", "2024-10-01", "2025-01-01",
                      "2025-04-01", "2025-07-01", "2025-10-01"))
# The last four quarters give the same picture (check below), so only the
# final one is plotted
bin_keep <- as.Date(c("2024-07-01", "2024-10-01", "2025-10-01"))

inc_src <- snapshot_dist |>
  filter(bin %in% bin_all) |>
  mutate(region = as.character(assign_region(district)))

# CHECK: districts with no region drop out of every panel. Should be empty.
inc_src |> filter(is.na(region)) |> distinct(district)

# cum_n counts cases to the end of each quarter, so quarters are labelled by
# bin_end, as in panel A of the outbreak figure
inc_all <- inc_src |>
  summarise(cum_n = sum(cum_n), n_census = sum(n_census),
            .by = c(region, bin, bin_end)) |>
  mutate(
    inc    = cum_n / n_census * 1e5,
    inc_lo = qchisq(0.025, 2 * cum_n) / 2 / n_census * 1e5,
    inc_hi = qchisq(0.975, 2 * (cum_n + 1)) / 2 / n_census * 1e5
  )

# CHECK: with one x scale per quarter, a panel looks the same as the next when
# the regions keep their rank and relative position (pos). Same ranks and
# similar pos across the last four quarters = safe to merge them.
inc_all |>
  mutate(rank = rank(-inc),
         pos  = round((inc - min(inc)) / diff(range(inc)), 2),
         .by = bin_end) |>
  select(region, bin_end, rank, pos) |>
  pivot_wider(names_from = bin_end, values_from = c(rank, pos))

# verify the ordering first
sero_src |> count(hospital, period_collect)

# last collection period, defined within each hospital
sero_last <- sero_src |>
  filter(period_collect == max(period_collect), .by = hospital)

sero_last |> summarise(period = first(period_collect), n = n(), .by = hospital)

sero_round <- sero_last |>
  filter(exact_age >= sero_age_lo, exact_age < sero_age_hi) |>
  sero_ci()

sero_round

sero_snap <- sero_round |> filter(n_sero >= n_min_round)

metric_lev <- c(metric_lvl, "Seroprevalence")

x_all <- bind_rows(
  ser$cov |>
    filter(week == snap_week) |>
    transmute(region = as.character(region),
              metric = unname(metric_of[as.character(dose)]), x = coverage),
  ser$evc |>
    filter(week == snap_week) |>
    transmute(region = as.character(region), metric = metric_of[["EVC"]],
              x = evc, x_lo = evc_lo, x_hi = evc_hi),
  sero_snap |>
    transmute(region, metric = "Seroprevalence",
              x = sero, x_lo = sero_lo, x_hi = sero_hi)
)

if (common_only) {
  x_all <- x_all |> filter(n_distinct(metric) == length(metric_lev), .by = region)
}

# Every metric is paired with every period within a region, on purpose
dat_inc <- x_all |>
  inner_join(filter(inc_all, bin %in% bin_keep), by = "region",
             relationship = "many-to-many") |>
  mutate(metric = factor(metric, levels = metric_lev),
         region = factor(region, levels = rank_order),
         period = factor(format(bin_end, "%b %Y"),
                         levels = format(sort(unique(bin_end)), "%b %Y")))

dat_inc |> count(metric, period)

# n = 5 per period, and the periods are nested cumulative counts, so these
# are not independent tests. cor.test() gives the exact p when there are no
# ties; the normal approximation (exact = FALSE) is poor at n = 5.
cor_inc <- dat_inc |>
  summarise(
    n   = sum(is.finite(x) & is.finite(inc)),
    rho = if (n >= 3) cor(x, inc, method = "spearman",
                          use = "complete.obs") else NA_real_,
    p   = if (n >= 3) suppressWarnings(
            cor.test(x, inc, method = "spearman")$p.value
          ) else NA_real_,
    .by = c(metric, period)
  )

cor_inc

# For the figure legend: Spearman rho (p) by metric and period
cor_inc |>
  mutate(val = sprintf("%.2f (p = %.2f)", rho, p)) |>
  select(metric, period, val) |>
  pivot_wider(names_from = period, values_from = val)

# rho label centred above each period's group of points
cor_lab <- dat_inc |>
  summarise(inc = (min(inc) + max(inc)) / 2, .by = c(metric, period)) |>
  left_join(cor_inc, by = join_by(metric, period)) |>
  mutate(lab = sprintf("rho = %.2f", rho))

# Fractional logit (quasibinomial GLM, logit link): the fitted curve stays
# inside 0-1 and bends near the bounds. Two parameters, like a straight line.
# Unlike Theil-Sen it is not robust to one outlying region.
fit_frac <- function(inc, x) {
  ok <- is.finite(inc) & is.finite(x)
  if (sum(ok) < 3) return(NULL)
  b    <- unname(coef(glm(x[ok] ~ inc[ok], family = quasibinomial)))
  grid <- seq(min(inc[ok]), max(inc[ok]), length.out = 100)
  tibble(inc = grid, x = plogis(b[1] + b[2] * grid))
}

fit_curve <- dat_inc |> reframe(fit_frac(inc, x), .by = c(metric, period))

# One panel per metric, x shared across panels. Colour = region, shape =
# period. Horizontal bars: exact 95% Poisson CI for incidence. Vertical bars:
# 95% VE-parameter interval (EVC) or exact 95% CI (seroprevalence); coverage
# is a registry census and carries no bar. Dashed curves: fractional logit fit
# per period.
p_inc_all <- ggplot(dat_inc, aes(inc, x, colour = region)) +
  geom_segment(aes(x = inc_lo, xend = inc_hi, y = x, yend = x)) +
  geom_segment(aes(x = inc, xend = inc, y = x_lo, yend = x_hi), na.rm = TRUE) +
  geom_point(aes(shape = period)) +
  # group = period keeps the three curves in a panel from joining up
  geom_line(data = fit_curve, aes(inc, x, group = period), inherit.aes = FALSE,
            linetype = "dashed", colour = "grey40", linewidth = 0.4) +
  geom_text(data = cor_lab, aes(inc, y = Inf, label = lab),
            inherit.aes = FALSE, colour = "black", vjust = 1.3) +
  facet_grid(metric ~ ., scales = "free_y", switch = "y") +
  scale_colour_manual(values = pal_region) +
  # space above the points for the rho labels
  scale_y_continuous(labels = label_percent(),
                     expand = expansion(mult = c(0.05, 0.25))) +
  labs(x = "Cumulative incidence per 100,000", y = NULL,
       colour = "Region", shape = "Cases up to") +
  theme_epi +
  theme(strip.placement = "outside", legend.position = "right")

p_inc_all

# ggsave("figs/chap1/fig_immunity_vs_incidence.png", p_inc_all,
#        width = 7.5, height = 8, dpi = 300)
```


## Under5: Sero vs cum

```{r}
# ===========================================================================
# 14. UNDER-5 CUMULATIVE INCIDENCE BY REGION
# ===========================================================================
# The all-age version compares child immunity against an outbreak measured in
# the whole population. Restricting both numerator and denominator to young
# children makes the incidence measure match what the immunity measures
# describe.
 
case_age_lo <- 0     # years, inclusive
case_age_hi <- 5     # years, exclusive
 
# Setting case_age_lo <- 1 excludes infants, who are not yet MCV1-eligible and
# so sit outside the coverage and EVC denominators (9-59 months). That is the
# cleaner age match; 0-5 keeps the group epidemiologically complete.
 
dedupe_cases <- TRUE
 
# ---- duplicate records ----------------------------------------------------
 
df_ep2 |> summarise(rows = n(), unique_ids = n_distinct(id))
df_ep2 |> count(id) |> filter(n > 1) |> nrow()
 
# ---- census denominator ---------------------------------------------------
# Census age band k covers [k-1, k), so under 5 is bands 1:5, not 1:4.
 
census_u5 <- df_census |>
  filter(age >= case_age_lo + 1, age <= case_age_hi) |>
  summarise(n = sum(n), .by = district) |>
  mutate(district = merge_thuduc(clean_district(district)),
         region   = assign_region(district)) |>
  summarise(n_census = sum(n), .by = region) |>
  drop_na(region)
 
census_u5
 
# sanity: under-5s should be a few percent of the all-age total
tibble(u5 = sum(census_u5$n_census), all_age = sum(df_census$n)) |>
  mutate(pct = 100 * u5 / all_age)
 
# ---- case numerator -------------------------------------------------------
 
cases_u5 <- df_ep2 |>
  (\(d) if (dedupe_cases) distinct(d, id, .keep_all = TRUE) else d)() |>
  filter(age >= case_age_lo, age < case_age_hi) |>
  mutate(district = merge_thuduc(.data[[col_district]]),
         region   = assign_region(district),
         date     = .data[[col_date]]) |>
  drop_na(region, date)
 
cases_u5 |> count(region)
cat("under-5 cases:", nrow(cases_u5), "of", nrow(df_ep2), "rows\n")
 
# ---- quarterly cumulative incidence ---------------------------------------
# Bins built from the full date range so the under-5 and all-age versions
# share a bin grid; regions with no cases in a quarter still get a row.
 
date_rng <- range(df_ep2[[col_date]], na.rm = TRUE)
bin_grid <- seq(floor_date(date_rng[1], "quarter"),
                floor_date(date_rng[2], "quarter"), by = "quarter")
 
inc_u5 <- cases_u5 |>
  mutate(bin = floor_date(date, "quarter")) |>
  count(region, bin) |>
  complete(region = census_u5$region, bin = bin_grid, fill = list(n = 0)) |>
  arrange(region, bin) |>
  mutate(cum_n = cumsum(n), .by = region) |>
  left_join(census_u5, by = "region") |>
  mutate(
    group  = as.character(region),
    inc    = cum_n / n_census * 1e5,
    inc_lo = qchisq(0.025, 2 * cum_n) / 2 / n_census * 1e5,
    inc_hi = qchisq(0.975, 2 * (cum_n + 1)) / 2 / n_census * 1e5
  )
 
inc_u5 |> select(group, bin, cum_n, n_census, inc) |> arrange(bin) |> print(n = Inf)
 
# ===========================================================================
# 15. IMMUNITY MEASURES VS UNDER-5 INCIDENCE
# ===========================================================================
# Reuses x_all from section 13 (coverage and EVC at snap_week, seroprevalence
# from each hospital's last collection period).
 
bin_keep <- as.Date(c("2024-07-01", "2024-10-01", "2025-01-01",
                      "2025-04-01", "2025-07-01", "2025-10-01"))
bin_lab  <- setNames(format(bin_keep, "%b %Y"), as.character(bin_keep))
 
dat_u5 <- x_all |>
  inner_join(inc_u5 |> filter(bin %in% bin_keep) |>
               select(group, bin, region, inc, inc_lo, inc_hi),
             by = "group") |>
  mutate(metric = factor(metric, levels = metric_lev),
         region = factor(as.character(region), levels = rank_order),
         bin    = factor(bin_lab[as.character(bin)], levels = unname(bin_lab)))
 
dat_u5 |> summarise(n = n(), .by = c(metric, bin)) |> print(n = Inf)
 
# ---- Spearman rho ---------------------------------------------------------
# Five regions per panel, and the six bins are nested cumulative counts, so
# read each row as a trajectory rather than as six independent tests.
 
cor_u5 <- dat_u5 |>
  summarise(
    n   = sum(is.finite(x) & is.finite(inc)),
    rho = if (n[1] >= 3) cor(x, inc, method = "spearman",
                             use = "complete.obs") else NA_real_,
    p   = if (n[1] >= 3) suppressWarnings(
            cor.test(x, inc, method = "spearman", exact = FALSE)$p.value
          ) else NA_real_,
    .by = c(metric, bin)
  ) |>
  mutate(lab = if_else(is.na(rho), "rho = NA",
                       sprintf("rho = %.2f\np = %.3f", rho, p)))
 
cor_u5 |> print(n = Inf)
 
# side by side with the all-age version, if section 13 was run
if (exists("cor_inc_all")) {
  cor_inc_all |>
    select(metric, bin, rho_all = rho) |>
    left_join(cor_u5 |> select(metric, bin, rho_u5 = rho), by = c("metric", "bin")) |>
    print(n = Inf)
}
 
# ---- Theil-Sen line, consistent with a rank correlation -------------------
 
theil_sen <- function(x, y) {
  ok <- is.finite(x) & is.finite(y)
  x <- x[ok]; y <- y[ok]
  if (length(x) < 3) return(NULL)
  cb   <- combn(length(x), 2)
  dx   <- x[cb[2, ]] - x[cb[1, ]]
  dy   <- y[cb[2, ]] - y[cb[1, ]]
  keep <- dx != 0
  if (!any(keep)) return(NULL)
  b <- median(dy[keep] / dx[keep])
  tibble(a = median(y - b * x), b = b, x1 = min(x), x2 = max(x))
}
 
fit_u5 <- dat_u5 |>
  reframe(theil_sen(inc, x), .by = c(metric, bin)) |>
  transmute(metric, bin, inc = x1, inc_end = x2,
            x = a + b * x1, x_end = a + b * x2)
 
# ---- figure ---------------------------------------------------------------
# Horizontal bars: exact 95% Poisson CI for the incidence rate.
# Vertical bars: 95% VE-parameter interval (EVC) or exact 95% CI
# (seroprevalence). Coverage is a registry census and carries no bar.
# Dashed line: Theil-Sen fit, an aid to reading direction, not a model.
 
p_u5 <- ggplot(dat_u5, aes(inc, x, colour = region)) +
  geom_segment(aes(x = inc_lo, xend = inc_hi, y = x, yend = x)) +
  geom_segment(aes(x = inc, xend = inc, y = x_lo, yend = x_hi), na.rm = TRUE) +
  geom_segment(data = fit_u5, inherit.aes = FALSE,
               aes(x = inc, xend = inc_end, y = x, yend = x_end),
               linetype = "dashed", colour = "grey40", linewidth = 0.4) +
  geom_point(size = 1.2) +
  geom_text(data = cor_u5, aes(x = -Inf, y = Inf, label = lab),
            inherit.aes = FALSE, colour = "black", hjust = -0.08, vjust = 1.2,
            lineheight = 0.9) +
  facet_grid(metric ~ bin, scales = "free", switch = "y") +
  scale_colour_manual(values = pal_region) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  labs(x = "Cumulative incidence per 100,000 children under 5",
       y = NULL, colour = "Region") +
  theme_epi() +
  theme(strip.placement = "inside")
 
p_u5
 
ggsave("figs/chap1/fig_immunity_vs_incidence_u5.png", p_u5,
       width = 9.5, height = 6.5, dpi = 300)
```



## Case negative coverage

```{r}
df_case_neg <- readRDS(file.path(path2data, cases_neg_file))

df_case_neg <- df_case_neg |> 
  filter(ngay_nv >= ep2$start, ngay_nv <= ep2$end)

df_summary <- df_case_neg |>
  mutate(
    age_group = cut(
      age,
      breaks = c(-Inf, 9 / 12, 5, 15, Inf),
      labels = c("<9m", "9m-5", "5-15", "15+")
    ),
    vax = factor(tiem_chung,
                 levels = c("Không", "Có", "Không rõ"),
                 labels = c("No", "Yes", "Unknown"))
  ) |>
  count(age_group, vax) |>
  pivot_wider(names_from = vax, values_from = n, values_fill = 0) |>
  mutate(
    n_total = No + Yes + Unknown,
    # complete case: unknown excluded from numerator and denominator
    pct_cc  = 100 * Yes / (Yes + No),
    # lower bound: unknown counted as unvaccinated
    pct_lo  = 100 * Yes / n_total,
    # upper bound: unknown counted as vaccinated
    pct_hi  = 100 * (Yes + Unknown) / n_total
  ) |>
  transmute(
    age_group,
    No      = sprintf("%d", No),
    Yes     = sprintf("%d", Yes),
    Unknown = sprintf("%d", Unknown),
    n_total,
    `Complete case`  = sprintf("%.1f%%", pct_cc),
    `Unknown as no`  = sprintf("%.1f%%", pct_lo),
    `Unknown as yes` = sprintf("%.1f%%", pct_hi)
  )
 
df_summary

# write.csv(df_summary, "figs/chap1/vax_cases_2425.csv", quote = F, row.names = F)
```

## Catchment vs cases address

```{r}
# ===========================================================================
# 16. SEROSURVEY CATCHMENT VS HOSPITAL ADMISSION CATCHMENT
# ===========================================================================
# Three maps per hospital: the district composition of collected serum
# samples, and the district composition of measles admissions in each
# outbreak period, plus a direct scatter of the two shares.
 
col_date <- "ngay_nv"     # admission date column in df_case
 
# ---- 16a. Hospital name normalisation -------------------------------------
# "Nhi đồng 1/2" anchored at end of string so "Nhi Đồng Đồng Nai" and
# "Nhi đồng Tp.Cần Thơ" cannot match. Diacritic case varies, so ignore_case.
 
hosp_map <- df_case |>
  as_tibble() |>
  mutate(
    hosp = case_when(
      str_detect(ten_bv, regex("nhi\\s*đồng\\s*1\\s*$", ignore_case = TRUE)) ~ "CH1",
      str_detect(ten_bv, regex("nhi\\s*đồng\\s*2\\s*$", ignore_case = TRUE)) ~ "CH2",
      str_detect(ten_bv, regex("nhi\\s*đồng\\s*thành\\s*phố\\s*$",
                               ignore_case = TRUE))                          ~ "CCH",
      .default = NA_character_
    )
  )
 
# CHECK: which names mapped where, and the totals
hosp_map |> count(hosp, ten_bv) |> filter(!is.na(hosp)) |> print(n = Inf)
hosp_map |> count(hosp)      # expect CH1 4422, CH2 3415, CCH 1185
 
df_case_3h <- hosp_map |> filter(!is.na(hosp))
 
# ---- 16b. Percentages from all three sources ------------------------------
 
ep_win <- tibble(
  panel = c("2018-2020", "2024-2025"),
  start = c(ep1$start, ep2$start),
  end   = c(ep1$end,   ep2$end)
)
 
ep_win      # confirm the labels match the actual episode date ranges
 
adm_pct <- ep_win |>
  reframe(
    df_case_3h |>
      filter(between(.data[[col_date]], start, end)) |>
      drop_na(district) |>
      count(hosp, district),
    .by = panel
  ) |>
  mutate(pct = 100 * n / sum(n), .by = c(panel, hosp)) |>
  transmute(panel, hospital = hosp, district, pct)
 
adm_pct |> summarise(n_dist = n(), total = sum(pct), .by = c(panel, hospital))
 
# CHECK: district strings must line up across sources
setdiff(unique(adm_pct$district), unique(df_map_pct$district))
setdiff(unique(df_map_pct$district), unique(adm_pct$district))
setdiff(unique(adm_pct$district), unique(df_map_hcmc$district_blank))
 
panel_lev <- c("Serum samples", "2018-2020", "2024-2025")
 
pct_all <- bind_rows(
  df_map_pct |> as_tibble() |>
    transmute(panel = "Serum samples", hospital = as.character(hospital),
              district, pct),
  adm_pct
) |>
  mutate(panel    = factor(panel, levels = panel_lev),
         hospital = factor(hospital, levels = levels(df_map_pct$hospital)))
 
# ---- 16c. Panel A: maps ---------------------------------------------------
# log2 of each district's share against the equal-share reference
# (100 / n districts), so white marks proportional representation, red above,
# blue below. Districts below pct_min draw grey. One shared scale.
 
pct_min <- 1                                             # below this % draws grey
mid     <- 100 / n_distinct(df_map_hcmc$district_blank)  # equal-share reference
 
map_pct <- pct_all |>
  complete(nesting(panel, hospital), district = df_map_hcmc$district_blank,
           fill = list(pct = 0)) |>
  mutate(lfc = if_else(pct < pct_min, NA_real_, log2(pct / mid))) |>
  left_join(select(df_map_hcmc, district_blank),
            by = c("district" = "district_blank")) |>
  st_as_sf()
 
lo <- log2(pct_min / mid)                 # bar starts exactly at pct_min
hi <- max(map_pct$lfc, na.rm = TRUE)
 
# four breaks only: six labels collide on a bar this width
brks_pct <- c(1, 5, 20, 40)
brks_pct <- brks_pct[log2(brks_pct / mid) >= lo & log2(brks_pct / mid) <= hi]
 
p_catch <- map_pct |>
  ggplot() +
  geom_sf(aes(fill = lfc), colour = "grey60", linewidth = 0.25) +
  facet_grid(hospital ~ panel, switch = "y") +
  coord_sf(expand = FALSE) +
  scale_fill_gradientn(
    colours  = pal_gsea()(12),
    limits   = c(lo, hi),
    na.value = "grey92",
    rescaler = \(x, to = c(0, 1), from = range(x, na.rm = TRUE))
                 scales::rescale_mid(x, to, from, mid = 0),
    breaks   = log2(brks_pct / mid), labels = paste0(brks_pct, "%"),
    name     = "Share of samples or admissions",
    guide    = guide_colourbar(title.position = "top",
                               barwidth  = unit(5, "cm"),
                               barheight = unit(0.3, "cm"),
                               ticks.colour = "white",
                               frame.colour = NA)
  ) +
  theme_void(base_size = 11) +
  theme(
    strip.text.x       = element_text(colour = "grey15", size = 11,
                                      margin = margin(b = 5)),
    strip.text.y.left  = element_text(colour = "grey15", size = 11, angle = 0,
                                      margin = margin(r = 6)),
    strip.placement    = "outside",
    strip.clip         = "off",
    legend.position    = "bottom",
    legend.title       = element_text(colour = "grey15", hjust = 0.5, size = 10),
    legend.text        = element_text(colour = "grey30", size = 9),
    legend.box.spacing = unit(2, "pt"),
panel.spacing.x    = unit(0.3, "lines"),
    panel.spacing.y    = unit(0.1, "lines"),
    plot.margin        = margin(4, 4, 2, 4)
  )
 
p_catch
 
# ---- 16d. Summary values for the description ------------------------------
 
n_dist  <- n_distinct(df_map_hcmc$district_blank)
tab_pct <- map_pct |> st_drop_geometry()
 
# sample size behind each panel
bind_rows(
  df_map_pct |> as_tibble() |>
    summarise(n = sum(n), .by = hospital) |> mutate(panel = "Serum samples"),
  ep_win |>
    reframe(df_case_3h |>
              filter(between(.data[[col_date]], start, end)) |>
              drop_na(district) |>
              count(hosp, name = "n") |> rename(hospital = hosp),
            .by = panel)
) |>
  pivot_wider(names_from = panel, values_from = n) |>
  print()
 
# concentration: how few districts make up most of each catchment
tab_pct |>
  summarise(
    top1     = max(pct),
    top1_d   = district[which.max(pct)],
    top3     = sum(sort(pct, decreasing = TRUE)[1:3]),
    n_to_80  = which(cumsum(sort(pct, decreasing = TRUE)) >= 80)[1],
    n_over_1 = sum(pct >= 1),
    .by = c(hospital, panel)
  ) |>
  arrange(hospital, panel) |>
  print(n = Inf)
 
# agreement: Spearman rho and total variation distance (half the sum of
# absolute share differences, so 0 = identical, 100 = disjoint)
wide <- tab_pct |>
  select(hospital, panel, district, pct) |>
  pivot_wider(names_from = panel, values_from = pct, values_fill = 0)
 
agree <- wide |>
  summarise(
    n       = n(),
    rho_ep1 = cor(`Serum samples`, `2018-2020`, method = "spearman"),
    p_ep1   = suppressWarnings(
                cor.test(`Serum samples`, `2018-2020`,
                         method = "spearman", exact = FALSE)$p.value),
    rho_ep2 = cor(`Serum samples`, `2024-2025`, method = "spearman"),
    p_ep2   = suppressWarnings(
                cor.test(`Serum samples`, `2024-2025`,
                         method = "spearman", exact = FALSE)$p.value),
    tvd_ep1 = sum(abs(`Serum samples` - `2018-2020`)) / 2,
    tvd_ep2 = sum(abs(`Serum samples` - `2024-2025`)) / 2,
    .by = hospital
  ) |>
  mutate(across(starts_with("rho"), \(x) round(x, 2)),
         across(starts_with("tvd"), \(x) round(x, 1)),
         across(starts_with("p_"),  \(x) format.pval(x, digits = 2, eps = 0.001)))
 
agree |> print()
 
# largest district-level discrepancies, which drive the TVD
wide |>
  mutate(d_ep1 = `Serum samples` - `2018-2020`,
         d_ep2 = `Serum samples` - `2024-2025`) |>
  slice_max(abs(d_ep2), n = 3, by = hospital) |>
  select(hospital, district, `Serum samples`, `2024-2025`, d_ep2) |>
  arrange(hospital, desc(abs(d_ep2))) |>
  print(n = Inf)
 
# districts each hospital effectively does not reach
tibble(equal_share = mid, n_districts = n_dist) |> print()
 
tab_pct |>
  summarise(n_below_1pct = sum(pct < 1), n_zero = sum(pct == 0),
            .by = c(hospital, panel)) |>
  arrange(hospital, panel) |>
  print(n = Inf)
 
# pairwise overlap between the three sample catchments
smp <- wide |> select(hospital, district, pct = `Serum samples`)
 
crossing(a = levels(smp$hospital), b = levels(smp$hospital)) |>
  filter(a < b) |>
  mutate(overlap = map2_dbl(a, b, \(x, y) {
    pa <- smp |> filter(hospital == x) |> arrange(district) |> pull(pct)
    pb <- smp |> filter(hospital == y) |> arrange(district) |> pull(pct)
    sum(pmin(pa, pb))          # shared share, 0 = disjoint, 100 = identical
  })) |>
  print()
 
# ---- 16e. Panel B: sample share vs admission share ------------------------
# aspect.ratio = 1 makes the panels square while letting them size to the
# available height, so B fills its cell instead of leaving side gaps.
 
p_agree <- ggplot(sc_dat, aes(pct_adm, pct_smp, colour = region)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", colour = "grey50") +
  geom_point(size = 1.1, alpha = 0.85) +
  geom_text(data = lab_pos, aes(x = pct_adm, y = pct_smp, label = lab),
            inherit.aes = FALSE, colour = "black", size = 2.2,
            hjust = 0, vjust = 1, lineheight = 0.9) +
  facet_grid(hospital ~ panel) +
  scale_colour_manual(values = pal_region) +
  {if (use_log) {
     list(scale_x_log10(labels = \(x) paste0(x, "%"),
                        breaks = c(0.1, 1, 10)),
          scale_y_log10(labels = \(x) paste0(x, "%"),
                        breaks = c(0.1, 1, 10)))
   } else {
     list(scale_x_continuous(labels = \(x) paste0(x, "%")),
          scale_y_continuous(labels = \(x) paste0(x, "%")))
   }} +
  coord_cartesian(xlim = ax_lim, ylim = ax_lim) +
  labs(x = "Share of admissions", y = "Share of serum samples",
       colour = "Region") +
  theme_epi() +
  theme(aspect.ratio  = 1,
        text          = element_text(size = 8),
        axis.title    = element_text(size = 8, face = "bold"),
        axis.text     = element_text(size = 6.5),
        strip.text    = element_text(size = 8),
        panel.spacing = unit(0.25, "lines"))
 
# ---- 16f. Combined figure at 8 x 4.5 in -----------------------------------
# All sizes below are chosen for that canvas. Both panels put their legend
# in a bottom band of similar height so the panel regions line up.
 
p_catch_c <- p_catch +
  guides(fill = guide_colourbar(title.position = "top",
                                barwidth  = unit(3.2, "cm"),
                                barheight = unit(0.22, "cm"),
                                ticks.colour = "white",
                                frame.colour = NA)) +
  theme(legend.position    = "bottom",
        legend.title       = element_text(size = 7.5, hjust = 0.5),
        legend.text        = element_text(size = 6.5),
        legend.box.spacing = unit(0, "pt"),
        legend.margin      = margin(0, 0, 0, 0),
        strip.text.x       = element_text(size = 8, margin = margin(b = 6)),
        strip.text.y.left  = element_text(size = 8, margin = margin(r = 3), face = "bold"),
        panel.spacing.x    = unit(0.9, "lines"),
        panel.spacing.y    = unit(0, "lines"),
        plot.margin        = margin(2, 4, 0, 2))
 
p_agree_c <- p_agree +
  guides(colour = guide_legend(nrow = 2, byrow = TRUE,
                               override.aes = list(size = 2))) +
  theme(strip.text.y       = element_blank(),
        legend.position    = "bottom",
        legend.direction   = "horizontal",
        legend.title       = element_text(size = 7.5, face = "bold"),
        legend.text        = element_text(size = 6.5),
        legend.key.size    = unit(0.55, "lines"),
        legend.box.spacing = unit(0, "pt"),
        legend.margin      = margin(0, 0, 0, 0),
        plot.margin        = margin(2, 2, 0, 4))
 
p_comb <- (p_catch_c | p_agree_c) +
  plot_layout(widths = c(1.2, 1)) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag          = element_text(face = "bold", size = 10),
        plot.tag.position = c(0.01, 0.99),
        plot.margin       = margin(t = 5, r = 4, b = 5, l = 2))
 
p_comb
 
ggsave("figs/chap1/fig_catchment_combined.png", p_comb,
       width = 6, height = 3.8, dpi = 300, bg = "white")
```

```{r}
tab_pct <- map_pct |> st_drop_geometry()
tab_pct |>
  summarise(
    top1     = max(pct),
    top1_d   = district[which.max(pct)],
    top3     = sum(sort(pct, decreasing = TRUE)[1:3]),
    n_to_80  = which(cumsum(sort(pct, decreasing = TRUE)) >= 80)[1],
    n_over_1 = sum(pct >= 1),
    .by = c(hospital, panel)
  ) |>
  arrange(hospital, panel) |>
  print(n = Inf)

agree <- wide |>
  summarise(
    n       = n(),
    rho_ep1 = cor(`Serum samples`, `2018-2020`, method = "spearman"),
    p_ep1   = suppressWarnings(
                cor.test(`Serum samples`, `2018-2020`,
                         method = "spearman", exact = FALSE)$p.value),
    rho_ep2 = cor(`Serum samples`, `2024-2025`, method = "spearman"),
    p_ep2   = suppressWarnings(
                cor.test(`Serum samples`, `2024-2025`,
                         method = "spearman", exact = FALSE)$p.value),
    tvd_ep1 = sum(abs(`Serum samples` - `2018-2020`)) / 2,
    tvd_ep2 = sum(abs(`Serum samples` - `2024-2025`)) / 2,
    .by = hospital
  ) |>
  mutate(across(starts_with("rho"), \(x) round(x, 2)),
         across(starts_with("tvd"), \(x) round(x, 1)),
         across(starts_with("p_"),  \(x) format.pval(x, digits = 2, eps = 0.001)))
 
agree |> print()
```


## Sero of infants

```{r}
df_sero_infant |> 
  # 1. Convert age inline
  mutate(age_month = age * 12) |> 
  ggplot(aes(x = age_month, y = titre)) +
  geom_hline(yintercept = 200, linetype = "dashed") +
  # Added alpha to prevent overplotting with your raw points
  geom_point(alpha = 0.4) + 
  # 2. Fit and plot the GLM in one step
  geom_smooth(
    method = "glm",
    method.args = list(family = Gamma(link = "log")),
    level = 0.95,
    colour = "blue" 
  ) +
  # Vaccination milestones
  geom_vline(xintercept = 9, linetype = "dashed", colour = "#D55E00", linewidth = 0.7) +
  geom_vline(xintercept = 12, linetype = "dashed", colour = "#0072B2", linewidth = 0.7) +
  scale_y_continuous(breaks = c(200, 500, 1000, 1500, 2000)) +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  labs(
    x = "Age (months)",
    y = "Titer (mIU/mL)"
  ) +
  theme_classic()

# ggsave("figs/infant.png", width = 5, height = 2, dpi = 300)
```

```{r}
# saveRDS(df_ep2, "C:/Users/ongph/github/data/measles/cleaned/measles_hcmc_24-25.rds")
```

Ward

```{r}
ward_ch2_cch <- read.xlsx("C:/Users/ongph/github/data/sero-meta/ch2_cch_addr.xlsx", startRow = 6, cols = 1:11) |> 
  janitor::clean_names()

ward_ch2_cch |> 
  count(benh_vien, mau_den_tu_khoa_phong) |> 
  mutate(n_pct = sprintf("%d (%.1f%%)", n, n / sum(n) * 100), .by = benh_vien) |> 
  arrange(benh_vien, desc(n)) |> 
  select(-n)
```

