---
title: "amy"
format: html
---

## Packages

Required packages:

```{r}
required <- c("openxlsx", "janitor", "tidyverse", "stringi")

# Installing those that are not installed yet:
to_install <- required[! required %in% installed.packages()[, "Package"]]
if (length(to_install)) install.packages(to_install)

# Loading the packages for interactive use:
invisible(sapply(required, library, character.only = TRUE))
```

## Functions

Clean Vietnamese district and province name

```{r}
clean_district <- function(x) {
  x |>
    stri_trans_general("Latin-ASCII") |> 
    str_to_lower() |>                      
    str_squish() |>                        
    str_replace_all("đ", "d") |>
    str_remove("^(quan |huyen |thanh pho |tp )")
}

clean_province <- function(x) {
  x |>
    stri_trans_general("Latin-ASCII") |> 
    str_to_lower() |>                      # Fix case mismatches (e.g., Ha Noi vs ha noi)
    str_squish() |>                        # Remove trailing, leading, or double spaces
    str_replace_all("đ", "d")              # Catch the Vietnamese 'đ' if the ASCII conversion misses it
}
```


## Constant

Path to data:

```{r}
path2data <- paste0("C:/Users/ongph/github/data")
```

Files names:

```{r}
# Case notification
cases_file <- "measles/measles_cases_2019-12.2025.xlsx"
```

```{r}
df_case <- file.path(path2data, cases_file) |>
  read.xlsx(sheet = 4) |>
  clean_names() |>
  mutate(
    ngay_nv = as.Date(as.numeric(ngay_nv), origin = "1899-12-30"),
    ngay_sinh = as.Date(as.numeric(ngay_sinh), origin = "1899-12-30"),
    age = time_length(interval(ngay_sinh, ngay_nv), "year"),
    district = clean_district(quan_huyen)
  )
```

What if you pulled the vaccination coverage from the suspected measles cases, but IgM negative? 

The assumption here is that measles IgM negative are representative of the hospital population broadly. If you don’t think it is a terrible assumption, then we can get vaccination coverage among that population by age (and potentially by age and geography) and apply this coverage to the whole of the hospital population.

Lab test negative: not sure IgM or PCR

```{r}
df_neg <- df_case |> 
  filter(ngay_nv >= dmy("01-04-2024") & pl_chan_doan == "Loại trừ sởi") |> 
  mutate(vax = factor(tiem_chung, levels = c("Không", "Có", "Không rõ"), labels = c("No", "Yes", "Unknown")))

table(df_neg$vax)
```

```{r}
df_summary <- df_neg |>
  mutate(age_group = cut(
    age,
    breaks = c(-Inf, 9 / 12, 5, 15, 30, Inf),
    labels = c("<9m", "9m-5", "5-15", "15-30", "30+")
  )) |>
  count(age_group, vax) |>
  # Calculate percentage within each age group
  mutate(pct = (n / sum(n)) * 100, .by = age_group) |>
  # Combine count and percentage into one string
  mutate(n_pct = sprintf("%d (%.1f%%)", n, pct)) |>
  select(age_group, vax, n_pct) |>
  pivot_wider(names_from = vax, values_from = n_pct, values_fill = "0 (0.0%)")

df_summary
```



