---
title: "Hospital-community model"
format: 
  html:
    code-fold: true
---

## Model description

![](img/hosp-com.svg)

- The model represents 2 environments: hospital ($H$) and community ($C$), each with $S$, $I$, and $R$ compartments

- Movement between environments is modelled implicitly through the force of infection: infectious individuals in one environment contribute to transmission in the other environment through a parameter $\rho$

- This allows constant population sizes in hospital and community, while also represents cross-environment infection

$$\begin{aligned}
\frac{dS_H}{dt} &= -\lambda_H S_H, \\
\frac{dI_H}{dt} &= \lambda_H S_H - \gamma_H I_H, \\
\frac{dR_H}{dt} &= \gamma_H I_H,
\end{aligned}
\qquad
\begin{aligned}
\frac{dS_C}{dt} &= -\lambda_C S_C, \\
\frac{dI_C}{dt} &= \lambda_C S_C - \gamma_C I_C, \\
\frac{dR_C}{dt} &= \gamma_C I_C.
\end{aligned}$$

with

$$\lambda_H = \beta_H \left( \frac{I_H}{N_H} + \rho \frac{I_C}{N_C} \right)$$
$$\lambda_C = \beta_C \left( \frac{I_C}{N_C} + \rho \frac{I_H}{N_H} \right)$$

with $N_H$ and $N_C$ are the total population 

$$N_H = S_H + I_H + R_H,
\qquad
N_C = S_C + I_C + R_C$$

The community size is larger than hospital size

$$N_C = a \cdot N_H$$

Effective contact rate in the hospital is larger than in the community

$$\beta_H = b \cdot \beta_C$$

The proportion of initial susceptibility in the hospital is larger than in the community

$$s_{H0} = c \cdot s_{C0}$$

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(deSolve)
```

## Model definition

```{r}
# ------------------------------------------------------------
# 1) TWO-ENVIRONMENT SIR MODEL
# ------------------------------------------------------------
sir_2env <- function(time, y, parms) {
  with(as.list(c(y, parms)), {
    NH <- SH + IH + RH
    NC <- SC + IC + RC

    lambda_H <- beta_H * (IH / NH + rho * IC / NC)
    lambda_C <- beta_C * (IC / NC + rho * IH / NH)

    dSH <- -lambda_H * SH
    dIH <-  lambda_H * SH - gamma_H * IH
    dRH <-  gamma_H * IH

    dSC <- -lambda_C * SC
    dIC <-  lambda_C * SC - gamma_C * IC
    dRC <-  gamma_C * IC

    list(c(dSH, dIH, dRH, dSC, dIC, dRC))
  })
}

# ------------------------------------------------------------
# 2) STOPPING RULE
# ------------------------------------------------------------
make_stop_condition <- function(threshold = 1) {
  function(time, y, parms) {
    # Force the solver to ignore the threshold for the first 1 day
    if (time < 1) return(1) 
    as.numeric(y["IH"] + y["IC"] - threshold)
  }
}

no_change_event <- function(time, y, parms) {
  y
}

# ------------------------------------------------------------
# 3) ONE-SIMULATION WRAPPER
# ------------------------------------------------------------
run_one_sim <- function(a, b, c, rho,
                        NH = 100,
                        R0_C = 18,
                        infectious_period = 7,
                        infectious_period_H = infectious_period,
                        infectious_period_C = infectious_period,
                        sC_prop = 0.05,
                        IH0 = 0,
                        IC0 = 1,
                        max_time = 365,
                        dt = 0.1,
                        stop_threshold = 1,
                        method = "lsoda",
                        use_root_stop = TRUE,
                        return_trajectory = TRUE) {

  # Separate recovery rates
  # For now they are equal by default because both periods default to infectious_period
  gamma_H <- 1 / infectious_period_H
  gamma_C <- 1 / infectious_period_C

  # Transmission rates
  # Community R0 is specified directly
  beta_C <- R0_C * gamma_C

  # Hospital transmission is scaled relative to community
  beta_H <- b * beta_C

  # Population sizes
  NC <- a * NH

  # Initial susceptible fraction in hospital
  # SH0/NH = c * (SC0/NC)
  sH_prop <- c * sC_prop

  if (sC_prop < 0 || sC_prop > 1 || sH_prop < 0 || sH_prop >= 1) {
    out_row <- data.frame(
      a = a, b = b, c = c, rho = rho,
      NH = NH, NC = NC,
      R0_C = R0_C,
      infectious_period_H = infectious_period_H,
      infectious_period_C = infectious_period_C,
      beta_H = beta_H, beta_C = beta_C,
      gamma_H = gamma_H, gamma_C = gamma_C,
      sC_prop = sC_prop, sH_prop = sH_prop,
      outbreak_size = NA_real_,
      attack_rate_total = NA_real_,
      attack_rate_H = NA_real_,
      attack_rate_C = NA_real_,
      peak_I_total = NA_real_,
      peak_IH = NA_real_,
      peak_IC = NA_real_,
      time_to_peak_total = NA_real_,
      end_time = NA_real_,
      status = "infeasible: invalid susceptible proportions"
    )
    if (return_trajectory) return(list(summary = out_row, trajectory = NULL))
    return(out_row)
  }

  SC0 <- sC_prop * NC
  SH0 <- sH_prop * NH
  RC0 <- NC - SC0 - IC0
  RH0 <- NH - SH0 - IH0

  if (any(c(SH0, IH0, RH0, SC0, IC0, RC0) < 0)) {
    out_row <- data.frame(
      a = a, b = b, c = c, rho = rho,
      NH = NH, NC = NC,
      R0_C = R0_C,
      infectious_period_H = infectious_period_H,
      infectious_period_C = infectious_period_C,
      beta_H = beta_H, beta_C = beta_C,
      gamma_H = gamma_H, gamma_C = gamma_C,
      sC_prop = sC_prop, sH_prop = sH_prop,
      outbreak_size = NA_real_,
      attack_rate_total = NA_real_,
      attack_rate_H = NA_real_,
      attack_rate_C = NA_real_,
      peak_I_total = NA_real_,
      peak_IH = NA_real_,
      peak_IC = NA_real_,
      time_to_peak_total = NA_real_,
      end_time = NA_real_,
      status = "infeasible: negative initial compartment"
    )
    if (return_trajectory) return(list(summary = out_row, trajectory = NULL))
    return(out_row)
  }

  y0 <- c(
    SH = SH0, IH = IH0, RH = RH0,
    SC = SC0, IC = IC0, RC = RC0
  )

  parms <- c(
    beta_H = beta_H,
    beta_C = beta_C,
    gamma_H = gamma_H,
    gamma_C = gamma_C,
    rho = rho
  )

  times <- seq(0, max_time, by = dt)

  use_root_stop <- isTRUE(use_root_stop) && method %in% c("lsoda", "lsode", "lsodes", "radau")

  if (use_root_stop) {
    out <- ode(
      y = y0,
      times = times,
      func = sir_2env,
      parms = parms,
      rootfun = make_stop_condition(stop_threshold),
      events = list(func = no_change_event, root = TRUE, terminalroot = 1),
      method = method
    )
    out <- as.data.frame(out)
    status <- if (tail(out$time, 1) < max_time) "stopped" else "max_time_reached"
  } else {
    out <- ode(
      y = y0,
      times = times,
      func = sir_2env,
      parms = parms,
      method = method
    )
    out <- as.data.frame(out)

    idx_stop <- which((out$IH + out$IC) < stop_threshold)
    if (length(idx_stop) > 0) {
      out <- out[1:min(idx_stop), , drop = FALSE]
      status <- "stopped_posthoc"
    } else {
      status <- "max_time_reached"
    }
  }

  out$NH <- out$SH + out$IH + out$RH
  out$NC <- out$SC + out$IC + out$RC
  out$I_total <- out$IH + out$IC

  last <- out[nrow(out), ]

  new_inf_H <- SH0 - last$SH
  new_inf_C <- SC0 - last$SC
  outbreak_size <- new_inf_H + new_inf_C + IH0 + IC0

  summary_row <- data.frame(
    a = a, b = b, c = c, rho = rho,
    NH = NH, NC = NC,
    R0_C = R0_C,
    infectious_period_H = infectious_period_H,
    infectious_period_C = infectious_period_C,
    beta_H = beta_H, beta_C = beta_C,
    gamma_H = gamma_H, gamma_C = gamma_C,
    sC_prop = sC_prop, sH_prop = sH_prop,
    outbreak_size = outbreak_size,
    attack_rate_total = outbreak_size / (NH + NC),
    attack_rate_H = (new_inf_H + IH0) / NH,
    attack_rate_C = (new_inf_C + IC0) / NC,
    peak_I_total = max(out$I_total),
    peak_IH = max(out$IH),
    peak_IC = max(out$IC),
    time_to_peak_total = out$time[which.max(out$I_total)][1],
    end_time = last$time,
    status = status
  )

  if (return_trajectory) return(list(summary = summary_row, trajectory = out))
  summary_row
}

# For sensitivity analysis
get_outbreak_size <- function(a, b, c, rho,
                              NH = 5000,
                              R0_C = 18,
                              infectious_period = 7,
                              infectious_period_H = infectious_period,
                              infectious_period_C = infectious_period,
                              sC_prop = 0.05,
                              IH0 = 0,
                              IC0 = 1,
                              max_time = 365,
                              dt = 0.1,
                              stop_threshold = 1,
                              method = "lsoda") {

  gamma_H <- 1 / infectious_period_H
  gamma_C <- 1 / infectious_period_C
  beta_C <- R0_C * gamma_C
  beta_H <- b * beta_C
  
  NC <- a * NH
  sH_prop <- c * sC_prop

  # Fast fail for invalid proportions
  if (sC_prop < 0 || sC_prop > 1 || sH_prop < 0 || sH_prop >= 1) {
    return(NA_real_)
  }

  SC0 <- sC_prop * NC
  SH0 <- sH_prop * NH
  RC0 <- NC - SC0 - IC0
  RH0 <- NH - SH0 - IH0

  # Fast fail for negative compartments
  if (any(c(SH0, IH0, RH0, SC0, IC0, RC0) < 0)) {
    return(NA_real_)
  }

  y0 <- c(
    SH = SH0, IH = IH0, RH = RH0,
    SC = SC0, IC = IC0, RC = RC0
  )

  parms <- c(
    beta_H = beta_H,
    beta_C = beta_C,
    gamma_H = gamma_H,
    gamma_C = gamma_C,
    rho = rho
  )

  times <- seq(0, max_time, by = dt)

  out <- ode(
    y = y0,
    times = times,
    func = sir_2env,
    parms = parms,
    rootfun = make_stop_condition(stop_threshold),
    events = list(func = no_change_event, root = TRUE, terminalroot = 1),
    method = method
  )

  # Extract the final row directly from the matrix output
  last_row <- out[nrow(out), ]

  new_inf_H <- SH0 - last_row["SH"]
  new_inf_C <- SC0 - last_row["SC"]
  
  # Strip names so pmap_dbl doesn't throw a warning
  unname(new_inf_H + new_inf_C + IH0 + IC0)
}
```

## Stress test the model

```{r}
plot_sim <- function(out, main_title = "") {
  # Reshape from wide to long for ggplot
  plot_data <- out |>
    select(time, IH, IC, I_total) |>
    pivot_longer(
      cols = -time,
      names_to = "compartment",
      values_to = "count"
    ) |>
    # Lock the factor levels so the legend stays in a logical order
    mutate(
      compartment = factor(compartment, levels = c("IH", "IC", "I_total"))
    )
  
  ggplot(plot_data, aes(x = time, y = count, color = compartment, linetype = compartment)) +
    geom_line(linewidth = 1) +
    # Match your original line types: solid for specific, dashed for total
    scale_linetype_manual(values = c("IH" = "solid", "IC" = "solid", "I_total" = "dashed")) +
    labs(
      title = main_title,
      x = "Time",
      y = "Number infectious",
      color = NULL,
      linetype = NULL
    ) +
    theme_minimal() +
    theme(
      legend.position = "top",
      legend.justification = "right"
    )
}
```

### If the stopping condition works

```{r}
# 1. Run a fast-burning simulation
# High R0 ensures the outbreak peaks and dies out quickly
test_sim <- run_one_sim(
  a = 10, b = 2, c = 1, rho = 0.1,
  IC0 = 2,
  R0_C = 8, 
  infectious_period_H = 5,
  infectious_period_C = 5,
  max_time = 365,
  stop_threshold = 1,
  use_root_stop = TRUE,
  return_trajectory = TRUE
)

# 2. Extract the data
traj <- test_sim$trajectory
summary_data <- test_sim$summary

plot_sim(traj)
```


### `rho = 0`

Outbreak in community won't affect hospital

```{r}
sim1 <- run_one_sim(
  a = 100, b = 2, c = 1.5, rho = 0,
  IH0 = 0, IC0 = 2
)

plot_sim(sim1$trajectory, "rho = 0")
```

### `rho = 0.001`

```{r}
sim2 <- run_one_sim(
  a = 100, b = 2, c = 1.5, rho = 0.001,
  IH0 = 0, IC0 = 2
)

plot_sim(sim2$trajectory, "rho = 0.001")
```

## Sensitivity analysis

The model is run with $N_H = 5000$, using measles $R_{0C} = 18$ in the community. Infectious period is set as 7 days ($\gamma_C = \frac{1}{7}$ therefore $\beta_C = R_{0C} \cdot \gamma_C = \frac{18}{7}$. Assuming infectious period of hospital-infected cases is the same as community-infected ($\gamma_H = \gamma_C$). Susceptibility fraction in the community is $s_{C0} = 0.05$ which is 95% measles vaccination coverage. The model stops (outbreak ends) when $I_H + I_C \leq 1$.

Sensitivity analysis is run on a grid of:

- $a$ from 100-150 (community size $N_C$ from 50,000-750,000)
- $b$ from 1-10 ($\beta_H$ is 1-10 times $\beta_C$)
- $c$ from 1-10 (susceptibility fraction in the hospital from 0.05-0.5)
- $rho$ from 0.001-0.01

```{r}
# grid <- expand_grid(
#   a = seq(100, 150, 10),
#   b = 1:10,
#   c = 1:10,
#   rho = seq(0.001, 0.01, 0.001)
# )
# 
# # Run the simulations
# results <- grid |>
#   mutate(
#     outbreak_size = pmap_dbl(
#       list(a = a, b = b, c = c, rho = rho), 
#       get_outbreak_size
#     )
#   )
# saveRDS(results, "data/sens.rds")
results <- readRDS("data/sens.rds")

# Assuming 'results' is your raw grid output
plot_data <- results |>
  filter(!is.na(outbreak_size)) |>
  mutate(
    a_label = a,
    rho_label = rho
  )

# 1. Remove factor() from x and y
# 2. Swap geom_tile for geom_raster
ggplot(plot_data, aes(x = b, y = c, fill = outbreak_size)) +
  geom_raster() + 
  # Use a perceptually uniform colour scale (viridis/magma) for better contrast
  scale_fill_viridis_c(option = "viridis", na.value = "transparent") + 
  facet_grid(rho_label ~ a_label) +
  labs(
    x = "Relative hospital transmission (b)",
    y = "Relative hospital susceptible fraction (c)",
    fill = "Total cases"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    strip.text = element_text(face = "bold", size = 10),
    # Angle x-axis text slightly if you still have overlapping numbers
    axis.text.x = element_text(angle = 90, hjust = 1) 
  )
```

### Parameter importance

Use partial rank correlation coefficient (PRCC):

```{r, warning=FALSE, message=FALSE}
library(sensitivity)

# 1. Isolate inputs and outputs from your successful runs
model_data <- results |>
  filter(!is.na(outbreak_size))

params <- model_data |> select(a, b, c, rho)
output <- model_data$outbreak_size

# 2. Calculate Partial Rank Correlation Coefficients
# rank = TRUE converts the raw values to ranks before calculating correlations
prcc_res <- pcc(X = params, y = output, rank = TRUE)

# 3. Format the output for ggplot2
prcc_plot_data <- prcc_res$PRCC |>
  rownames_to_column(var = "parameter") |>
  rename(prcc_value = original) |>
  # Reorder factors by absolute importance to create a neat tornado shape
  mutate(parameter = fct_reorder(parameter, abs(prcc_value)))

# 4. Plot the tornado chart
ggplot(prcc_plot_data, aes(x = prcc_value, y = parameter)) +
  geom_col() +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1) +
  # PRCC is bounded strictly between -1 and 1
  scale_x_continuous(limits = c(-1, 1)) +
  labs(
    title = "Parameter importance on outbreak size",
    subtitle = "Partial Rank Correlation Coefficients (PRCC)",
    x = "PRCC (Magnitude = importance, Sign = direction of effect)",
    y = NULL
  ) +
  theme_minimal()
```

## Surrogate modelling

```{r, message=FALSE, warning=FALSE}
library(mgcv)
```

```{r}
# 1. Clean the data
# The GAM cannot handle NA values from infeasible ODE runs
model_data <- results |>
  filter(!is.na(outbreak_size))

# 2. Fit the tensor product GAM
# We use family = gaussian(link = "identity") because the data is deterministic continuous.
# k sets the basis dimension per parameter. Keep it low (e.g., 4 or 5) initially 
# to avoid an explosion in degrees of freedom (total df = k^4).
# surrogate_gam <- gam(
#   outbreak_size ~ te(a, b, c, rho, k = 5),
#   data = model_data,
#   family = gaussian(link = "identity"),
#   method = "REML" # REML is standard for penalised smoothers to prevent overfitting
# )
# saveRDS(surrogate_gam, "data/surrogate_gam.rds")
surrogate_gam <- readRDS("data/surrogate_gam.rds")

# 3. Check the mechanical fit
# This will output the R-squared (deviance explained). 
# For a deterministic surrogate, you want this > 95%.
summary(surrogate_gam)

# Create a prediction sequence for 'rho'
pred_grid <- tibble(
  rho = seq(min(model_data$rho), max(model_data$rho), length.out = 100),
  a = 100,
  b = 1,
  c = 1
)

# Generate predictions and standard errors
predictions <- bind_cols(
  pred_grid,
  as_tibble(predict(surrogate_gam, newdata = pred_grid, se.fit = TRUE))
) |>
  mutate(
    lower = fit - (1.96 * se.fit),
    upper = fit + (1.96 * se.fit)
  )

# Plot the marginal effect of rho
ggplot(predictions, aes(x = rho, y = fit)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
  geom_line(linewidth = 1) +
  labs(
    title = "Marginal effect of cross-transmission (rho) on outbreak size",
    subtitle = "When a=100, b=1, c=1",
    x = "Mixing fraction (rho)",
    y = "Predicted outbreak size"
  ) +
  theme_minimal()
```



